---
title: "Experiment 12: Rare-Disagreement Intervals"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/archive/pre-redesign/12-quadratic-rare-disagreement/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

method_order <- c(
  "moss_fisher", "delta2_fisher",
  "garwood_plug", "garwood_chance",
  "jeffreys_plug", "jeffreys_chance",
  "anscombe_plug", "anscombe_chance"
)
summary$method <- factor(summary$method, levels = method_order)
focus <- summary[summary$coefficient %in% c("Conger", "Fleiss"), ]
focus <- focus[order(focus$target, focus$coefficient, focus$R, focus$n, focus$method), ]

overall <- aggregate(
  cbind(coverage, miss_below, miss_above, mean_length, zero_disagreement_rate) ~ target + method,
  data = focus,
  FUN = mean
)
overall <- overall[order(overall$target, overall$method), ]
overall_value <- function(target, method, column) {
  overall[overall$target == target & as.character(overall$method) == method, column][[1]]
}
cell_value <- function(target, R, n, coefficient, method, column) {
  focus[
    focus$target == target &
      focus$R == R &
      focus$n == n &
      focus$coefficient == coefficient &
      as.character(focus$method) == method,
    column
  ][[1]]
}
overall_tab <- data.frame(
  Target = overall$target,
  Method = overall$method,
  Coverage = fmt(overall$coverage),
  `Miss Below` = fmt(overall$miss_below),
  `Miss Above` = fmt(overall$miss_above),
  `Mean Length` = fmt(overall$mean_length),
  `Zero Disagree` = fmt(overall$zero_disagreement_rate),
  check.names = FALSE
)

lattice_diag <- aggregate(
  cbind(zero_disagreement_rate, mean_loss_units, mean_chance_disagreement) ~ target + R + n,
  data = focus[focus$method == "moss_fisher" & focus$coefficient == "Conger", ],
  FUN = mean
)
lattice_diag <- lattice_diag[order(lattice_diag$target, lattice_diag$R, lattice_diag$n), ]
lattice_tab <- data.frame(
  Target = lattice_diag$target,
  R = lattice_diag$R,
  n = lattice_diag$n,
  `Zero Disagree` = fmt(lattice_diag$zero_disagreement_rate),
  `Mean Loss Units` = fmt(lattice_diag$mean_loss_units, 1),
  `Mean Chance Disagree` = fmt(lattice_diag$mean_chance_disagreement),
  check.names = FALSE
)

worst <- focus[order(focus$coverage), ][1, ]
worst_tab <- head(focus[order(focus$coverage), ], 14)
worst_tab <- data.frame(
  Target = worst_tab$target,
  Coefficient = worst_tab$coefficient,
  R = worst_tab$R,
  n = worst_tab$n,
  Method = worst_tab$method,
  Coverage = fmt(worst_tab$coverage),
  `Miss Below` = fmt(worst_tab$miss_below),
  `Miss Above` = fmt(worst_tab$miss_above),
  `Mean Length` = fmt(worst_tab$mean_length),
  check.names = FALSE
)

best_by_cell <- do.call(rbind, lapply(split(
  focus,
  interaction(focus$target, focus$coefficient, focus$R, focus$n)
), function(d) {
  d[which.min(abs(d$coverage - 0.95)), ]
}))
best_tab <- data.frame(
  Target = best_by_cell$target,
  Coefficient = best_by_cell$coefficient,
  R = best_by_cell$R,
  n = best_by_cell$n,
  Method = best_by_cell$method,
  Coverage = fmt(best_by_cell$coverage),
  `Mean Length` = fmt(best_by_cell$mean_length),
  check.names = FALSE
)
```

## Question

Can a boundary-aware rare-disagreement interval repair the high-agreement quadratic-kappa coverage failures without bootstrap?

## Short Answer

The smooth delta expansion from Experiment 11 was too small because it still collapses when observed disagreement is zero. This experiment attacks that corner directly: it treats the observed quadratic disagreement numerator as a lattice count and uses standard Poisson count intervals before mapping back through `theta = 1 - A / B`.

The boundary correction helps exactly where it should, but it is not a general fix. In the pathological target `0.95`, `R = 2`, `n = 10` Conger cell, Moss Fisher covers only `r fmt(cell_value(0.95, 2, 10, "Conger", "moss_fisher", "coverage"))`, while Garwood with a plug-in chance denominator covers `r fmt(cell_value(0.95, 2, 10, "Conger", "garwood_plug", "coverage"))` and Garwood with chance-denominator widening covers `r fmt(cell_value(0.95, 2, 10, "Conger", "garwood_chance", "coverage"))`. That is the zero-disagreement spike being fixed.

But averaged over the grid, the count intervals are worse than Fisher/delta Fisher. At target `0.95`, average coverage is `r fmt(overall_value(0.95, "moss_fisher", "coverage"))` for Moss Fisher, `r fmt(overall_value(0.95, "delta2_fisher", "coverage"))` for smooth delta Fisher, and only `r fmt(overall_value(0.95, "garwood_chance", "coverage"))` for Garwood with chance widening. The pairwise loss-unit count is too clustered to behave like a simple Poisson count, especially for `R = 5`.

## Evidence

`Miss Below` means the interval upper endpoint is below the true coefficient; `Miss Above` means the lower endpoint is above it.

Average over coefficients, rater counts, and sample sizes:

```{r}
knitr::kable(overall_tab, align = "rlrrrrr")
```

Lattice diagnostics for Conger:

```{r}
knitr::kable(lattice_tab, align = "rrrrrr")
```

Worst cells:

```{r}
knitr::kable(worst_tab, align = "rlrrlrrrr")
```

Closest-to-nominal method by cell:

```{r}
knitr::kable(best_tab, align = "rlrrlrr")
```

## Methods

For five equally spaced categories `-2,-1,0,1,2`, the normalized quadratic loss is `(x-y)^2 / 16`. The observed disagreement numerator can therefore be written as a scaled nonnegative integer:

```text
A = M / (16 * choose(R, 2) * n),
```

where `M` is the sum of squared category-distance units across all item-rater pairs. Near perfect agreement, `M` is small and often zero. The usual influence variance for `A` is then zero, producing degenerate upper-boundary intervals.

The count methods form a count interval for `M`, convert it to an interval for `A`, and then map it through `theta = 1 - A / B`. The `plug` versions hold the chance disagreement `B` fixed at its estimate. The `chance` versions also use a t interval for `B` from its influence function.

The count intervals are:

- `garwood`: exact Poisson count interval.
- `jeffreys`: Jeffreys-prior Poisson count interval.
- `anscombe`: Poisson variance-stabilizing square-root interval.

The baselines are Moss's Fisher transform and the Fisher interval with the second-order delta variance from Experiment 11.

## Caveats

The observed-disagreement count is not literally Poisson. It is bounded, weighted, and generated by clustered raters within items. This is a rare-event boundary approximation, not a final theory. Its job is to test whether the zero-disagreement lattice is the source of the worst coverage failures.

## Reproduce

Smoke run:

```sh
Rscript experiments/archive/pre-redesign/12-quadratic-rare-disagreement/run_experiment.R --smoke
```

Exploratory run used here:

```sh
Rscript experiments/archive/pre-redesign/12-quadratic-rare-disagreement/run_experiment.R --reps 500 --target-grid 0.8,0.9,0.95 --n-grid 10,20,40 --r-grid 2,5 --progress
```

This report used n grid `r meta[["n_grid"]]`, R grid `r meta[["r_grid"]]`, reps `r meta[["reps"]]`, target grid `r meta[["target_grid"]]`, and seed base `r meta[["seed_base"]]`.
