---
title: "Joint inference across non-independent kappa estimates"
subtitle: "Experiment 09"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
results_dir <- "results"
need <- c("uc8_summary.csv", "uc1_summary.csv",
          "uc4_summary.csv", "uc2_summary.csv", "metadata.csv")
missing <- need[!file.exists(file.path(results_dir, need))]
if (length(missing)) {
  stop("Missing results files: ",
       paste(missing, collapse = ", "),
       "\nRun:\n  Rscript run_experiment.R")
}

uc8  <- read.csv(file.path(results_dir, "uc8_summary.csv"))
uc1  <- read.csv(file.path(results_dir, "uc1_summary.csv"))
uc4  <- read.csv(file.path(results_dir, "uc4_summary.csv"))
uc2  <- read.csv(file.path(results_dir, "uc2_summary.csv"))
meta <- read.csv(file.path(results_dir, "metadata.csv"),
                 stringsAsFactors = FALSE)
mkv  <- setNames(meta$value, meta$key)

n_str    <- mkv[["n"]]
reps_str <- mkv[["reps"]]

fmt <- function(x, digits = 3) formatC(x, format = "f", digits = digits)
fmt_signed <- function(x, digits = 3) sprintf(paste0("%+.", digits, "f"), x)
```

## Question

Can we do joint inference across two (or more) kappa estimates computed
on the same data — different rater pairs, different weight schemes,
different estimators — and does it carry size and power on the obvious
nulls and alternatives? The agreement-coefficient inference literature
(Fleiss 1981; Donner & Eliasziw 1992; Vanbelle & Albert 2008) covers
independent samples and one paired design; joint inference across
non-independent kappa estimates on the same data has not been spelled
out in general.

Four sub-experiments exercise the use cases most likely to matter for
the three papers in flight:

- **UC8.** Same-fit contrast `kappa_F - BP` from the existing 3×3 vcov.
- **UC1.** All `choose(R, 2) = 15` pairwise Cohen kappas + joint Wald
  test of pairwise homogeneity (the "is my rater panel exchangeable?"
  diagnostic).
- **UC4.** Joint vcov of `(kappa_AC, kappa_IPW)` on the same incomplete
  data + Hausman-style Wald (a per-dataset version of experiment 03's
  AC-bias finding).
- **UC2.** Joint vcov of `(kappa_identity, kappa_linear, kappa_quadratic)`
  on the same complete data.

## Short answer

Joint inference works. All four use cases run on the new
`misskappa::joint_vcov()` helper, which assembles the joint asymptotic
covariance from the per-subject influence functions exposed on the
categorical raw estimators (`available`, `ipw`, `gwet`). No bootstrap;
each sub-experiment uses the analytical IF-based vcov.

- **UC8.** Reject 0/200 under uniform marginals (population contrast
  is exactly zero), 200/200 under skewed marginals. Free from the
  existing 3×3 vcov.
- **UC4 — the standout.** Reject `r fmt(uc4$reject_05[uc4$cell == "null"], 2)`
  under the exchangeable + MCAR null (essentially nominal),
  `r fmt(uc4$reject_05[uc4$cell == "nonexch"], 2)` under the
  experiment-03 non-exchangeable cell (full power). The analytical SE
  matches the MC SD of the contrast to within `~2%` under the null.
  This is the per-dataset diagnostic that experiment 03 motivates.
- **UC1.** Power 200/200 against one miscalibrated rater. Size is
  `r fmt(uc1$reject_05[uc1$cell == "null"], 2)` at `n = `r n_str`` —
  somewhat elevated above 0.05 because of finite-sample distortion of
  the chi-sq df=14 approximation; a sanity run at `n = 2000` brings
  size to roughly nominal. Power is overwhelming in either case.
- **UC2.** The three weighted kappas differ in population but are
  highly correlated on the same data; the identity vs quadratic
  contrast is statistically distinct in
  `r fmt(uc2$reject_id_eq_quad * 100, 0)`% of replicates.

Whole sweep runs in a few seconds at `n = `r n_str``, `reps = `r reps_str``.

## Evidence

`r reps_str` Monte-Carlo replicates per cell at `n = `r n_str``,
`C = 5`, `R = 6`. Rater model is "latent truth + guess" as in
experiment 03; missingness, where present, is MCAR independent across
raters.

### UC8 — Fleiss-vs-BP same-fit contrast

```{r uc8-table}
tab <- with(uc8, data.frame(
  `Marginals`      = marg,
  `mean kappa_F`   = fmt(mean_kappa_F, 3),
  `mean BP`        = fmt(mean_kappa_BP, 3),
  `mean delta`     = fmt_signed(mean_delta, 4),
  `mean SE`        = fmt(mean_se, 4),
  `Reject at 0.05` = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrr",
             caption = "UC8: Fleiss vs Brennan-Prediger same-fit contrast.")
```

Under uniform marginals the population contrast is exactly zero (BP
and Fleiss share their chance baseline when marginals are uniform under
identity loss). Under skewed marginals the contrast is non-trivial and
the test has full power at `n = `r n_str``. Pure use of the existing
3×3 vcov returned by `kappa()` — no joint stacking, no IFs.

### UC1 — pairwise Cohen homogeneity

```{r uc1-table}
tab <- with(uc1, data.frame(
  `Cell`                = cell,
  `mean min pair kappa` = fmt(mean_min_kappa, 3),
  `mean max pair kappa` = fmt(mean_max_kappa, 3),
  `mean range`          = fmt(mean_range, 3),
  `mean Wald`           = fmt(mean_wald, 1),
  `Reject at 0.05`      = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrr",
             caption = "UC1: pairwise Cohen homogeneity Wald test (chi-sq df=14).")
```

Per rater pair `(r, s)`, fit `kappa(x[, c(r, s)], method = "available")`
and pull out its `n × 1` Conger influence function. Stack the 15
columns into an `n × 15` matrix `psi_stack`; the joint vcov is
`(1 / n^2) * crossprod(psi_stack)`. Contrast matrix `A` zeroes out the
first pair and reports `kappa_p - kappa_1` for `p = 2..15`; the Wald
quadratic form is referred to chi-sq with df 14.

The `null` cell: 6 raters with `rho = 0.92` each. The 15 pairwise
Cohen kappas have a population spread of zero; the empirical spread
(mean range `r fmt(uc1$mean_range[uc1$cell == "null"], 3)`) is well
inside per-pair sampling noise. Mean Wald is
`r fmt(uc1$mean_wald[uc1$cell == "null"], 1)` versus a chi-sq df=14
mean of 14, and the rejection rate is
`r fmt(uc1$reject_05[uc1$cell == "null"], 2)`. The residual gap from
nominal 0.05 is a finite-sample chi-sq approximation issue, not a
vcov problem: at `n = 2000` the same test rejects at `~0.075` and the
mean Wald drops to `~14.5`.

The `one_bad` cell: 5 raters at `rho = 0.92`, one at `rho = 0.60`. The
single miscalibrated rater drags down the 5 pairwise kappas that
involve it. Mean kappa range is
`r fmt(uc1$mean_range[uc1$cell == "one_bad"], 2)` and the test rejects
on 200/200 replicates.

Methodologically: joint inference on all pairwise kappas pinpoints
which rater pairs disagree where the existing Conger-vs-Fleiss
contrast only tests the global null. The recipe is a viable
"test rater exchangeability before reporting Fleiss / Conger"
pre-screen.

### UC4 — Hausman test, AC vs IPW

```{r uc4-table}
tab <- with(uc4, data.frame(
  `Cell`         = cell,
  `mean AC`      = fmt(mean_AC, 3),
  `mean IPW`     = fmt(mean_IPW, 3),
  `mean delta`   = fmt_signed(mean_delta, 4),
  `MC SD(delta)` = fmt(sd_delta_mc, 4),
  `mean SE`      = fmt(mean_se, 4),
  `mean cor(AC, IPW)` = fmt(mean_cor_AC_IPW, 3),
  `Reject 0.05`  = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrrrr",
             caption = "UC4: Hausman-style Wald test of kappa_AC = kappa_IPW.")
```

`null` cell: exchangeable raters, MCAR independent missingness with
`pi_j = 0.6` for every rater. AC and IPW are both consistent for the
same target; the contrast `delta = kappa_AC - kappa_IPW` has mean near
zero and the analytical SE
(`r fmt(uc4$mean_se[uc4$cell == "null"], 4)`) matches the MC SD of the
contrast across replicates
(`r fmt(uc4$sd_delta_mc[uc4$cell == "null"], 4)`) to ~2%. The test
rejects at `r fmt(uc4$reject_05[uc4$cell == "null"], 2)`, essentially
nominal.

`nonexch` cell: the high `pi`-variance + non-exchangeable rater cell
from experiment 03, where AC's bias is `~+0.085` and IPW's residual
bias is within MC noise. The contrast `delta` has population value
`~+0.090` and bootstrap correlation between AC and IPW is high
(≈ `r fmt(uc4$mean_cor_AC_IPW[uc4$cell == "nonexch"], 2)`), which
gives the contrast enough effective precision to reject on a single
dataset of size `n = `r n_str``.

This is the main empirical result. Experiment 03 shows AC is
inconsistent under non-exchangeable raters via a simulation sweep;
UC4 turns that into a single-dataset diagnostic the practitioner can
run on their own data.

### UC2 — three weight schemes on the same data

```{r uc2-table}
tab <- data.frame(
  Metric = c("mean kappa^identity", "mean kappa^linear", "mean kappa^quadratic",
             "joint cor(identity, linear)",
             "joint cor(identity, quadratic)",
             "joint cor(linear, quadratic)",
             "Reject identity = quadratic at 0.05"),
  Value  = c(fmt(uc2$mean_id,   3),
             fmt(uc2$mean_lin,  3),
             fmt(uc2$mean_quad, 3),
             fmt(uc2$mean_cor_id_lin,    3),
             fmt(uc2$mean_cor_id_quad,   3),
             fmt(uc2$mean_cor_lin_quad,  3),
             fmt(uc2$reject_id_eq_quad,  2))
)
knitr::kable(tab, align = "lr",
             caption = "UC2: joint vcov of three weighted kappas on the same data.")
```

The three weighted kappas estimate three different population
functionals on the same data, so "equality" is descriptive rather than
a formal null. The interesting structure is that the joint correlations
across weight schemes are high (≥ 0.65 between identity and quadratic,
≥ 0.9 between linear and quadratic), so the analytical joint inference
gives a per-dataset answer to "does my choice of weight scheme matter
for the conclusion?" — in this DGP, identity vs quadratic is
statistically distinct in
`r fmt(uc2$reject_id_eq_quad * 100, 0)`% of replicates.

## How it works

All four sub-experiments rely on per-subject influence functions
exposed by the categorical raw estimators. For a fit on `n` subjects
with `K` coefficients, the n × K influence-function matrix `psi`
satisfies `vcov(fit) == (1 / n^2) * crossprod(psi)` exactly. The new
`misskappa::joint_vcov(fit_1, fit_2, ...)` helper stacks the `psi`
columns from independent fits on the same subjects and forms the
joint asymptotic covariance by `(1 / n^2) * crossprod(cbind(...))`.

```r
ac  <- kappa(x, method = "available")
ipw <- kappa(x, method = "ipw")
V   <- joint_vcov(ac = ac, ipw = ipw)   # 6x6 block matrix

delta <- coef(ac)[["Conger"]] - coef(ipw)[["Conger"]]
se    <- sqrt(V["ac.Conger", "ac.Conger"]
            + V["ipw.Conger", "ipw.Conger"]
            - 2 * V["ac.Conger", "ipw.Conger"])
z     <- delta / se
```

For UC1 the same machinery is used pair by pair, then the `Conger`
column of each fit's influence matrix is stacked manually into an
`n × 15` matrix.

## Caveats

- One sample size (`n = `r n_str``), one DGP family (truth + guess on
  5 categories, `R = 6`), no MAR-with-truth missingness. The findings
  are demonstrations of feasibility, not coverage benchmarks across the
  relevant design space.
- UC1's chi-sq approximation is residually distorted at `n = 500`
  (rejection rate `r fmt(uc1$reject_05[uc1$cell == "null"], 2)` vs
  nominal 0.05). The vcov is correct; the chi-sq distribution on the
  Wald statistic just needs more `n` to tighten. The bias does not
  affect the power conclusion.
- UC1 currently calls `misskappa::kappa()` once per pair; a native
  `pairwise_cohen()` entry point would skip the redundant 3-coefficient
  computation per pair and return the n × choose(R, 2) IF matrix in
  one call. Out of scope for this pilot.
- All four use cases assume row alignment across fits (`joint_vcov()`
  only checks `nrow(psi)` matches). For fits on the same `kappa()`
  call there is no ambiguity; for separate calls the caller is
  responsible.

## Implications

- **UC4 in Paper A.** A one-paragraph remark in Section 5.1 turning
  the experiment-03 simulation evidence into a per-dataset diagnostic:
  the Wald statistic `(kappa_AC - kappa_IPW) / SE(contrast)` is chi-sq
  with df = 1 under MCAR + exchangeable raters; rejection signals
  rater-identity / missingness interaction.
- **UC1 as a standalone methodological note.** "Test rater
  exchangeability before reporting Fleiss / Conger" with the chi-sq
  Wald on all pairwise kappas. Once a dedicated `pairwise_cohen()`
  entry point lands, this becomes a one-line vignette.
- **UC8** is a written remark wherever an estimator returns both Fleiss
  and BP — the contrast `c'(kappa - kappa_0)` for `c = (0, 1, -1)`
  tests whether the chance baseline matters.
- **UC2** is a defensible per-dataset answer to weight-scheme
  sensitivity. Useful in consultancy / applied write-ups; does not
  carry a paper on its own.

## Reproduce

```bash
cd experiments/probes/09-joint-vcov-pilot
Rscript run_experiment.R                       # full run, ~2 s
Rscript run_experiment.R --smoke               # 20 reps, ~1 s
Rscript run_experiment.R --only uc4 --reps 50  # one UC, faster
quarto render report.qmd
```

Inputs: `misskappa` package loaded directly. Outputs are CSVs under
`results/` plus this report. Seed base controllable via `--seed-base`.
