---
title: "Component-separable vector kappa on CRACKLES"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
read_results <- function(f) {
  p <- file.path("results", f)
  if (!file.exists(p)) {
    stop("Missing ", p, ". Run: Rscript run_experiment.R --smoke", call. = FALSE)
  }
  read.csv(p, check.names = FALSE)
}
complete <- read_results("complete_estimates.csv")
masked <- read_results("masked_summary.csv")
synthetic <- read_results("synthetic_summary.csv")
site_layout <- read_results("site_layout.csv")
meta <- read_results("metadata.csv")
mv <- function(k) meta$value[meta$key == k]
fmt <- function(x, d = 3) formatC(as.numeric(x), digits = d, format = "f")
```

## Question

Can the component-separable vector kappa estimator support patient-level
CRACKLES profiles and component-wise missingness without exposing a new public
R API?

## Short Answer

Yes. The internal `kappa_vector()` wrapper estimates Conger and Fleiss
coefficients for rectangular `subjects x raters x features` arrays, and the
CRACKLES pivot is a natural complete-data pilot. Synthetic component masking
keeps the estimator numerically stable in this small data set. A separate
synthetic DGP with heterogeneous feature/rater missingness checks that the IPW
path is wired to a target where observation rates differ across components.

## Estimand

The implemented class is **component-separable vector loss**. For rater vectors
$x=(x_1,\ldots,x_p)$ and $y=(y_1,\ldots,y_p)$,

$$
D(x,y) =
h\left\{
\frac{\sum_{\ell=1}^p v_\ell d_\ell(x_\ell,y_\ell)}
     {\sum_{\ell=1}^p v_\ell}
\right\}.
$$

With missing components, the pairwise estimator uses the same ratio over
jointly observed components. The IPW estimator replaces each component-pair
term by an inverse observation-rate weighted term: rater-pair-feature rates for
observed disagreement and rater-feature marginal rates for chance disagreement.
The current implementation uses diagonal feature weights only.

This report uses Hamming loss for CRACKLES. For the binary site ratings this is
component mismatch, averaged over the six body sites.

## Data Pivot

`CRACKLES.rda` has `r mv("n_rows")` rows and `r mv("n_rating_columns")`
binary rating columns. Each of the `r mv("n_patients")` patients has six rows:
two upper posterior, two lower posterior, and two anterior recordings.

```{r sites}
site_tab <- site_layout[, c("site", "site_label", "location_code",
                            "location", "side_index")]
names(site_tab) <- c("Site", "Site label", "Location code",
                     "Thorax location", "Within-location side index")
knitr::kable(site_tab, caption = "Six-site patient-level layout.")
```

## Complete-Data Estimates

```{r complete}
tab <- complete[, c("analysis_label", "observer_code", "method", "loss",
                    "R", "features", "kappa_C", "se_C", "kappa_F", "se_F")]
names(tab)[1:4] <- c("Analysis", "Prefix", "Estimator", "Loss")
for (j in c("kappa_C", "se_C", "kappa_F", "se_F")) tab[[j]] <- fmt(tab[[j]])
knitr::kable(tab, align = "llllrrrrrr",
             caption = "Complete CRACKLES vector-kappa estimates.")
```

The pooled row treats all 28 rating columns as one descriptive panel. The
observer-class rows are the cleaner interchangeable-rater analyses because
each prefix identifies a separate clinical or training group.

## Synthetic CRACKLES Masking

```{r masked}
keep <- masked[, c("analysis_label", "observer_code", "method", "loss",
                   "R", "features", "kappa_C_mean", "err_C_mean",
                   "kappa_F_mean", "err_F_mean", "missing_fraction_mean")]
names(keep)[1:4] <- c("Analysis", "Prefix", "Estimator", "Loss")
for (j in c("kappa_C_mean", "err_C_mean", "kappa_F_mean", "err_F_mean",
            "missing_fraction_mean")) {
  keep[[j]] <- fmt(keep[[j]])
}
knitr::kable(keep, align = "llllrrrrrrr",
             caption = "Masked CRACKLES estimates relative to the complete-data pairwise estimate.")
```

The CRACKLES data are naturally complete. The masking here is only an
engineering and estimand check: it deletes components independently at rate
`r fmt(mv("mask_prop"), 2)` for `r mv("reps")` replicates.

## Synthetic Verification

```{r synthetic}
stab <- synthetic[, c("dgp", "method", "loss", "R", "features",
                      "kappa_C_mean", "err_C_mean", "kappa_F_mean",
                      "err_F_mean", "missing_fraction_mean")]
names(stab)[1:3] <- c("DGP", "Estimator", "Loss")
for (j in c("kappa_C_mean", "err_C_mean", "kappa_F_mean", "err_F_mean",
            "missing_fraction_mean")) {
  stab[[j]] <- fmt(stab[[j]])
}
knitr::kable(stab, align = "lllrrrrrrr",
             caption = "Synthetic heterogeneous-MCAR check.")
```

The synthetic DGP uses binary component profiles with feature weights
$(1,2,4)$ and deliberately uneven rater-feature missingness. This is not a
paper simulation; it is a smoke test that the pairwise and IPW paths behave
under component-wise observation rates.

## Caveats

Component-separable losses are not arbitrary vector distances. Whole-profile
exact match, max-component loss, cosine distance, and non-diagonal feature
metrics need different missing-data estimands or model-based integration.

Categorical full-profile FIML is deferred. It is the natural model-based
extension for finite components, but the profile support grows
combinatorially in raters and features.

## Reproduce

```sh
cd experiments/26-crackles-vector-kappa
Rscript run_experiment.R --smoke
quarto render report.qmd
```

For the fuller default:

```sh
Rscript run_experiment.R --reps 25 --mask-prop 0.25 --seed-base 26000
```
