---
title: "Vector kappa on CRACKLES"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
read_results <- function(f) {
  p <- file.path("results", f)
  if (!file.exists(p)) {
    stop("Missing ", p, ". Run: Rscript run_experiment.R --smoke", call. = FALSE)
  }
  read.csv(p, check.names = FALSE)
}
complete <- read_results("complete_estimates.csv")
masked <- read_results("masked_summary.csv")
synthetic <- read_results("synthetic_summary.csv")
quadratic_complete <- read_results("quadratic_complete_estimates.csv")
quadratic_sim <- read_results("quadratic_sim_summary.csv")
quadratic_truth <- read_results("quadratic_sim_truth.csv")
site_layout <- read_results("site_layout.csv")
meta <- read_results("metadata.csv")
mv <- function(k) meta$value[meta$key == k]
fmt <- function(x, d = 3) formatC(as.numeric(x), digits = d, format = "f")
```

## Question

Can the component-separable vector kappa estimator support patient-level
CRACKLES profiles, component-wise missingness, and full-weight quadratic vector
loss without exposing a new public R API?

## Short Answer

Yes. The internal `kappa_vector()` wrapper estimates Conger and Fleiss
coefficients for rectangular `subjects x raters x features` arrays, and the
CRACKLES pivot is a natural complete-data pilot. Synthetic component masking
keeps the estimator numerically stable in this small data set. A separate
synthetic DGP with heterogeneous feature/rater missingness checks that the IPW
path is wired to a target where observation rates differ across components.

The full-weight quadratic route also works internally. It treats
$(x-y)'W(x-y)$ as a moment problem in the stacked rater-feature mean and
covariance, with a pairwise-available covariance estimator and a saturated
normal-theory FIML estimator for the paper-like missing-data experiment.

## Estimand

The implemented class is **component-separable vector loss**. For rater vectors
$x=(x_1,\ldots,x_p)$ and $y=(y_1,\ldots,y_p)$,

$$
D(x,y) =
h\left\{
\frac{\sum_{\ell=1}^p v_\ell d_\ell(x_\ell,y_\ell)}
     {\sum_{\ell=1}^p v_\ell}
\right\}.
$$

With missing components, the pairwise estimator uses the same ratio over
jointly observed components. The IPW estimator replaces each component-pair
term by an inverse observation-rate weighted term: rater-pair-feature rates for
observed disagreement and rater-feature marginal rates for chance disagreement.
The current implementation uses diagonal feature weights only.

## Full-Weight Quadratic Route

For a full symmetric positive-semidefinite feature metric $W$, the squared
vector loss is

$$
D_W(x,y)=(x-y)'W(x-y).
$$

Stack one subject's ratings as $X_i=(X_{i1}',\ldots,X_{iR}')'$ with mean
$\mu$ and covariance $\Sigma$. Let
$P_R=I_R-R^{-1}{\bf 1}_R{\bf 1}_R'$. The internal quadratic backend uses the
unsimplified tensor moment summaries

$$
B_W=\operatorname{tr}\{({\bf 1}_R{\bf 1}_R'\otimes W)\Sigma\},\qquad
T_W=\operatorname{tr}\{(I_R\otimes W)\Sigma\},\qquad
G_W=\mu'(P_R\otimes W)\mu .
$$

The reported coefficients are

$$
\kappa_C^W=\frac{B_W-T_W}{(R-1)T_W+RG_W},\qquad
\kappa_F^W=\frac{B_W-T_W-G_W}{(R-1)(T_W+G_W)} .
$$

The pairwise method estimates $\mu$ and $\Sigma$ with available covariance
moments. The `nt_fiml` method estimates the same $\mu,\Sigma$ under a saturated
normal model and then applies the same map.

This report uses Hamming loss for CRACKLES. For the binary site ratings this is
component mismatch, averaged over the six body sites.

## Data Pivot

`CRACKLES.rda` has `r mv("n_rows")` rows and `r mv("n_rating_columns")`
binary rating columns. Each of the `r mv("n_patients")` patients has six rows:
two upper posterior, two lower posterior, and two anterior recordings.

```{r sites}
site_tab <- site_layout[, c("site", "site_label", "location_code",
                            "location", "side_index")]
names(site_tab) <- c("Site", "Site label", "Location code",
                     "Thorax location", "Within-location side index")
knitr::kable(site_tab, caption = "Six-site patient-level layout.")
```

## Complete-Data Estimates

```{r complete}
tab <- complete[, c("analysis_label", "observer_code", "method", "loss",
                    "R", "features", "kappa_C", "se_C", "kappa_F", "se_F")]
names(tab)[1:4] <- c("Analysis", "Prefix", "Estimator", "Loss")
for (j in c("kappa_C", "se_C", "kappa_F", "se_F")) tab[[j]] <- fmt(tab[[j]])
knitr::kable(tab, align = "llllrrrrrr",
             caption = "Complete CRACKLES vector-kappa estimates.")
```

The pooled row treats all 28 rating columns as one descriptive panel. The
observer-class rows are the cleaner interchangeable-rater analyses because
each prefix identifies a separate clinical or training group.

## Complete-Data Full-W Estimates

```{r quadratic-complete}
qtab <- quadratic_complete[, c("analysis_label", "observer_code", "method",
                               "loss", "R", "features", "W_rank", "kappa_C",
                               "se_C", "kappa_F", "se_F")]
names(qtab)[1:4] <- c("Analysis", "Prefix", "Estimator", "Loss")
for (j in c("kappa_C", "se_C", "kappa_F", "se_F")) qtab[[j]] <- fmt(qtab[[j]])
knitr::kable(qtab, align = "llllrrrrrrr",
             caption = "Complete CRACKLES full-W quadratic estimates.")
```

The six-site CRACKLES metric is identity plus a small same-location
off-diagonal weight for each left/right site pair. These estimates are
pairwise-covariance plug-ins; pooled saturated FIML would require a 168-variate
normal moment fit and is not a useful default run.

## Synthetic CRACKLES Masking

```{r masked}
keep <- masked[, c("analysis_label", "observer_code", "method", "loss",
                   "R", "features", "kappa_C_mean", "err_C_mean",
                   "kappa_F_mean", "err_F_mean", "missing_fraction_mean")]
names(keep)[1:4] <- c("Analysis", "Prefix", "Estimator", "Loss")
for (j in c("kappa_C_mean", "err_C_mean", "kappa_F_mean", "err_F_mean",
            "missing_fraction_mean")) {
  keep[[j]] <- fmt(keep[[j]])
}
knitr::kable(keep, align = "llllrrrrrrr",
             caption = "Masked CRACKLES estimates relative to the complete-data pairwise estimate.")
```

The CRACKLES data are naturally complete. The masking here is only an
engineering and estimand check: it deletes components independently at rate
`r fmt(mv("mask_prop"), 2)` for `r mv("reps")` replicates.

## Synthetic Verification

```{r synthetic}
stab <- synthetic[, c("dgp", "method", "loss", "R", "features",
                      "kappa_C_mean", "err_C_mean", "kappa_F_mean",
                      "err_F_mean", "missing_fraction_mean")]
names(stab)[1:3] <- c("DGP", "Estimator", "Loss")
for (j in c("kappa_C_mean", "err_C_mean", "kappa_F_mean", "err_F_mean",
            "missing_fraction_mean")) {
  stab[[j]] <- fmt(stab[[j]])
}
knitr::kable(stab, align = "lllrrrrrrr",
             caption = "Synthetic heterogeneous-MCAR check.")
```

The synthetic DGP uses binary component profiles with feature weights
$(1,2,4)$ and deliberately uneven rater-feature missingness. This is not a
paper simulation; it is a smoke test that the pairwise and IPW paths behave
under component-wise observation rates.

## Quadratic Missing-Data Simulation

```{r quadratic-truth}
truth_tab <- quadratic_truth
truth_tab$truth <- fmt(truth_tab$truth)
knitr::kable(truth_tab,
             caption = "Analytic tensor-moment truth for the normal quadratic DGP.")
```

```{r quadratic-sim}
qsim <- quadratic_sim[, c("mechanism", "method", "coefficient", "truth",
                          "bias", "rmse", "sd_est", "mean_se", "coverage95",
                          "missing_fraction_mean", "n_used_mean",
                          "valid_reps", "error_reps")]
names(qsim)[1:3] <- c("Missingness", "Estimator", "Coefficient")
for (j in c("truth", "bias", "rmse", "sd_est", "mean_se", "coverage95",
            "missing_fraction_mean", "n_used_mean")) {
  qsim[[j]] <- fmt(qsim[[j]])
}
knitr::kable(qsim, align = "lllrrrrrrrrrr",
             caption = "Normal-theory quadratic simulation with layered missingness.")
```

The simulation uses four raters and three vector components with a full
positive-definite feature metric. `mcar_layered` combines rater-profile dropout
with component-specific dropout. `mar_anchor` keeps rater 1 observed and lets
the other raters' profile and component missingness depend on those observed
anchor values. This is closer to the quadratic paper's normal-moment setup
than the binary CRACKLES masking, but adds the extra rater-feature missingness
layer needed for vector ratings.

## Caveats

Component-separable losses are not arbitrary vector distances. Whole-profile
exact match, max-component loss, and cosine distance need different
missing-data estimands or model-based integration. Full feature metrics are
currently covered only for the quadratic covariance route.

Categorical full-profile FIML is deferred. It is the natural model-based
extension for finite components, but the profile support grows
combinatorially in raters and features.

## Reproduce

```sh
cd experiments/26-crackles-vector-kappa
Rscript run_experiment.R --smoke
quarto render report.qmd
```

For the fuller default:

```sh
Rscript run_experiment.R --reps 25 --quad-reps 25 --mask-prop 0.25 --seed-base 26000
```
