---
title: "Vector Fleiss/Conger kappa on CRACKLES"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
read_results <- function(f) {
  p <- file.path("results", f)
  if (!file.exists(p)) {
    stop("Missing ", p, ". Run: Rscript run_experiment.R", call. = FALSE)
  }
  read.csv(p, check.names = FALSE)
}
complete <- read_results("complete_estimates.csv")
masked <- read_results("masked_summary.csv")
site_layout <- read_results("site_layout.csv")
meta <- read_results("metadata.csv")
mv <- function(k) meta$value[meta$key == k]
fmt <- function(x, d = 3) formatC(as.numeric(x), digits = d, format = "f")
```

## Question

Can the scalar Fréchet/quadratic agreement coefficient be lifted directly
to vector ratings and used as a CRACKLES pilot, where each patient carries
multiple binary site ratings? The experiment uses only the empirical
mean-covariance pair $(\widehat\mu,\widehat\Sigma)$ and then asks what
happens if site ratings are made incomplete by synthetic cell masking.

## Estimand

For $R$ raters and $S$ sites, stack the rater vectors as
$X=(X_1',\ldots,X_R')'$ with $X_r\in\mathbb{R}^S$,
$\mu=(\mu_1',\ldots,\mu_R')'$, and block covariance
$\Sigma^{rs}=\operatorname{Cov}(X_r,X_s)$. For a site metric $W\succeq 0$,
define

$$
T_W=\sum_r \operatorname{tr}(W\Sigma^{rr}),\qquad
B_W=\sum_{r,s}\operatorname{tr}(W\Sigma^{rs}),\qquad
G_W=\sum_r(\mu_r-\bar\mu)'W(\mu_r-\bar\mu),
$$

with $\bar\mu=R^{-1}\sum_r\mu_r$. The two contractions are

$$
\kappa_F^W =
\frac{B_W-T_W-G_W}{(R-1)(T_W+G_W)},\qquad
\kappa_C^W =
\frac{B_W-T_W}{(R-1)T_W+R\,G_W}.
$$

This pilot sets $W=I_S$. On binary CRACKLES ratings this is Hamming loss
over sites. The trace in $B_W-T_W$ keeps same-site cross-rater
covariances and drops cross-site covariances; the normal-FIML missing-data
plug-in can still use the full covariance when imputing missing site
entries.

## Data Pivot

`CRACKLES.rda` has `r mv("n_rows")` rows and `r mv("n_rating_columns")`
binary rating columns. Each of the `r mv("n_patients")` patients has six
rows: two upper posterior, two lower posterior, and two anterior thorax
recordings. The pivot treats those six recordings as the patient-level site
vector and keeps the `UP`/`LO` indicators as location labels:

```{r sites}
knitr::kable(site_layout, caption = "Inferred six-site layout from row order.")
```

The two rows with `UP=LO=0` are anterior thorax recordings; Vanbelle's
tables label this location `A`.

## Short Answer

The vector coefficient is well-defined on the complete CRACKLES pivot and
gives stable, interpretable values. The all-rater pooled analysis has a
smaller $\kappa_F$ than $\kappa_C$ because the rater mean vectors are not
equal; within four-rater panels, the spread is substantial, with `RUS` and
`STU` showing the largest site-level agreement and `WAL`/`NLD` the lowest
in this pilot.

Under synthetic missingness (`r fmt(100 * as.numeric(mv("mask_prop")), 1)`% cell
masking, `r mv("reps")` replicates), complete-case analysis is mostly a
diagnostic because 20 patients vanish quickly in the pooled 28-by-6 matrix.
Pairwise moments remain usable for the pooled matrix. The normal-FIML
moment plug-in is run on the seven four-rater panels, where the dimension
is small enough for this pure-R pilot. FIML is not a categorical model
here; it is the same robust-normal moment idea as the quadratic paper,
used only to produce a full $(\mu,\Sigma)$ for the contraction.

## Complete-Data Estimates

```{r complete}
tab <- complete[, c("analysis", "R", "S", "T", "B", "G", "kappa_F", "kappa_C")]
tab[, c("T", "B", "G", "kappa_F", "kappa_C")] <-
  lapply(tab[, c("T", "B", "G", "kappa_F", "kappa_C")], fmt, d = 3)
knitr::kable(tab, align = "lrrrrrrr",
             caption = "Vector kappa plug-ins on the complete CRACKLES data.")
```

The pooled row treats all 28 columns as raters. The `panel_*` rows treat
each four-rater prefix as a separate interchangeable-rater panel. Those
within-panel rows are the cleaner reading if the column prefixes denote
distinct clinical or reader groups rather than exchangeable raters.

## Synthetic Missing Sites

```{r masked}
keep <- masked[, c("analysis", "method", "R", "S",
                   "kappa_F_mean", "err_F_mean", "err_F_sd",
                   "kappa_C_mean", "err_C_mean", "err_C_sd",
                   "n_used_mean")]
for (j in c("kappa_F_mean", "err_F_mean", "err_F_sd",
            "kappa_C_mean", "err_C_mean", "err_C_sd", "n_used_mean")) {
  keep[[j]] <- fmt(keep[[j]], ifelse(j == "n_used_mean", 1, 3))
}
knitr::kable(keep, align = "llrrrrrrrrr",
             caption = "Mean masked-data estimates and errors relative to the complete-data plug-in.")
```

The masked run is deliberately modest. It is a feasibility check for the
moment contraction, not a missingness theorem. The data are naturally
complete, so MCAR deletion is created by random masking; MAR/site-dependent
masking would be the next experiment if this pilot becomes paper-facing.

## Caveats

The site labels are documented at the location level, but left/right side is
not represented by a separate variable in this copy of the data. The six-row
order is consistent by patient and should be interpreted as two recordings
per location: upper posterior, lower posterior, and anterior.

The all-28-rater analysis is descriptive. If the prefixes are distinct
training cohorts, specialties, or countries, the within-prefix R=4 panels
are the primary interchangeable-rater analyses and the pooled result is a
stress test.

## Reproduce

```sh
cd experiments/26-crackles-vector-kappa
Rscript run_experiment.R --smoke
quarto render report.qmd
```

For the fuller default used here:

```sh
Rscript run_experiment.R --reps 25 --mask-prop 0.25 --seed-base 26000
```
