---
title: "Experiment 06: Raw Estimator Scaling"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
timings_path <- file.path(base_dir, "results", "timings.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(timings_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/probes/06-raw-estimator-scaling/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

table <- summary[order(summary$input, summary$method, summary$n), ]
table$Estimator <- paste(table$input, table$method, sep = ":")
out <- data.frame(
  Estimator = table$Estimator,
  n = table$n,
  Reps = table$reps,
  `Median Time (s)` = fmt(table$median_elapsed_sec),
  `Mean Time (s)` = fmt(table$mean_elapsed_sec),
  Errors = table$n_error,
  check.names = FALSE
)
```

## Question

How steep is the current raw-estimator runtime curve, and how much faster are the counts-format estimators on the same simulated ratings?

## Short Answer

For this run, the log-log scaling exponent for raw available-case runtime is `r fmt(as.numeric(meta[["raw_available_loglog_exponent"]]), 2)`. Values near 2 indicate the all-subject-pairs chance-kernel path is dominating.

## Evidence

```{r}
knitr::kable(out, align = "llrrrr")
```

The raw timings are the performance baseline for the planned `src/estimate_raw.cpp` cleanup. Counts timings are not a replacement for raw rater-identified estimators, but they show the order of magnitude available from aggregated category algebra.

## Caveats

These are wall-clock timings from one local machine and one simulated MCAR design. They are meant to guide engineering priority, not to benchmark hardware. The fitted exponent is only descriptive over the sampled `n` grid.

## Reproduce

Smoke run:

```sh
Rscript experiments/probes/06-raw-estimator-scaling/run_experiment.R --smoke
```

Overnight run:

```sh
Rscript experiments/probes/06-raw-estimator-scaling/run_experiment.R --overnight --progress
```

This report used n grid `r meta[["n_grid"]]`, reps `r meta[["reps"]]`, and seed base `r meta[["seed_base"]]`.
