---
title: "Kappa MSE and Coverage Pilot"
subtitle: "Experiment 27"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
pick_results_dir <- function() {
  override <- Sys.getenv("KAPPA_RESULTS_DIR", unset = "")
  if (nzchar(override)) return(override)

  medium_dirs <- list.dirs("results", full.names = TRUE, recursive = FALSE)
  medium_dirs <- medium_dirs[grepl("medium-", basename(medium_dirs))]
  if (length(medium_dirs)) {
    info <- file.info(file.path(medium_dirs, "summary.csv"))
    medium_dirs <- medium_dirs[!is.na(info$mtime)]
    if (length(medium_dirs)) return(medium_dirs[which.max(info$mtime)])
  }
  "results"
}

results_dir <- pick_results_dir()
summary_path <- file.path(results_dir, "summary.csv")
truth_path <- file.path(results_dir, "truth.csv")
metadata_path <- file.path(results_dir, "metadata.csv")
dgps_path <- file.path(results_dir, "dgps.csv")
mechanisms_path <- file.path(results_dir, "mechanisms.csv")

if (!file.exists(summary_path)) {
  stop(
    "No summary.csv found. Run, for example:\n",
    "  Rscript run_experiment.R --small --progress\n",
    "or set KAPPA_RESULTS_DIR to a result directory before rendering."
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
truth <- read.csv(truth_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
dgps <- read.csv(dgps_path, stringsAsFactors = FALSE)
mechanisms <- read.csv(mechanisms_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_signed <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
}
fmt_int <- function(x) formatC(x, format = "d", big.mark = ",")

method_label <- c(
  ipw_nominal = "IPW",
  cat_fiml_nominal = "Cat-FIML",
  ipw_absolute = "IPW",
  cat_fiml_absolute = "Cat-FIML",
  ipw_quadratic = "IPW",
  cat_fiml_quadratic = "Cat-FIML",
  pairwise_quadratic = "Pairwise",
  nt_fiml_quadratic = "NT-FIML"
)

summary$method_label <- unname(method_label[summary$method])
summary$mechanism_family <- mechanisms$family[match(summary$mechanism, mechanisms$mechanism)]
summary$abs_bias <- abs(summary$bias)

coef_focus <- "Conger"
conger <- subset(summary, coefficient == coef_focus)
```

## Question

How do the current public kappa estimators behave for finite-sample MSE,
standard-error calibration, and Wald coverage under MCAR and clean MAR
missingness? The grid compares IPW and saturated categorical FIML for nominal,
absolute, and quadratic loss, plus pairwise deletion and normal-theory FIML for
the quadratic scored coefficient.

## Run

The report is reading `r results_dir`. Set `KAPPA_RESULTS_DIR` before rendering
to inspect a different run.

```{r run-summary}
rep_path <- file.path(results_dir, "replicates.csv")
replicate_rows <- if (file.exists(rep_path)) {
  nrow(read.csv(rep_path, stringsAsFactors = FALSE))
} else {
  NA_integer_
}

run_tab <- data.frame(
  Quantity = c(
    "Replications per cell",
    "Sample size",
    "Truth sample size",
    "DGPs",
    "Missingness mechanisms",
    "Estimator rows",
    "Summary rows",
    "Replicate rows",
    "Failures",
    "Elapsed minutes"
  ),
  Value = c(
    meta[["reps"]],
    meta[["n_grid"]],
    meta[["truth_n"]],
    meta[["dgps"]],
    meta[["mechanisms"]],
    meta[["methods"]],
    fmt_int(nrow(summary)),
    if (is.na(replicate_rows)) "not checked" else fmt_int(replicate_rows),
    fmt_int(sum(summary$failures)),
    fmt(as.numeric(meta[["elapsed_seconds"]]) / 60, 2)
  ),
  check.names = FALSE
)
knitr::kable(run_tab, align = "ll")
```

```{r dgp-table}
dgp_tab <- dgps[, c("dgp", "C", "R", "support_size", "category_prob", "skill")]
names(dgp_tab) <- c("DGP", "Categories", "Raters", "Full patterns",
                    "Category probabilities", "Rater skills")
knitr::kable(dgp_tab, align = "lrrrll")
```

The MAR mechanisms are all clean observed-data mechanisms. The anchor variants
keep rater 1 observed and delete later raters as a function of that observed
rating. The sequential mechanism deletes each later rater as a function of the
previous observed rating or missingness status. The MCAR mechanisms include
ordinary iid deletion, rater-heterogeneous deletion, and subject-clustered
deletion.

## Short Answer

At this pilot size, the qualitative pattern is coherent. Under MCAR the
quadratic estimators mostly agree. Under MAR in the non-exchangeable and sparse
DGPs, IPW and pairwise quadratic estimators show clear downward bias because
they are MCAR estimators. Cat-FIML and NT-FIML largely remove that MAR bias in
the four-category DGP. In the sparse five-category DGP, Cat-FIML still shows
finite-sample downward bias, while NT-FIML performs well for the scored
quadratic target.

Do not over-read exact coverage percentages from `B = r meta[["reps"]]`; one
coverage step is `r fmt(1 / as.numeric(meta[["reps"]]), 2)`. The table is good
for finding regimes to expand, not for final manuscript claims.

## Quadratic Focus

The clearest stress test is Conger's quadratic coefficient in `biased4` and
`sparse5`. These rows compare the two MCAR routes (`IPW`, `Pairwise`) with the
two likelihood routes (`Cat-FIML`, `NT-FIML`) under one MCAR mechanism and two
MAR mechanisms.

```{r quadratic-focus}
quad <- subset(
  conger,
  weight_label == "quadratic" &
    dgp %in% c("biased4", "sparse5") &
    mechanism %in% c("mcar30", "mar_anchor30", "mar_shifted30")
)
quad$Method <- factor(
  quad$method_label,
  levels = c("Cat-FIML", "IPW", "NT-FIML", "Pairwise")
)
quad <- quad[order(quad$dgp, quad$mechanism, quad$Method), ]
quad_tab <- data.frame(
  DGP = quad$dgp,
  Mechanism = quad$mechanism,
  Method = as.character(quad$Method),
  Bias = fmt_signed(quad$bias, 3),
  RMSE = fmt(quad$rmse, 3),
  `Mean SE` = fmt(quad$mean_se, 3),
  Coverage = fmt(quad$coverage95, 2),
  check.names = FALSE
)
knitr::kable(quad_tab, align = "lllrrrr")
```

The important contrast is not a tiny efficiency difference. It is that the
MCAR estimators are stable under MCAR and biased under clean MAR. In `biased4`
with shifted MAR, IPW and pairwise quadratic are around `-0.08` biased, while
Cat-FIML and NT-FIML are near zero on the same target. In `sparse5`, Cat-FIML
also improves over IPW and pairwise under MAR, but it retains noticeable
downward finite-sample bias.

## All-Loss Summary

The next table averages the Conger summaries across all MCAR mechanisms and
all MAR mechanisms. This is a compression of the full grid, intended to show
which method classes are worth expanding.

```{r family-summary}
family_summary <- do.call(rbind, lapply(split(conger, interaction(
  conger$dgp, conger$mechanism_family, conger$weight_label, conger$method,
  drop = TRUE
)), function(g) {
  data.frame(
    dgp = g$dgp[[1L]],
    family = g$mechanism_family[[1L]],
    weight = g$weight_label[[1L]],
    method = g$method_label[[1L]],
    mean_bias = mean(g$bias),
    mean_abs_bias = mean(abs(g$bias)),
    mean_rmse = mean(g$rmse),
    mean_coverage = mean(g$coverage95),
    stringsAsFactors = FALSE
  )
}))
family_summary <- subset(family_summary, family %in% c("mcar", "mar"))
family_summary <- family_summary[order(
  family_summary$dgp, family_summary$family, family_summary$weight,
  family_summary$method
), ]

family_tab <- data.frame(
  DGP = family_summary$dgp,
  Missingness = toupper(family_summary$family),
  Weight = family_summary$weight,
  Method = family_summary$method,
  `Mean Bias` = fmt_signed(family_summary$mean_bias, 3),
  `Mean |Bias|` = fmt(family_summary$mean_abs_bias, 3),
  `Mean RMSE` = fmt(family_summary$mean_rmse, 3),
  `Mean Coverage` = fmt(family_summary$mean_coverage, 2),
  check.names = FALSE
)
knitr::kable(family_tab, align = "llllrrrr")
```

For `balanced4`, the MAR and MCAR averages are close because the rating model is
exchangeable and mild. For `biased4`, the MAR averages separate: IPW and
pairwise quadratic show larger absolute bias than the likelihood routes. For
`sparse5`, Cat-FIML does not fully inherit the clean four-category FIML
behavior, which is consistent with the known sparse-support finite-sample
issue.

## Worst Coverage Rows

These are the lowest natural-scale Wald coverage rows for Conger. The same
cells are usually the largest-bias cells, so undercoverage is mostly bias, not
just a noisy SE estimate.

```{r worst-coverage}
worst <- conger[order(conger$coverage95, -conger$abs_bias), ]
worst <- head(worst, 18)
worst_tab <- data.frame(
  DGP = worst$dgp,
  Mechanism = worst$mechanism,
  Weight = worst$weight_label,
  Method = worst$method_label,
  Bias = fmt_signed(worst$bias, 3),
  RMSE = fmt(worst$rmse, 3),
  `Mean SE` = fmt(worst$mean_se, 3),
  Coverage = fmt(worst$coverage95, 2),
  check.names = FALSE
)
knitr::kable(worst_tab, align = "llllrrrr")
```

## Runtime

The sparse `5^5 = 3125` categorical FIML cells dominate the run. Shifted MAR is
the slowest sparse mechanism because it combines an observed anchor with
rater-specific observation shifts, producing a harder observed-pattern mix for
the saturated EM.

```{r runtime}
runtime <- subset(summary, coefficient == "Conger")
runtime <- do.call(rbind, lapply(split(runtime, interaction(
  runtime$dgp, runtime$mechanism, runtime$method, drop = TRUE
)), function(g) {
  data.frame(
    dgp = g$dgp[[1L]],
    mechanism = g$mechanism[[1L]],
    method = g$method_label[[1L]],
    weight = g$weight_label[[1L]],
    median_ms = median(g$median_elapsed_ms, na.rm = TRUE),
    stringsAsFactors = FALSE
  )
}))
runtime <- runtime[order(-runtime$median_ms), ]
runtime <- head(runtime, 18)
runtime_tab <- data.frame(
  DGP = runtime$dgp,
  Mechanism = runtime$mechanism,
  Weight = runtime$weight,
  Method = runtime$method,
  `Median Fit Time (ms)` = fmt(runtime$median_ms, 0),
  check.names = FALSE
)
knitr::kable(runtime_tab, align = "llllr")
```

## Caveats

The run used only `B = r meta[["reps"]]` replications at one sample size. It is
large enough to find regimes, but not large enough for final coverage
statements. The next implementation improvement should be checkpointed
per-cell writes, because one slow sparse cell currently holds the whole run in
memory until the end. The next statistical pass should expand only the regimes
that matter: MCAR calibration rows, biased-DGP MAR rows, and the sparse
Cat-FIML finite-sample cells.

## Reproduce

```bash
cd experiments/workbench/27-kappa-coverage-sim

Rscript run_experiment.R --big \
  --reps 50 \
  --n-grid 200 \
  --truth-n 100000 \
  --dgps balanced4,biased4,sparse5 \
  --mechanisms mcar15,mcar30,mcar_rater30,mcar_subject30,mar_anchor15,mar_anchor30,mar_anchor_nonlinear30,mar_shifted30,mar_sequential30 \
  --progress \
  --out-dir results/medium-YYYYMMDD-HHMM

KAPPA_RESULTS_DIR=results/medium-YYYYMMDD-HHMM quarto render report.qmd
```
