---
title: "Kappa MSE and Coverage Pilot"
subtitle: "Experiment 27"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
pick_results_dir <- function() {
  override <- Sys.getenv("KAPPA_RESULTS_DIR", unset = "")
  if (nzchar(override)) return(override)

  medium_dirs <- list.dirs("results", full.names = TRUE, recursive = FALSE)
  medium_dirs <- medium_dirs[grepl("medium-", basename(medium_dirs))]
  if (length(medium_dirs)) {
    info <- file.info(file.path(medium_dirs, "summary.csv"))
    medium_dirs <- medium_dirs[!is.na(info$mtime)]
    if (length(medium_dirs)) return(medium_dirs[which.max(info$mtime)])
  }
  "results"
}

results_dir <- pick_results_dir()
summary_path <- file.path(results_dir, "summary.csv")
truth_path <- file.path(results_dir, "truth.csv")
metadata_path <- file.path(results_dir, "metadata.csv")
dgps_path <- file.path(results_dir, "dgps.csv")
mechanisms_path <- file.path(results_dir, "mechanisms.csv")

if (!file.exists(summary_path)) {
  stop(
    "No summary.csv found. Run, for example:\n",
    "  Rscript run_experiment.R --small --progress\n",
    "or set KAPPA_RESULTS_DIR to a result directory before rendering."
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
truth <- read.csv(truth_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
dgps <- read.csv(dgps_path, stringsAsFactors = FALSE)
mechanisms <- read.csv(mechanisms_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_signed <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
}
fmt_int <- function(x) formatC(x, format = "d", big.mark = ",")

method_label <- c(
  ipw_nominal = "IPW",
  cat_fiml_nominal = "Cat-FIML",
  cat_fiml_jk5_nominal = "Cat-FIML JK5",
  cat_fiml_jk10_nominal = "Cat-FIML JK10",
  ipw_absolute = "IPW",
  cat_fiml_absolute = "Cat-FIML",
  cat_fiml_jk5_absolute = "Cat-FIML JK5",
  cat_fiml_jk10_absolute = "Cat-FIML JK10",
  ipw_quadratic = "IPW",
  cat_fiml_quadratic = "Cat-FIML",
  cat_fiml_jk5_quadratic = "Cat-FIML JK5",
  cat_fiml_jk5_hot_quadratic = "Cat-FIML JK5 hot",
  cat_fiml_jk10_quadratic = "Cat-FIML JK10",
  cat_fiml_jk10_hot_quadratic = "Cat-FIML JK10 hot",
  pairwise_quadratic = "Pairwise",
  nt_fiml_quadratic = "NT-FIML"
)

summary$method_label <- unname(method_label[summary$method])
summary$method_label[is.na(summary$method_label)] <- summary$method[is.na(summary$method_label)]
summary$mechanism_family <- mechanisms$family[match(summary$mechanism, mechanisms$mechanism)]
summary$abs_bias <- abs(summary$bias)

coef_focus <- "Conger"
conger <- subset(summary, coefficient == coef_focus)
```

## Question

How do the current public kappa estimators behave for finite-sample MSE,
standard-error calibration, and Wald coverage under MCAR and clean MAR
missingness? The grid compares IPW and saturated categorical FIML for nominal,
absolute, and quadratic loss, plus pairwise deletion and normal-theory FIML for
the quadratic scored coefficient.

## Run

The report is reading `r results_dir`. Set `KAPPA_RESULTS_DIR` before rendering
to inspect a different run.

```{r run-summary}
rep_path <- file.path(results_dir, "replicates.csv")
replicate_rows <- if (file.exists(rep_path)) {
  nrow(read.csv(rep_path, stringsAsFactors = FALSE))
} else {
  NA_integer_
}

run_tab <- data.frame(
  Quantity = c(
    "Replications per cell",
    "Sample size",
    "Truth sample size",
    "DGPs",
    "Missingness mechanisms",
    "Estimator rows",
    "Summary rows",
    "Replicate rows",
    "Failures",
    "Elapsed minutes"
  ),
  Value = c(
    meta[["reps"]],
    meta[["n_grid"]],
    meta[["truth_n"]],
    meta[["dgps"]],
    meta[["mechanisms"]],
    meta[["methods"]],
    fmt_int(nrow(summary)),
    if (is.na(replicate_rows)) "not checked" else fmt_int(replicate_rows),
    fmt_int(sum(summary$failures)),
    fmt(as.numeric(meta[["elapsed_seconds"]]) / 60, 2)
  ),
  check.names = FALSE
)
knitr::kable(run_tab, align = "ll")
```

```{r dgp-table}
dgp_tab <- dgps[, c("dgp", "C", "R", "support_size", "category_prob", "skill")]
names(dgp_tab) <- c("DGP", "Categories", "Raters", "Full patterns",
                    "Category probabilities", "Rater skills")
knitr::kable(dgp_tab, align = "lrrrll")
```

The MAR mechanisms are all clean observed-data mechanisms. The anchor variants
keep rater 1 observed and delete later raters as a function of that observed
rating. The sequential mechanism deletes each later rater as a function of the
previous observed rating or missingness status. The MCAR mechanisms include
ordinary iid deletion, rater-heterogeneous deletion, and subject-clustered
deletion.

## Short Answer

At this pilot size, the qualitative pattern is coherent. Under MCAR the
quadratic estimators mostly agree. Under MAR in the non-exchangeable and sparse
DGPs, IPW and pairwise quadratic estimators show clear downward bias because
they are MCAR estimators. Cat-FIML and NT-FIML largely remove that MAR bias in
the four-category DGP. In the sparse five-category DGP, Cat-FIML still shows
finite-sample downward bias, while NT-FIML performs well for the scored
quadratic target. Fisher intervals were recorded, but they barely move the
coverage story here because the estimates are not near the boundary; the
problem cells are biased-center cells, not transformed-scale length cells.

Do not over-read exact coverage percentages from `B = r meta[["reps"]]`; one
coverage step is `r fmt(1 / as.numeric(meta[["reps"]]), 2)`. The table is good
for finding regimes to expand, not for final manuscript claims.

## Quadratic Focus

The clearest stress test is Conger's quadratic coefficient in `biased4` and
`sparse5`. These rows compare the two MCAR routes (`IPW`, `Pairwise`) with the
two likelihood routes (`Cat-FIML`, `NT-FIML`) under one MCAR mechanism and two
MAR mechanisms.

```{r quadratic-focus}
quad <- subset(
  conger,
  weight_label == "quadratic" &
    dgp %in% c("biased4", "sparse5") &
    mechanism %in% c("mcar30", "mar_anchor30", "mar_shifted30")
)
quad$Method <- factor(
  quad$method_label,
  levels = c("Cat-FIML", "Cat-FIML JK5", "Cat-FIML JK5 hot",
             "Cat-FIML JK10", "Cat-FIML JK10 hot",
             "IPW", "NT-FIML", "Pairwise")
)
quad <- quad[order(quad$dgp, quad$mechanism, quad$Method), ]
quad_tab <- data.frame(
  DGP = quad$dgp,
  Mechanism = quad$mechanism,
  Method = as.character(quad$Method),
  Bias = fmt_signed(quad$bias, 3),
  RMSE = fmt(quad$rmse, 3),
  `Mean SE` = fmt(quad$mean_se, 3),
  `Wald Cov.` = fmt(quad$coverage95, 2),
  `Fisher Cov.` = fmt(quad$fisher_coverage95, 2),
  `Wald Len.` = fmt(quad$mean_ci_length, 3),
  `Fisher Len.` = fmt(quad$mean_fisher_length, 3),
  check.names = FALSE
)
knitr::kable(quad_tab, align = "lllrrrrrr")
```

The important contrast is not a tiny efficiency difference. It is that the
MCAR estimators are stable under MCAR and biased under clean MAR. In `biased4`
with shifted MAR, IPW and pairwise quadratic are around `-0.08` biased, while
Cat-FIML and NT-FIML are near zero on the same target. In `sparse5`, Cat-FIML
also improves over IPW and pairwise under MAR, but it retains noticeable
downward finite-sample bias.

## All-Loss Summary

The next table averages the Conger summaries across all MCAR mechanisms and
all MAR mechanisms. This is a compression of the full grid, intended to show
which method classes are worth expanding.

```{r family-summary}
family_summary <- do.call(rbind, lapply(split(conger, interaction(
  conger$dgp, conger$mechanism_family, conger$weight_label, conger$method,
  drop = TRUE
)), function(g) {
  data.frame(
    dgp = g$dgp[[1L]],
    family = g$mechanism_family[[1L]],
    weight = g$weight_label[[1L]],
    method = g$method_label[[1L]],
    mean_bias = mean(g$bias),
    mean_abs_bias = mean(abs(g$bias)),
    mean_rmse = mean(g$rmse),
    mean_coverage = mean(g$coverage95),
    stringsAsFactors = FALSE
  )
}))
family_summary <- subset(family_summary, family %in% c("mcar", "mar"))
family_summary <- family_summary[order(
  family_summary$dgp, family_summary$family, family_summary$weight,
  family_summary$method
), ]

family_tab <- data.frame(
  DGP = family_summary$dgp,
  Missingness = toupper(family_summary$family),
  Weight = family_summary$weight,
  Method = family_summary$method,
  `Mean Bias` = fmt_signed(family_summary$mean_bias, 3),
  `Mean |Bias|` = fmt(family_summary$mean_abs_bias, 3),
  `Mean RMSE` = fmt(family_summary$mean_rmse, 3),
  `Mean Coverage` = fmt(family_summary$mean_coverage, 2),
  check.names = FALSE
)
knitr::kable(family_tab, align = "llllrrrr")
```

## Interval Behavior

Fisher intervals were kept as `fisher_coverage95` and
`mean_fisher_length`. They are nearly identical to the natural-scale Wald
intervals in this run. Averaged over Conger rows, Fisher coverage differs from
natural coverage by less than one percentage point for every method-weight
combination, and Fisher lengths are usually within one percent of Wald lengths.

```{r interval-summary}
interval_summary <- aggregate(
  cbind(coverage95, fisher_coverage95, mean_ci_length, mean_fisher_length)     method + method_label + weight_label,
  conger,
  mean
)
interval_summary$cov_diff <- interval_summary$fisher_coverage95 -
  interval_summary$coverage95
interval_summary$len_ratio <- interval_summary$mean_fisher_length /
  interval_summary$mean_ci_length
interval_summary <- interval_summary[order(
  interval_summary$weight_label, interval_summary$method_label
), ]
interval_tab <- data.frame(
  Weight = interval_summary$weight_label,
  Method = interval_summary$method_label,
  `Wald Cov.` = fmt(interval_summary$coverage95, 3),
  `Fisher Cov.` = fmt(interval_summary$fisher_coverage95, 3),
  `Cov. Diff` = fmt_signed(interval_summary$cov_diff, 3),
  `Wald Len.` = fmt(interval_summary$mean_ci_length, 3),
  `Fisher Len.` = fmt(interval_summary$mean_fisher_length, 3),
  `Length Ratio` = fmt(interval_summary$len_ratio, 3),
  check.names = FALSE
)
knitr::kable(interval_tab, align = "llrrrrrr")
```

For this grid, Fisher transformation is not a rescue method. The bad rows are
not high-agreement boundary rows with distorted endpoint behavior. They are
mostly MAR rows where the point estimator is shifted away from the full-data
target.

## Cat-FIML Undercoverage

Cat-FIML looks better here than in the alpha-missing simulations because most
of this grid is much lower-dimensional. The four-category kappa DGPs have
`4^4 = 256` full rating patterns; the sparse stress DGP has `5^5 = 3125`
patterns. The alpha-missing categorical paper grid includes six-item response
spaces, including larger support cells, and the coefficient is a covariance
functional over item batteries rather than an agreement functional over raters.

The remaining Cat-FIML undercoverage is mainly finite-sample bias. In the
sparse DGP, the Cat-FIML standard errors are not obviously too small: the
`se_over_sd` ratios are often near or above one. Coverage is poor because the
intervals are centered too low.

```{r cat-fiml-diagnostics}
cf <- subset(conger, estimator == "cat_fiml")
cf$bias_over_se <- cf$bias / cf$mean_se
cf$abs_bias_over_se <- abs(cf$bias_over_se)
cf_worst <- cf[order(cf$coverage95, -cf$abs_bias_over_se), ]
cf_worst <- head(cf_worst, 14)
cf_tab <- data.frame(
  DGP = cf_worst$dgp,
  Mechanism = cf_worst$mechanism,
  Weight = cf_worst$weight_label,
  Bias = fmt_signed(cf_worst$bias, 3),
  `Bias/SE` = fmt_signed(cf_worst$bias_over_se, 2),
  `SE/SD` = fmt(cf_worst$se_over_sd, 2),
  `Wald Cov.` = fmt(cf_worst$coverage95, 2),
  `Fisher Cov.` = fmt(cf_worst$fisher_coverage95, 2),
  `Wald Len.` = fmt(cf_worst$mean_ci_length, 3),
  `Fisher Len.` = fmt(cf_worst$mean_fisher_length, 3),
  check.names = FALSE
)
knitr::kable(cf_tab, align = "lllrrrrrrr")
```

This points to the same sparse-saturated-likelihood story as the alpha work,
but in a milder form. The fitted saturated distribution has enough unobserved
or thinly observed cells to bias the plug-in moments downward at `n = 200`.
The Wald variance is roughly measuring sampling spread around that biased
center, so the nominal interval misses the full-data target.

For `balanced4`, the MAR and MCAR averages are close because the rating model is
exchangeable and mild. For `biased4`, the MAR averages separate: IPW and
pairwise quadratic show larger absolute bias than the likelihood routes. For
`sparse5`, Cat-FIML does not fully inherit the clean four-category FIML
behavior, which is consistent with the known sparse-support finite-sample
issue.

## Grouped Jackknife

The optional `cat_fiml_jk5_*` and `cat_fiml_jk10_*` rows use delete-group
jackknife corrections as experiment-side bias diagnostics. Plain JK rows use
cold-start delete refits, matching the ordinary Cat-FIML estimator. Rows with
`_hot_` in the method name hot-start the delete refits from the full-data EM
solution and should be read as a speed/sensitivity diagnostic. The reported
interval keeps the full-fit Cat-FIML standard error and recenters it at the
bias-corrected estimate, so this section isolates whether recentring helps
before treating jackknife variance as a separate problem.

```{r grouped-jackknife}
jk <- subset(conger, grepl("^cat_fiml_jk[0-9]+(_hot)?_", method))
if (!nrow(jk)) {
  cat("No grouped jackknife rows are present in this run.")
} else {
  jk$raw_bias_display <- if ("raw_bias" %in% names(jk)) jk$raw_bias else NA_real_
  jk$correction <- jk$bias - jk$raw_bias_display
  jk <- jk[order(jk$dgp, jk$mechanism, jk$weight_label), ]
  jk_tab <- data.frame(
    DGP = jk$dgp,
    Mechanism = jk$mechanism,
    Weight = jk$weight_label,
    Method = jk$method_label,
    `Raw Bias` = fmt_signed(jk$raw_bias_display, 3),
    `JK Bias` = fmt_signed(jk$bias, 3),
    `Correction` = fmt_signed(jk$correction, 3),
    RMSE = fmt(jk$rmse, 3),
    `Wald Cov.` = fmt(jk$coverage95, 2),
    `Mean SE` = fmt(jk$mean_se, 3),
    `Mean Refits` = fmt(jk$mean_jackknife_refits, 1),
    `Mean Fit Time (s)` = fmt(jk$mean_elapsed_ms / 1000, 2),
    check.names = FALSE
  )
  knitr::kable(jk_tab, align = "llllrrrrrrrr")
}
```

## Worst Coverage Rows

These are the lowest natural-scale Wald coverage rows for Conger. The same
cells are usually the largest-bias cells, so undercoverage is mostly bias, not
just a noisy SE estimate.

```{r worst-coverage}
worst <- conger[order(conger$coverage95, -conger$abs_bias), ]
worst <- head(worst, 18)
worst_tab <- data.frame(
  DGP = worst$dgp,
  Mechanism = worst$mechanism,
  Weight = worst$weight_label,
  Method = worst$method_label,
  Bias = fmt_signed(worst$bias, 3),
  RMSE = fmt(worst$rmse, 3),
  `Mean SE` = fmt(worst$mean_se, 3),
  `Wald Cov.` = fmt(worst$coverage95, 2),
  `Fisher Cov.` = fmt(worst$fisher_coverage95, 2),
  `Wald Len.` = fmt(worst$mean_ci_length, 3),
  `Fisher Len.` = fmt(worst$mean_fisher_length, 3),
  check.names = FALSE
)
knitr::kable(worst_tab, align = "llllrrrrrr")
```

## Runtime

The sparse `5^5 = 3125` categorical FIML cells dominate the run. Shifted MAR is
the slowest sparse mechanism because it combines an observed anchor with
rater-specific observation shifts, producing a harder observed-pattern mix for
the saturated EM.

```{r runtime}
runtime <- subset(summary, coefficient == "Conger")
runtime <- do.call(rbind, lapply(split(runtime, interaction(
  runtime$dgp, runtime$mechanism, runtime$method, drop = TRUE
)), function(g) {
  data.frame(
    dgp = g$dgp[[1L]],
    mechanism = g$mechanism[[1L]],
    method = g$method_label[[1L]],
    weight = g$weight_label[[1L]],
    median_ms = median(g$median_elapsed_ms, na.rm = TRUE),
    stringsAsFactors = FALSE
  )
}))
runtime <- runtime[order(-runtime$median_ms), ]
runtime <- head(runtime, 18)
runtime_tab <- data.frame(
  DGP = runtime$dgp,
  Mechanism = runtime$mechanism,
  Weight = runtime$weight,
  Method = runtime$method,
  `Median Fit Time (ms)` = fmt(runtime$median_ms, 0),
  check.names = FALSE
)
knitr::kable(runtime_tab, align = "llllr")
```

## Caveats

The run used only `B = r meta[["reps"]]` replications at one sample size. It is
large enough to find regimes, but not large enough for final coverage
statements. The next implementation improvement should be checkpointed
per-cell writes, because one slow sparse cell currently holds the whole run in
memory until the end. The next statistical pass should expand only the regimes
that matter: MCAR calibration rows, biased-DGP MAR rows, and the sparse
Cat-FIML finite-sample cells.

## Reproduce

```bash
cd experiments/workbench/27-kappa-coverage-sim

Rscript run_experiment.R --big \
  --reps 50 \
  --n-grid 200 \
  --truth-n 100000 \
  --dgps balanced4,biased4,sparse5 \
  --mechanisms mcar15,mcar30,mcar_rater30,mcar_subject30,mar_anchor15,mar_anchor30,mar_anchor_nonlinear30,mar_shifted30,mar_sequential30 \
  --progress \
  --out-dir results/medium-YYYYMMDD-HHMM

KAPPA_RESULTS_DIR=results/medium-YYYYMMDD-HHMM quarto render report.qmd
```
