---
title: "Kappa MSE and Coverage Pilot"
subtitle: "Experiment 27"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
results_dir <- "results"
summary_path <- file.path(results_dir, "summary.csv")
replicates_path <- file.path(results_dir, "replicates.csv")
truth_path <- file.path(results_dir, "truth.csv")
metadata_path <- file.path(results_dir, "metadata.csv")

if (!file.exists(summary_path)) {
  stop(
    "results/summary.csv not found. Run:\n",
    "  Rscript run_experiment.R --small --progress\n",
    "from experiments/workbench/27-kappa-coverage-sim first."
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
truth <- read.csv(truth_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_signed <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
}
```

## Question

How do the current public kappa estimators behave for finite-sample MSE and
Wald coverage under MCAR and clean MAR missingness? The estimator grid is IPW
and categorical FIML for nominal, absolute, and quadratic loss, plus pairwise
and normal-theory FIML for the quadratic scored coefficient.

## Short Answer

This report is a workbench scaffold. Interpret only runs with enough
replications for Monte Carlo error to be small. The current run used
`r meta[["mode"]]` mode, `B = r meta[["reps"]]`, sample sizes
`r meta[["n_grid"]]`, and DGPs `r meta[["dgps"]]`.

## Evidence

```{r headline}
headline <- subset(summary, coefficient == "Conger" &
                     mechanism %in% c("mcar30", "mar_anchor30"))
headline <- headline[order(headline$dgp, headline$mechanism,
                           headline$weight_label, headline$method), ]
headline <- head(headline, 48)
tab <- data.frame(
  DGP = headline$dgp,
  Mechanism = headline$mechanism,
  Method = headline$method,
  Weight = headline$weight_label,
  Truth = fmt(headline$truth, 3),
  Bias = fmt_signed(headline$bias, 3),
  RMSE = fmt(headline$rmse, 3),
  `Mean SE` = fmt(headline$mean_se, 3),
  `SE/SD` = fmt(headline$se_over_sd, 2),
  Coverage = fmt(headline$coverage95, 2),
  `Fisher Cov.` = fmt(headline$fisher_coverage95, 2),
  `Valid` = headline$n_valid,
  check.names = FALSE
)
knitr::kable(tab, align = "llllrrrrrrr")
```

The full rectangular outputs are:

- `results/truth.csv`
- `results/replicates.csv`
- `results/summary.csv`
- `results/mechanisms.csv`
- `results/dgps.csv`
- `results/metadata.csv`

## Caveats

The smoke and small modes are for plumbing, signs, and crude ordering only.
The big mode is a starting grid, not a final manuscript design. The sparse
five-category DGP is intentionally included to expose finite-sample saturated
FIML costs before expanding the grid.

## Reproduce

```bash
cd experiments/workbench/27-kappa-coverage-sim
Rscript run_experiment.R --smoke --progress
Rscript run_experiment.R --small --progress
Rscript run_experiment.R --big --progress
quarto render report.qmd
```
