---
title: "Experiment 19: Coefficient alpha under missing data --- overnight regime map"
format:
  html:
    toc: true
    number-sections: false
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
if (!dir.exists(res_dir)) res_dir <- file.path("experiments", "probes", "19-alpha-overnight-regimes", "results")
needed <- c("summary.csv", "truth.csv", "metadata.csv")
missing <- needed[!file.exists(file.path(res_dir, needed))]
if (length(missing)) {
  stop("Missing results: ", paste(missing, collapse = ", "),
       ". Run: Rscript experiments/probes/19-alpha-overnight-regimes/run_experiment.R --smoke")
}
summary <- read.csv(file.path(res_dir, "summary.csv"), stringsAsFactors = FALSE)
truth   <- read.csv(file.path(res_dir, "truth.csv"),   stringsAsFactors = FALSE)
meta    <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)
has_gg  <- requireNamespace("ggplot2", quietly = TRUE)

show <- function(df, cols, digits = 3) {
  df <- df[, intersect(cols, names(df)), drop = FALSE]
  num <- vapply(df, is.numeric, logical(1))
  df[num] <- lapply(df[num], round, digits)
  knitr::kable(df, row.names = FALSE)
}
core <- c("label", "estimator", "truth_alpha", "bias", "mc_sd", "mean_se",
          "se_sd_ratio", "coverage", "mean_ci_width", "pct_undefined", "pct_npd", "mean_time")
```

Run mode: **`r meta$value[meta$key=="mode"]`**, reps = **`r meta$value[meta$key=="reps"]`**,
cells = **`r meta$value[meta$key=="n_cells"]`**, magmaan = **`r meta$value[meta$key=="magmaan"]`**
(misskappa `r meta$value[meta$key=="misskappa_version"]`).

In a smoke run the numbers are noise --- this view exists to prove the harness
is wired end to end (every estimator returns an estimate, an SE, a coverage, a
time). Read the boundaries only on a full overnight run.

## S1 --- pairwise overlap SE vs the average-n Feldt strawman

```{r s1}
show(summary[summary$substudy == "s1", ], core)
```

The headline: pairwise (overlap SE) coverage should track 0.95 where `feldt`
drifts, and the win should be largest under `mcar_planned`. `se_sd_ratio`
isolates whether the SE *formula* is right (→1) from alpha-hat bias.

## S2 --- MCAR -> MAR -> MNAR consistency split

```{r s2}
show(summary[summary$substudy == "s2", ], core)
```

Pairwise bias grows from MCAR to MAR; FIML stays flat until MNAR, where every
estimator is biased (the honesty marker).

## S3 --- nonnormality (IG/Pearson) and the two footnotes

```{r s3}
show(summary[summary$substudy == "s3", ], core)
```

MCAR + nonnormal: `fiml_normal` SE undercovers, `fiml_sandwich` rescues.
MAR + nonnormal: `fiml_normal` *point* bias appears (sandwich does not fix it);
on the discrete companion `cat_fiml` should recover.

## S4 --- off-congeneric x small-n bias

```{r s4}
show(summary[summary$substudy == "s4", ], core)
```

```{r s4-plot, eval=has_gg, fig.width=7, fig.height=4}
library(ggplot2)
d <- summary[summary$substudy == "s4", ]
ggplot(d, aes(n, bias, colour = estimator)) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  geom_line() + geom_point() +
  facet_wrap(~ model) + scale_x_log10() +
  labs(title = "alpha-hat bias vs n by measurement model", y = "bias") +
  theme_minimal()
```

## S5 --- efficiency and cat-FIML cost

```{r s5}
show(summary[summary$substudy == "s5", ],
     c("label", "estimator", "mc_sd", "mean_se", "coverage", "mean_time"))
```

`mc_sd` ratios (pairwise vs FIML) give the ARE; `mean_time` and the discrete
`cost/*` cells trace the C^p wall.

## Torture gallery

```{r torture}
show(summary[summary$substudy == "torture", ], core)
```

`zero_overlap` is the identification boundary: pairwise `pct_undefined` ~ 1 AND
saturated FIML is singular (all NA) --- alpha is unidentified without a
structural model.
`t3_no_4th_moment` stresses the pairwise SE's finite-fourth-moment assumption
(compare `mean_se` to `mean_se_boot`); `alpha_0.97` sits near the irregular
point.

```{r boot-vs-analytic}
b <- summary[is.finite(summary$mean_se_boot) & summary$estimator == "pairwise",
             c("substudy", "label", "mean_se", "mean_se_boot")]
if (nrow(b)) { b$ratio <- round(b$mean_se / b$mean_se_boot, 3); show(b, names(b)) }
```

Analytic-vs-bootstrap SE ratio for the pairwise estimator: ~1 validates the
van Praag overlap formula independent of coverage.
