---
title: "FIML Bias and Sparsity Scaling"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- tryCatch(knitr::current_input(), error = function(e) NA_character_)
base_dir <- if (is.na(input)) getwd() else dirname(normalizePath(input, mustWork = FALSE))
candidate_dirs <- c(
  file.path(base_dir, "results"),
  "results",
  file.path("experiments", "05-fiml-sparsity-scaling", "results")
)
result_dir <- candidate_dirs[file.exists(file.path(candidate_dirs, "summary.csv"))][1L]
if (is.na(result_dir)) {
  stop("Missing results. Run: Rscript experiments/05-fiml-sparsity-scaling/run_experiment.R --smoke")
}

summary <- read.csv(file.path(result_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(result_dir, "metadata.csv"), stringsAsFactors = FALSE)
truth <- read.csv(file.path(result_dir, "truth.csv"), stringsAsFactors = FALSE)

meta <- function(key) metadata$value[match(key, metadata$key)]
fmt <- function(x, digits = 4) ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
fmt_tol <- function(x) formatC(x, format = "e", digits = 0)

headline <- subset(summary, phase == "scaling" & coefficient == "Conger")
headline <- headline[order(headline$profile, headline$n), ]
last_by_profile <- do.call(rbind, lapply(split(headline, headline$profile), function(x) x[which.max(x$n), ]))
```

## Question

Does the FIML point-estimate bias under an exchangeable MCAR rating model
shrink at the expected rate, or does the `C^R` pattern space create a
slow sparse regime?

The runner fixes a truth-plus-guess DGP with exchangeable raters,
identity loss, and rater-specific MCAR observation rates. P1 is the
paper-sized grid (`C = 5`, `R = 6`, `C^R = 15625`). P2 is a low-dimensional
sanity grid (`C = 3`, `R = 4`, `C^R = 81`).

## Short Answer

This is a starter run, not a full-grade answer. It uses
`r meta("reps")` replicates per scaling cell, so the Monte Carlo error is
too large to decide whether bias times `n` is stable. The runner and
report are in place for the full replication budget.

At the largest starter sample size in each profile, Conger bias was:

```{r}
largest_display <- last_by_profile
largest_display$Profile <- largest_display$profile_label
largest_display$`n` <- largest_display$n
largest_display$`Truth` <- fmt(largest_display$truth)
largest_display$`Mean Estimate` <- fmt(largest_display$mean_estimate)
largest_display$`Bias` <- fmt(largest_display$bias)
largest_display$`MC SE(Bias)` <- fmt(largest_display$mc_se_bias)
largest_display$`Bias x n` <- fmt(largest_display$bias_times_n, 2)
largest_display <- largest_display[, c("Profile", "n", "Truth", "Mean Estimate",
                                       "Bias", "MC SE(Bias)", "Bias x n")]
knitr::kable(largest_display, align = "lrrrrrr")
```

## Evidence

The scaling grid below uses the default `prune_tol = 1e-9`.

```{r}
scaling_display <- subset(summary, phase == "scaling" & coefficient == "Conger")
scaling_display <- scaling_display[order(scaling_display$profile, scaling_display$n), ]
scaling_display$Profile <- scaling_display$profile_label
scaling_display$`n` <- scaling_display$n
scaling_display$`Bias` <- fmt(scaling_display$bias)
scaling_display$`SD` <- fmt(scaling_display$sd)
scaling_display$`MC SE(Bias)` <- fmt(scaling_display$mc_se_bias)
scaling_display$`Bias x n` <- fmt(scaling_display$bias_times_n, 2)
scaling_display$`Failures` <- scaling_display$failures
scaling_display <- scaling_display[, c("Profile", "n", "Bias", "SD",
                                       "MC SE(Bias)", "Bias x n", "Failures")]
knitr::kable(scaling_display, align = "lrrrrrr")
```

The pruning sweep holds $n = `r meta("prune_n")`$ and varies the EM
support threshold. If pruning only affects the Louis covariance, the
point-estimate rows should be nearly identical within Monte Carlo error
while `Mean SE` can move.

```{r}
prune_display <- subset(summary, phase == "prune" & coefficient == "Conger")
prune_display <- prune_display[order(prune_display$profile, prune_display$prune_tol), ]
prune_display$Profile <- prune_display$profile_label
prune_display$`Prune Tol` <- fmt_tol(prune_display$prune_tol)
prune_display$`Bias` <- fmt(prune_display$bias)
prune_display$`Mean SE` <- fmt(prune_display$mean_se)
prune_display$`MC SE(Bias)` <- fmt(prune_display$mc_se_bias)
prune_display$`Failures` <- prune_display$failures
prune_display <- prune_display[, c("Profile", "Prune Tol", "Bias", "Mean SE",
                                   "MC SE(Bias)", "Failures")]
knitr::kable(prune_display, align = "lrrrrr")
```

## Caveats

The committed outputs are a starter run. A full-grade run should raise
`--reps` enough that `mc_se_bias` is below `0.001` for the cells used in
the manuscript decision. The high-dimensional P1 cells are feasible on
this machine, but the replication budget is the real cost.

Only identity-loss FIML is included. The analytic truth values are for
the complete-data distribution under the same rating model.

## Reproduce

For the starter run:

```sh
Rscript experiments/05-fiml-sparsity-scaling/run_experiment.R
```

For a cheap mechanical check:

```sh
Rscript experiments/05-fiml-sparsity-scaling/run_experiment.R --smoke
```

For a higher-grade run:

```sh
Rscript experiments/05-fiml-sparsity-scaling/run_experiment.R --reps 200 --prune-reps 60
```
