---
title: "Validation of normal-FIML coefficient alpha (`alpha_continuous`)"
format: html
execute:
  echo: false
  warning: false
---

```{r setup}
res_dir <- "results"
val_path <- file.path(res_dir, "validation.csv")
if (!file.exists(val_path)) {
  stop("No results. Run `Rscript run_experiment.R` first.")
}
validation <- read.csv(val_path, stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)
```

## Question

Does `misskappa::alpha_continuous()` -- Cronbach's alpha from a saturated
multivariate-normal covariance fitted by FIML/EM, with a delta-method standard
error -- agree with the established implementations of the same estimand?

The fit is checked against **lavaan**'s saturated `h1` estimator and
**magmaan**'s `estimate_saturated_em_moments()`; the alpha point and the
**sandwich** SE against **coefficientalpha** (`varphi = 0`, normal-theory ML);
and the SE scale against a nonparametric **case bootstrap**. The bootstrap is a
qualitative sanity check in the checked-in smoke run; the deterministic oracle
comparisons are the main reproducibility checks. The `nonnormal/mcar` cell is
the discriminator.

## Moments

Maximum absolute difference between the `alpha_continuous` EM covariance and the
two oracles, per cell:

```{r moments}
knitr::kable(
  validation[, c("dist", "mech", "max_cov_err_lavaan", "max_cov_err_magmaan")],
  digits = 9,
  col.names = c("dist", "mech", "vs lavaan", "vs magmaan"))
```

The covariance matches lavaan to roughly `1e-8` (a tolerance gap; lavaan runs at
`tol = 1e-12`) and magmaan to roughly `1e-6`.

## Standard errors

```{r ses}
knitr::kable(
  validation[, c("dist", "mech", "alpha", "se_sandwich", "se_normal",
                 "se_ca", "se_boot")],
  digits = 5,
  col.names = c("dist", "mech", "alpha", "SE sandwich", "SE normal",
                "SE coef.alpha", "SE bootstrap"))
```

Reading the table:

- Under **normal complete/MCAR** data, the sandwich and normal-theory SEs are
  close; under **normal/MAR** they need not be identical in finite samples, but
  both are on the same empirical scale as the smoke-run bootstrap.
- The **sandwich** SE reproduces `coefficientalpha(varphi = 0)` to
  near-numerical precision in this run -- an independent confirmation through a
  different code path (`rsem`).
- In the **`nonnormal/mcar`** cell the normal-theory delta SE undershoots the
  sandwich and bootstrap scale, while the sandwich SE stays close to
  `coefficientalpha`. This is the expected behaviour and the reason the package
  defaults to `se_type = "sandwich"`; the
  normal-theory SE (van Zyl, Neudecker & Nel, 2000) is exposed via
  `se_type = "normal"` for the multivariate-normal regime.

The provenance table records whether this is the cheap smoke run (`n = 300`,
`boot = 100`) or a heavier validation run. Use the smoke bootstrap only to catch
gross scale failures; paper-facing coverage and SE claims should come from the
larger Monte Carlo experiments.

## Provenance

```{r meta}
knitr::kable(metadata, col.names = c("key", "value"))
```
