---
title: "Testing Equality Of Agreement Coefficients"
---

```{r}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Every `misskappa` estimate carries an asymptotic covariance matrix, and (for
the influence-function estimators) a per-subject influence matrix `fit$psi`
satisfying `vcov == crossprod(psi) / n^2`. That is exactly what a Wald test of a
linear hypothesis needs, so `kappa_test()` and `alpha_test()` can compare a
coefficient across fits and return a standard `htest`. Only the way the
covariance of the contrast is assembled changes between designs:

* **one sample**: the coefficient's own variance
* **independent samples**: the per-fit variances add (block diagonal)
* **paired / same-subject**: the cross-fit dependence is read from the
  stacked influence functions `fit$psi`

```{r}
#| message: false
library(misskappa)
```

## One sample: is a coefficient equal to a fixed value?

Pass a single fit and a `value`. This tests whether the (linearly weighted)
Conger coefficient for the Gwet (2014) data equals 0.5:

```{r}
kappa_test(kappa(dat.gwet2014, estimator = "ipw"), coef = "Conger", value = 0.5)
```

## Independent samples: two studies, different subjects

When the fits come from disjoint sets of subjects, set `paired = FALSE`. The
variances add. Here the Conger coefficient is compared between the Gwet (2014)
and Klein (2018) studies:

```{r}
kappa_test(
  gwet  = kappa(dat.gwet2014,  estimator = "ipw"),
  klein = kappa(dat.klein2018, estimator = "ipw"),
  coef = "Conger", paired = FALSE
)
```

The same works for coefficient alpha, and `psych::bfi` makes the point on **real
missing data**: 2800 respondents answer the 25-item Big Five inventory and about
360 of them skip at least one item. We reverse-key the negatively-worded items
and treat each five-item scale as a battery. The normal-theory FIML estimator
handles the missing responses under ignorable missingness.

```{r}
data(bfi, package = "psych")
neg <- c("A1", "C4", "C5", "E1", "E2", "O2", "O5")
bfi[neg] <- 7 - bfi[neg]
scale_items <- function(letter) as.matrix(bfi[, paste0(letter, 1:5)])
```

Is the Conscientiousness scale equally reliable for men and women? Different
respondents, so the samples are independent:

```{r}
g <- split(seq_len(nrow(bfi)), bfi$gender)
alpha_test(
  men   = alpha(scale_items("C")[g[["1"]], ], estimator = "nt_fiml"),
  women = alpha(scale_items("C")[g[["2"]], ], estimator = "nt_fiml"),
  paired = FALSE
)
```

## Paired: same subjects, two coefficients

When both coefficients are computed on the *same* subjects, the two estimates
are correlated and ignoring that dependence gives the wrong standard error.
`paired = TRUE` (the default) takes the dependence from the influence functions.

A clean illustration on **real missing data**: in `dat.mcduff2019`, MTurk judges
rate smiling images on two attributes (*smile presence* and *image
positivity*), but only about 4% of the item-by-judge grid is observed. Because
both coefficients are estimated on the same items, the contrast is dependent.

```{r}
to_wide <- function(attribute) {
  m <- with(dat.mcduff2019,
            tapply(get(attribute), list(item, judge), function(z) z[1]))
  matrix(as.numeric(m), nrow = nrow(m))
}
smile    <- kappa(to_wide("rating_smile"),    estimator = "pairwise")
positive <- kappa(to_wide("rating_positive"), estimator = "pairwise")

kappa_test(smile = smile, positive = positive, coef = "Conger")
```

The headline is that this paired test runs on a genuinely incomplete design: the
pairwise-available estimator supplies both the point estimates and the influence
functions that encode their dependence.

Coefficient alpha has the same structure. Every `bfi` respondent answers all five
scales, so two scales' alphas are estimated on the same people and the comparison
is paired. Are Neuroticism and Extraversion equally reliable?

```{r}
alpha_test(N = alpha(scale_items("N"), estimator = "nt_fiml"),
           E = alpha(scale_items("E"), estimator = "nt_fiml"),
           paired = TRUE)
```

## *G*-way homogeneity: are several coefficients all equal?

With more than two fits the test becomes the joint hypothesis that they are all
equal, with `G - 1` degrees of freedom.

For coefficient alpha, all five `bfi` scales are answered by the same
respondents, so testing that they are equally reliable is a paired *G* = 5
homogeneity test (`df = 4`):

```{r}
scales <- c("A", "C", "E", "N", "O")
fits <- lapply(scales, function(L) alpha(scale_items(L), estimator = "nt_fiml"))
names(fits) <- scales
do.call(alpha_test, c(fits, list(paired = TRUE)))
```

The test decisively rejects: Neuroticism (`r round(coef(fits$N), 2)`) is far more
reliable than Openness (`r round(coef(fits$O), 2)`).

The kappa analogue is "is the two-rater quadratic coefficient constant across all
rater pairs?". The Zapf (2016) data has four raters and hence six pairs, all on
the same 50 subjects:

```{r}
pairs <- combn(4, 2)
pair_fits <- apply(pairs, 2, function(rs)
  kappa(dat.zapf2016[, rs], estimator = "pairwise"), simplify = FALSE)
names(pair_fits) <- apply(pairs, 2, paste0, collapse = "")
do.call(kappa_test, c(pair_fits, list(coef = "Conger", paired = TRUE)))
```

## Under the hood

All four cases are built only on the exported `coef()`, `vcov()`, and `fit$psi`.
If you need an arbitrary linear hypothesis (general contrasts with non-zero
margins across several fits), the internal `joint_vcov()` and `wald_test()`
provide the general engine that `kappa_test()` and `alpha_test()` are
convenience wrappers around.
