---
title: "Multirater And Vector-Valued Agreement"
---

```{r}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Ordinary weighted kappa is a *pairwise* construction: chance-corrected
agreement is built from how often two raters disagree. `misskappa` reaches two
generalizations through the same `kappa()` call --- agreement among *more than
two* raters at once, and ratings that are *vectors* rather than single
categories --- without any new functions. Both still report the Conger and
Fleiss coefficients with an asymptotic covariance matrix.

```{r}
#| message: false
library(misskappa)
```

## g-wise (multirater) agreement

The `g` argument is the *arity* of the disagreement kernel: how many raters are
compared at once. The default `g = 2` is the familiar pairwise kappa. With
`g > 2`, observed disagreement averages a symmetric $g$-rater distance over the
within-subject rater combinations, and the chance baseline uses the matching
$g$-rater $V$-statistic --- the Fréchet / Hubert multirater family. The
Conger coefficient uses distinct-rater-combination chance; the Fleiss
coefficient pools the rater margins.

```{r}
# Pairwise (g = 2) versus triple-wise (g = 3) nominal agreement on the
# Gwet (2014) categorical data (five raters, some ratings missing).
kappa(dat.gwet2014, estimator = "ipw", weight = "nominal")
kappa(dat.gwet2014, estimator = "ipw", weight = "nominal", g = 3)
```

The estimator menu carries over. `"ipw"` reweights for missing entries under
MCAR, `"cat_fiml"` is the saturated-multinomial FIML for categorical ratings
under ignorable missingness, and `"pairwise"` is the complete-data estimator
(it errors if any rating is missing). For `g > 2` the supported `weight`s are
`"nominal"`, `"linear"`, and `"hubert"` (the all-raters-equal kernel, defined
only for `g > 2`).

### Quadratic weighting is g-invariant

Quadratic (squared-error) disagreement decomposes into pairwise terms, so the
combinatorial constant cancels in the kappa ratio: the $g$-wise coefficient
equals the pairwise one for *every* $g$. `kappa()` exploits this --- with
`weight = "quadratic"` the `g` argument is ignored and the cheap closed form is
used instead of the $g^{\text{th}}$-order enumeration.

```{r}
g2 <- coef(kappa(dat.gwet2014, estimator = "pairwise", weight = "quadratic"))
g4 <- coef(kappa(dat.gwet2014, estimator = "pairwise", weight = "quadratic", g = 4))
rbind(g2 = g2, g4 = g4)
```

## Vector-valued ratings

Sometimes a single rating is itself a vector. In the CRACKLES auscultation
study shipped as `dat.vanbelle2019`, each of 28 observers classifies the
presence of crackles (`0`/`1`) at six chest sites per patient, so one observer's
verdict on one patient is a six-component vector. The data are a
subjects-by-raters-by-features **array** --- and an array keeps its labels on
every axis:

```{r}
dim(dat.vanbelle2019)
dimnames(dat.vanbelle2019)$site
dat.vanbelle2019["1", "EXP1", ]   # patient 1, observer EXP1, the six sites
```

Pass the array and `kappa()` takes the vector-valued path automatically. The
disagreement between two rater vectors $x$ and $y$ is a component-separable loss

$$
D(x, y) = \sum_{\ell=1}^{p} w_\ell\, d(x_\ell, y_\ell),
$$

and `weight` selects the per-component loss $d$: `"nominal"` is Hamming (count
of disagreeing sites), `"linear"` is absolute difference, `"quadratic"` is
squared difference. The expert panel:

```{r}
kappa(dat.vanbelle2019[, 1:4, ], estimator = "pairwise")
```

The Conger coefficient by observer group reproduces the kind of breakdown
Vanbelle (2019) reports --- experts and pulmonologists agree most, students
least:

```{r}
groups <- c("EXP", "NOR", "RUS", "WAL", "NLD", "PUL", "STU")
conger <- sapply(groups, function(grp) {
  cols <- paste0(grp, 1:4)
  coef(kappa(dat.vanbelle2019[, cols, ], estimator = "pairwise"))["Conger"]
})
round(sort(conger, decreasing = TRUE), 3)
```

Two knobs are passed through `...`: `feature_weights` (non-negative per-site
weights, e.g. to down-weight the anterior sites) and `loss = "rms"` (the
root-mean-square component loss). For missing components, use
`estimator = "ipw"`, which reweights by the per-rater-feature observation rates;
on the complete CRACKLES data it agrees with the pairwise estimator.

```{r}
kappa(dat.vanbelle2019[, 1:4, ], estimator = "ipw",
      feature_weights = c(1, 1, 1, 1, 0.5, 0.5))
```

See the [getting-started vignette](../misskappa.html) for the pairwise
estimators and tests, and the mathematical guides for the loss-matrix and
missing-data foundations these generalizations build on.
