---
title: "Vector-Valued Agreement"
---

```{r}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

::: {.callout-warning}
## Frontier / experimental

Vector-valued agreement with **component-separable feature weights**, and its
missing-data estimators, is our own ongoing research rather than settled, published
methodology. The interface shown here may change, some routes are internal, and the
supporting manuscript is in preparation. Treat results as exploratory. The general
Fréchet framework for multirater agreement (Moss 2024; the multirater /
vector-valued-data paper) is the loose anchor, but it does **not** cover the
separable-weight construction below — that is new here.
:::

```{r}
#| message: false
library(misskappa)
```

Sometimes a single rating is itself a vector. In the CRACKLES auscultation study
shipped as `dat.vanbelle2019`, each of 28 observers classifies the presence of
crackles (`0`/`1`) at six chest sites per patient, so one observer's verdict on one
patient is a six-component vector. The data are a subjects-by-raters-by-features
**array**, and an array keeps its labels on every axis:

```{r}
dim(dat.vanbelle2019)
dimnames(dat.vanbelle2019)$site
dat.vanbelle2019["1", "EXP1", ]   # patient 1, observer EXP1, the six sites
```

Pass the array and `kappa()` takes the vector-valued path automatically.

## Mathematical description

Write a rater's vector rating on feature $\ell$ as $x_\ell$, with non-negative
feature weights $v_\ell$. The disagreement between two rater vectors $x$ and $y$ is
a **component-separable** loss

$$
D(x, y) \;=\; h\!\left(
  \frac{\sum_{\ell=1}^{p} v_\ell\, d_\ell(x_\ell, y_\ell)}
       {\sum_{\ell=1}^{p} v_\ell}
\right),
$$

a weighted average of per-component losses $d_\ell$ passed through a scalar
transform $h$. The `weight` argument selects the component loss: `"nominal"` is
Hamming (count of disagreeing sites, $d_\ell = \mathbf 1[x_\ell \neq y_\ell]$,
$h(t)=t$), `"linear"` is absolute difference, and `"quadratic"` is squared
difference; passing `loss = "rms"` keeps the squared component loss but sets
$h(t)=\sqrt t$. Per-component weights enter through `feature_weights`, e.g. to
down-weight some sites.

The Conger and Fleiss coefficients are then the usual $\kappa = 1 - D_o/D_e$, with
observed disagreement, distinct-rater-pair chance, and pooled-margin chance all
built from these component losses. The separability is deliberate: because
disagreement is a sum over features, a *missing component* simply drops that
feature from the affected rater pair, and the estimator stays a ratio of
feature-level moments — which is what makes the missing-data estimators below
tractable. The closest published precedent is the single-loss multivariate
agreement coefficient of Janson & Olsson (2001); the weighted separable family and
its missing-data theory are the new part, sketched in the proto-paper note under
the project's `dev/notes/`.

## The expert panel

```{r}
kappa(dat.vanbelle2019[, 1:4, ], estimator = "pairwise")
```

The Conger coefficient by observer group reproduces the kind of breakdown
Vanbelle (2019) reports, with experts and pulmonologists agreeing most and students
least:

```{r}
groups <- c("EXP", "NOR", "RUS", "WAL", "NLD", "PUL", "STU")
conger <- sapply(groups, function(grp) {
  cols <- paste0(grp, 1:4)
  coef(kappa(dat.vanbelle2019[, cols, ], estimator = "pairwise"))["Conger"]
})
round(sort(conger, decreasing = TRUE), 3)
```

For missing components, use `estimator = "ipw"`, which reweights by the
per-rater-feature observation rates. Two knobs ride along through `...`:
`feature_weights` (non-negative per-site weights) and `loss = "rms"` (the
root-mean-square component loss). On the complete CRACKLES data IPW agrees with the
pairwise estimator:

```{r}
kappa(dat.vanbelle2019[, 1:4, ], estimator = "ipw",
      feature_weights = c(1, 1, 1, 1, 0.5, 0.5))
```

## The quadratic case

With the squared component loss the construction generalizes further. Allow a
**full** symmetric positive-semidefinite feature-weight matrix $W$, so that
features may be *coupled* rather than weighted one at a time,

$$
D(x, y) \;=\; (x - y)^{\top} W (x - y),
$$

which reduces to the separable squared loss when $W$ is diagonal. As in the scalar
quadratic kappa, the coefficient is then a smooth function of the stacked rater mean
$\mu$ and covariance $\Sigma$: with
$B_W = \sum_{j,k}\operatorname{tr}(W\Sigma_{jk})$,
$T_W = \sum_{j}\operatorname{tr}(W\Sigma_{jj})$, and
$G_W = \sum_{j}(\mu_j-\bar\mu)^{\top} W (\mu_j-\bar\mu)$,

$$
\kappa_{C}^{W} = \frac{B_W - T_W}{(R-1)\,T_W + R\,G_W}, \qquad
\kappa_{F}^{W} = \frac{B_W - T_W - G_W}{(R-1)\,(T_W + G_W)}.
$$

Because the estimand depends on the data only through $(\mu, \Sigma)$, it admits a
**FIML** estimator: fit the saturated normal mean and covariance by EM and use every
partially observed vector, exactly as the scalar `nt_fiml` quadratic kappa does. This
is the vector analogue of `kappa_quadratic_fiml()`.

The publicly supported vector path uses *diagonal* weights, so the squared-loss
coefficient with per-feature weights is reachable directly:

```{r}
kappa(dat.vanbelle2019[, 1:4, ], weight = "quadratic", estimator = "pairwise")
```

The **full-matrix-$W$** quadratic estimators — including the FIML route — currently
live in an internal backend, `kappa_vector_quadratic()`, and are *not* exposed
through `kappa()`. They are experimental: the normal-theory FIML path inverts
per-pattern covariance blocks and is singular when a feature has no within-pattern
variation (which the complete binary CRACKLES data triggers). The call shape, for
reference only, is

```{r}
#| eval: false
# Experimental, internal, may fail to converge on degenerate data — not part of
# the public API. Shown only to document the full-W / FIML route.
fit <- misskappa:::kappa_vector_quadratic(
  dat.vanbelle2019[, 1:4, ],
  method = "nt_fiml",            # or "pairwise"
  W = diag(6)                    # full symmetric PSD feature-weight matrix
)
coef(fit)
```

See the proto-paper note in the project's `dev/notes/` for the estimands, the
delta-method derivations, and the open problems (categorical full-profile FIML,
nonregular RMS at perfect agreement, FIML on degenerate data).

## References

Janson, H., & Olsson, U. (2001). A measure of agreement for interval or nominal
multivariate observations. *Educational and Psychological Measurement*, 61(2),
277–289.

Moss, J. (2024). Measures of agreement with multiple raters: Fréchet variances and
inference. *Psychometrika*. <https://doi.org/10.1007/s11336-023-09945-2>

Vanbelle, S. (2019). Asymptotic variability of (multilevel) multirater kappa
coefficients. *Statistical Methods in Medical Research*, 28(10–11), 3012–3026.
<https://doi.org/10.1177/0962280218794733>
