---
title: "Equality tests for quadratic Conger's kappa: real-data examples"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
read_results <- function(f) {
  p <- file.path("results", f)
  if (!file.exists(p))
    stop("Missing ", p, ". Run: Rscript run_experiment.R", call. = FALSE)
  read.csv(p, check.names = FALSE)
}
A_pairs <- read_results("caseA_pairs.csv")
A_test  <- read_results("caseA_test.csv")
B_kappa <- read_results("caseB_kappa.csv")
B_test  <- read_results("caseB_test.csv")
D_kappa <- read_results("caseD_kappa.csv")
D_test  <- read_results("caseD_test.csv")
C_stat  <- read_results("caseC_status.csv")
C_ran   <- isTRUE(C_stat$ran)
if (C_ran) {
  C_kappa <- read_results("caseC_kappa.csv")
  C_test  <- read_results("caseC_test.csv")
}
meta    <- read_results("metadata.csv")
mv  <- function(k) meta$value[meta$key == k]
fmt <- function(x, d = 3) formatC(as.numeric(x), digits = d, format = "f")
```

## Question

The quadratic-kappa paper promises two equality tests as extensions: a
**two-independent-sample** Wald test on $\widehat{\kappa}^{(1)} -
\widehat{\kappa}^{(2)}$, and a **within-matrix** joint test of whether
the two-rater quadratic kappa is constant across rater pairs. Do both
work on real ordinal rating data using only the package's existing
influence-function machinery? We add a third, substantive *dependent*
contrast --- equal agreement on two constructs rated by the same judges,
on genuinely incomplete data --- and an external benchmark.

## Short answer

Yes, with no new package code. The within-data homogeneity test runs on
`irr::anxiety` through the NT-FIML estimator's influence functions
and the public `kappa_test()` verb; the two-independent-sample test runs on the classic
multiple-sclerosis data of Westlund & Kurland (via Landis & Koch 1977),
where variances simply add; and a *dependent* two-construct test runs on
the genuinely incomplete `dat.mcduff2019` (smile vs positivity, same
judges and items), where the cross-covariance is what makes the test
correct. None shows significant heterogeneity, but all exercise the full
pipeline and reproduce known unweighted-kappa values as a check. As an
external benchmark, we reproduce Vanbelle's (2017) FEES weighted kappas
**exactly** against the `multiagree` package, with a matching
homogeneity-test conclusion.

## Evidence

### Case A --- constant kappa across rater pairs (within one matrix)

`irr::anxiety`: 3 raters score 20 subjects on a 6-point anxiety scale.
We fit the two-rater quadratically weighted Conger kappa on each rater
pair with the paper's NT-FIML estimator (`kappa(estimator = "nt_fiml")`),
then `kappa_test(p12, p13, p23, coef = "Conger", paired = TRUE)` stacks the
three same-subject influence functions and tests that all three are equal.

```{r caseA-pairs}
a <- A_pairs
a$`rater pair` <- sprintf("(%d, %d)", a$rater_1, a$rater_2)
a$`Conger kappa` <- fmt(a$Conger, 3)
a$`SE` <- fmt(a$se, 3)
knitr::kable(a[, c("rater pair", "Conger kappa", "SE")],
  align = "lrr",
  caption = "Per-pair quadratic Conger kappa, anxiety data.")
```

Joint homogeneity test across the `r A_test$n_pairs` pairs (mean pairwise
correlation of the kappa estimates ≈ `r fmt(A_test$mean_pairwise_corr, 2)`):
**Wald** $\chi^2$ = `r fmt(A_test$wald_chisq, 2)` on
`r A_test$df` df, *p* = `r fmt(A_test$p_value, 3)`. At n = 20 the
pairwise kappas range over `r fmt(A_test$range, 2)` but the standard
errors are wide, so there is no evidence the rater pairs differ in
agreement. The point of the example is that the joint test assembles and
runs correctly on a single real rating matrix.

### Case B --- equal kappa across two independent samples

The Westlund & Kurland multiple-sclerosis data (Landis & Koch 1977): two
neurologists place patients into four ordinal diagnostic classes,
separately in Winnipeg (n = 149) and New Orleans (n = 69). These are two
independent samples, so a difference in quadratic Conger kappa is tested
with $z = (\widehat\kappa_1 - \widehat\kappa_2)/\sqrt{\,\widehat{\mathrm{se}}_1^2 + \widehat{\mathrm{se}}_2^2}$.

```{r caseB-kappa}
b <- B_kappa
tab <- reshape(b[, c("weight", "city", "Conger", "se")],
               idvar = "weight", timevar = "city", direction = "wide")
disp <- data.frame(
  `weight`               = tab$weight,
  `kappa (Winnipeg)`     = fmt(tab$`Conger.Winnipeg`, 3),
  `SE (Winnipeg)`        = fmt(tab$`se.Winnipeg`, 3),
  `kappa (New Orleans)`  = fmt(tab$`Conger.New Orleans`, 3),
  `SE (New Orleans)`     = fmt(tab$`se.New Orleans`, 3),
  check.names = FALSE)
knitr::kable(disp, align = "lrrrr",
  caption = "Conger kappa by city and weight scheme.")
```

```{r caseB-test}
t <- B_test
disp2 <- data.frame(
  `weight`     = t$weight,
  `difference` = fmt(t$diff, 3),
  `SE of diff` = fmt(t$se_diff, 3),
  `z`          = fmt(t$z, 2),
  `p-value`    = fmt(t$p_value, 3),
  check.names = FALSE)
knitr::kable(disp2, align = "lrrrr",
  caption = "Two-sample equal-kappa tests (Winnipeg minus New Orleans).")
```

The quadratically weighted agreement is higher in New Orleans
(`r fmt(B_test$kappa_neworleans[B_test$weight == "quadratic"], 2)`) than
Winnipeg
(`r fmt(B_test$kappa_winnipeg[B_test$weight == "quadratic"], 2)`), but
the gap is about one standard error
(*z* = `r fmt(B_test$z[B_test$weight == "quadratic"], 2)`,
*p* = `r fmt(B_test$p_value[B_test$weight == "quadratic"], 2)`): no
significant difference in how well the two neurologist pairs agree, on
any weight scheme. As a correctness check, the unweighted Conger kappa
for the New Orleans table is
`r fmt(B_kappa$Conger[B_kappa$weight == "identity" & B_kappa$city == "New Orleans"], 3)`,
the value reported for this table in the standard references.

### Case D --- two constructs on the same incomplete data (dependent)

`dat.mcduff2019`: MTurk judges rate smiling images on two attributes ---
whether the person is *smiling* and whether the image is *positive*. The
same judges rate the same items, and the design is genuinely incomplete:
only `r fmt(100*D_kappa$observed_frac[1], 1)`% of the
`r D_kappa$n_items[1]` items $\times$ `r D_kappa$n_judges[1]` judges grid
is observed. Because both kappas are computed on the same items, "is
agreement equal for the two constructs?" is a *dependent* contrast, and
`kappa_test(paired = TRUE)` takes the cross-covariance from the influence
functions --- here on real missing data, which is the point.

```{r caseD}
d <- D_kappa
knitr::kable(
  data.frame(
    construct = d$construct,
    `Conger kappa` = fmt(d$Conger, 3),
    `SE` = fmt(d$se, 3),
    check.names = FALSE),
  align = "lrr",
  caption = sprintf("Quadratic Conger kappa per construct (%d items, %d judges, %.1f%% observed).",
                    d$n_items[1], d$n_judges[1], 100*d$observed_frac[1]))
```

The two constructs are agreed upon almost identically (difference
`r fmt(D_test$diff, 3)`), so the dependent Wald test does not reject
($\chi^2$ = `r fmt(D_test$wald_chisq, 3)`, *p* = `r fmt(D_test$p_dependent, 2)`).
The dependence is real but modest (correlation of the two kappa estimates
`r fmt(D_test$corr, 2)`): the dependent standard error
(`r fmt(D_test$se_dependent, 3)`) is about
`r fmt(D_test$se_independent/D_test$se_dependent, 2)`$\times$ tighter than
the naive independent one (`r fmt(D_test$se_independent, 3)`). The
estimator runs stably at ~4% density because the available-case
covariance pools item-level concordance rather than relying on dense
rater-pair overlap.

### Case C --- FEES external benchmark (Vanbelle 2017, `multiagree`)

Vanbelle's (2017) FEES study compares dependent weighted kappas and ships
with the `multiagree` package, so it is the natural external check. Two
observers rate swallows on several ordinal criteria; `multiagree` forms
three rater-pair kappas and tests their equality with a multilevel
Hotelling $T^2$. It aligns the pairs to their common complete cases
(`r if (C_ran) C_test$n_aligned[1] else NA` subjects here), which is also
what `kappa_test(paired = TRUE)` needs, so we can run the same comparison.

```{r caseC, eval=C_ran}
mw <- C_kappa[C_kappa$weight == "quadratic", ]
disp <- data.frame(
  `rater pair`         = mw$pair,
  `our Conger`         = fmt(mw$our_Conger, 4),
  `multiagree kappa`   = fmt(mw$multiagree_kappa, 4),
  `our SE`             = fmt(mw$our_se, 4),
  `multiagree SE`      = fmt(mw$multiagree_se, 4),
  check.names = FALSE)
knitr::kable(disp, align = "lrrrr",
  caption = "Quadratically weighted kappa, FEES: misskappa vs multiagree (aligned cases).")
```

```{r caseC-test, eval=C_ran}
t <- C_test
disp2 <- data.frame(
  `weight`               = t$weight,
  `our Wald chi-sq (df 2)` = fmt(t$our_wald_chisq, 2),
  `our p`                = fmt(t$our_p, 3),
  `multiagree T-sq`      = fmt(t$multiagree_T2, 2),
  `multiagree p`         = fmt(t$multiagree_p, 3),
  check.names = FALSE)
knitr::kable(disp2, align = "lrrrr",
  caption = "Homogeneity of the three rater-pair kappas: our Wald test vs multiagree's Hotelling T-squared.")
```

`r if (!C_ran) as.character(C_stat$note)`

```{r caseC-prose, eval=C_ran, results="asis"}
cat(sprintf(
"The point estimates match **exactly** across all three weight schemes
(the quadratic kappas are %s, %s, %s in both packages), confirming our
Conger kappa is the standard Cohen weighted kappa. The standard errors
agree closely; the small differences are `multiagree`'s multilevel
correction versus our delta-method influence functions. The two
omnibus tests reach the same conclusion --- no significant heterogeneity
among the rater pairs --- with our asymptotic Wald chi-square running
slightly more liberal than the finite-sample Hotelling $T^2$.",
fmt(C_kappa$our_Conger[C_kappa$weight=='quadratic'][1], 3),
fmt(C_kappa$our_Conger[C_kappa$weight=='quadratic'][2], 3),
fmt(C_kappa$our_Conger[C_kappa$weight=='quadratic'][3], 3)))
```

The one caveat is the multilevel design: on this aligned subset the
clustering is negligible (`multiagree`'s multilevel and non-multilevel
fits coincide), but in general two-swallows-per-patient dependence is
outside our one-row-per-subject `kappa_test()`, so `multiagree` remains
the right tool when clustering bites.

## Caveats

* Both A and B are *illustrations of the machinery*, not power studies.
  Calibrated size/power for these tests live in
  `experiments/probes/09-joint-vcov-pilot/` (joint
  vcov) and the simulation experiments.
* Every estimator now exposes per-subject influence functions, so any of
  them can feed `kappa_test()`. Case A uses `nt_fiml`; on complete
  integer-score data the `pairwise` and `nt_fiml` paths give identical
  point estimates.
* Cases A and B use complete data; Case D carries native item
  missingness (~4% observed) and is where the pairwise-available
  estimator earns its keep. Calibrated MCAR robustness of weighted-kappa
  inference is covered in experiments 14 / 22 / 23, and the
  equality-test-under-MCAR story in the companion alpha example
  (experiment 24).
* `dat.zapf2016` (50 subjects, 4 raters, 5-point ordinal) is an
  in-package alternative for a richer within-data Case A if more power is
  wanted.

## Reproduce

```bash
cd experiments/studies/25-kappa-equal-examples
Rscript run_experiment.R            # runs Cases A, B; C if multiagree present
Rscript run_experiment.R --only B   # one case
quarto render report.qmd
```

Run with `misskappa` `r mv("misskappa_version")`, `irr`
`r mv("irr_version")`, R `r mv("R_version")`.
