---
title: "Experiment 15: Categorical Alpha Smoke Checks"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- tryCatch(knitr::current_input(), error = function(e) NA_character_)
base_dir <- if (is.na(input)) getwd() else dirname(normalizePath(input, mustWork = FALSE))
candidate_dirs <- c(
  file.path(base_dir, "results"),
  "results",
  file.path("experiments", "probes", "15-alpha-categorical-smoke", "results")
)
result_dir <- candidate_dirs[file.exists(file.path(candidate_dirs, "summary.csv"))][1L]
if (is.na(result_dir)) {
  stop("Missing results. Run: Rscript experiments/probes/15-alpha-categorical-smoke/run_experiment.R --smoke",
       call. = FALSE)
}

summary <- read.csv(file.path(result_dir, "summary.csv"), stringsAsFactors = FALSE)
timing <- read.csv(file.path(result_dir, "timing_summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(result_dir, "metadata.csv"), stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_ms <- function(x) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = 1))
}

complete_fiml <- subset(summary, mechanism == "complete" & method == "fiml")
complete_gap <- if (nrow(complete_fiml)) max(abs(complete_fiml$mean_delta_complete)) else NA_real_

main_tab <- summary[, c(
  "profile_label", "mechanism", "method", "patterns", "n", "reps",
  "n_valid", "complete_reference_mean", "mean_estimate",
  "mean_delta_complete", "mc_se_delta", "mean_se",
  "cover_complete_ref95", "median_elapsed_ms", "mean_observed_fraction",
  "failures"
)]
main_tab <- main_tab[order(main_tab$profile_label, main_tab$mechanism, main_tab$method), ]
```

## Question

Do the new categorical coefficient-alpha estimators behave mechanically
as expected on small ordinal item batteries?

The runner compares `misskappa::alpha(estimator = "pairwise")` and
`misskappa::alpha(estimator = "cat_fiml")` against each replicate's
complete-data alpha under three mechanisms: complete data, MCAR item
missingness, and an anchor-item MAR mechanism. It also records small
`C^R` timing cells for the saturated categorical EM. This feeds the
standalone alpha-missing paper as a smoke/feasibility check, not as the
full simulation suite.

## Short Answer

This is a mechanical starter run. The complete-data FIML path agrees
with the complete-data available estimator to within
`r fmt(complete_gap, 6)` on the committed run, so the basic R-to-C++
routing and alpha map are coherent. MCAR and MAR rows should be read as
paired deviations from the simulated complete matrix, not as formal
coverage evidence; the later sweep should use fixed population truths
and larger replication budgets.

The useful next step is a full alpha experiment suite only after the
`paper5x6` timing cell is run with enough repeats to set practical
guardrails around `C^R`.

## Evidence

```{r}
display <- main_tab
display$`Profile` <- display$profile_label
display$`Mechanism` <- display$mechanism
display$`Method` <- display$method
display$`C^R` <- display$patterns
display$`n` <- display$n
display$`Reps` <- display$reps
display$`Valid` <- display$n_valid
display$`Complete Ref` <- fmt(display$complete_reference_mean)
display$`Mean Estimate` <- fmt(display$mean_estimate)
display$`Mean Delta` <- fmt(display$mean_delta_complete)
display$`MC SE Delta` <- fmt(display$mc_se_delta)
display$`Mean SE` <- fmt(display$mean_se)
display$`Ref Cover 95` <- fmt(display$cover_complete_ref95)
display$`Median ms` <- fmt_ms(display$median_elapsed_ms)
display$`Obs Frac` <- fmt(display$mean_observed_fraction)
display$`Failures` <- display$failures
display <- display[, c(
  "Profile", "Mechanism", "Method", "C^R", "n", "Reps", "Valid",
  "Complete Ref", "Mean Estimate", "Mean Delta", "MC SE Delta",
  "Mean SE", "Ref Cover 95", "Median ms", "Obs Frac", "Failures"
)]
knitr::kable(display, row.names = FALSE, align = "lllrrrrrrrrrrrrr")
```

The timing rows fit FIML under the MAR mechanism, which exercises the
EM over compatible full-response patterns rather than the trivial
complete-data likelihood.

```{r}
timing_display <- timing
timing_display$`Profile` <- timing_display$profile_label
timing_display$`C^R` <- timing_display$patterns
timing_display$`n` <- timing_display$n
timing_display$`Reps` <- timing_display$reps
timing_display$`Successes` <- timing_display$successes
timing_display$`Failures` <- timing_display$failures
timing_display$`Median ms` <- fmt_ms(timing_display$median_elapsed_ms)
timing_display$`Max ms` <- fmt_ms(timing_display$max_elapsed_ms)
timing_display$`Obs Frac` <- fmt(timing_display$mean_observed_fraction)
timing_display <- timing_display[, c(
  "Profile", "C^R", "n", "Reps", "Successes", "Failures",
  "Median ms", "Max ms", "Obs Frac"
)]
knitr::kable(timing_display, row.names = FALSE, align = "lrrrrrrrr")
```

## Caveats

- The reference is each replicate's complete-data alpha, so the
  `Ref Cover 95` column is a smoke diagnostic, not a population
  coverage estimate.
- Only categorical multinomial EM is exercised. The normal-covariance
  alpha path described in the paper TODO does not exist in the package
  yet.
- The MAR mechanism is ignorable because item 1 is always observed and
  item-specific missingness depends on item 1. It is one sanity cell, not
  a general MAR study.
- The committed smoke run is intentionally small. Use the default or a
  larger `--reps` run before citing any number in a manuscript.

## Reproduce

```bash
Rscript experiments/probes/15-alpha-categorical-smoke/run_experiment.R --smoke
```

Larger starter run:

```bash
Rscript experiments/probes/15-alpha-categorical-smoke/run_experiment.R --progress
```

```{r}
cat(sprintf(
  "Metadata: smoke=%s; profiles=%s; mechanisms=%s; n=%s; reps=%s; timing_profiles=%s; timing_n=%s; timing_reps=%s; seed_base=%s; misskappa=%s\n",
  meta[["smoke"]], meta[["profiles"]], meta[["mechanisms"]], meta[["n"]],
  meta[["reps"]], meta[["timing_profiles"]], meta[["timing_n"]],
  meta[["timing_reps"]], meta[["seed_base"]], meta[["misskappa_version"]]
))
```
