---
title: "Experiment 02: Rater Model Sensitivity"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/02-rater-model-sensitivity/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

headline <- subset(summary, coefficient == "Conger")
headline$model_family <- factor(
  headline$model_family,
  levels = c("guess", "dawid_skene"),
  labels = c("Truth + Guess", "Dawid-Skene")
)
headline$method <- factor(
  headline$method,
  levels = c("available", "ipw", "fiml", "gwet"),
  labels = c("AC", "IPW", "FIML", "Gwet")
)

rank_signature <- function(dat, field) {
  dat <- dat[order(dat[[field]], as.character(dat$method)), ]
  paste(as.character(dat$method), collapse = " > ")
}

rank_compare <- do.call(rbind, lapply(split(headline, headline$dgp), function(g) {
  guess <- subset(g, model_family == "Truth + Guess")
  ds <- subset(g, model_family == "Dawid-Skene")
  data.frame(
    DGP = unique(g$dgp),
    bias_guess = rank_signature(guess, "rank_abs_bias"),
    bias_ds = rank_signature(ds, "rank_abs_bias"),
    sd_guess = rank_signature(guess, "rank_sd"),
    sd_ds = rank_signature(ds, "rank_sd"),
    stringsAsFactors = FALSE
  )
}))

stable_bias <- all(rank_compare$bias_guess == rank_compare$bias_ds)
stable_sd <- all(rank_compare$sd_guess == rank_compare$sd_ds)
verdict <- if (stable_bias && stable_sd) {
  "stable"
} else if (stable_bias || stable_sd) {
  "partially stable"
} else {
  "unstable"
}

fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

evidence <- headline[order(headline$dgp, headline$model_family, headline$rank_abs_bias), ]
evidence_table <- data.frame(
  DGP = evidence$dgp,
  Model = as.character(evidence$model_family),
  Method = as.character(evidence$method),
  Truth = fmt(evidence$truth),
  Bias = fmt(evidence$bias),
  `Monte Carlo SD` = fmt(evidence$sd),
  `Bias Rank` = evidence$rank_abs_bias,
  `SD Rank` = evidence$rank_sd,
  `Error Rate` = fmt(evidence$error_rate, 2),
  check.names = FALSE
)

rank_table <- data.frame(
  DGP = rank_compare$DGP,
  `Bias Order, Truth + Guess` = rank_compare$bias_guess,
  `Bias Order, Dawid-Skene` = rank_compare$bias_ds,
  `SD Order, Truth + Guess` = rank_compare$sd_guess,
  `SD Order, Dawid-Skene` = rank_compare$sd_ds,
  check.names = FALSE
)
```

## Question

Do the A/B/C bias and efficiency orderings from the paper's current simulation design survive when the rating model is changed from latent truth plus guessing to Dawid-Skene per-rater confusion matrices?

## Short Answer

For this run, the Conger identity-loss ordering is **`r verdict`** across the two rating models.

`r if (identical(meta[["smoke"]], "TRUE")) "This is a smoke run, so the result is only a mechanical check of the design and output path." else "This run is large enough to treat the table as experiment evidence, subject to the caveats below."`

## Evidence

```{r}
knitr::kable(evidence_table, align = "lllrrrrrr")
```

The ranking comparison is:

```{r}
knitr::kable(rank_table, align = "lllll")
```

Full per-replicate estimates are in `results/per_rep.csv`; grouped summaries are in `results/summary.csv`.

## Caveats

The Dawid-Skene cells are representative stress tests, not an exhaustive family of rater models. The complete-data truth is computed analytically from the known rater model, so Monte Carlo error comes from the incomplete-data replicates. The C cell uses missingness driven by the previous observed rating, so it is MAR through observed history rather than truth-dependent missingness.

## Reproduce

Smoke run:

```sh
Rscript experiments/02-rater-model-sensitivity/run_experiment.R --smoke
```

Larger run:

```sh
Rscript experiments/02-rater-model-sensitivity/run_experiment.R --reps 50 --n 1000
```

This report used n = `r meta[["n"]]`, reps = `r meta[["reps"]]`, analytic complete-data truth, and seed base `r meta[["seed_base"]]`.
