---
title: "Coverage of IPW and FIML Wald CIs for Conger's kappa"
author: "misskappa experiments"
date: today
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr)

results_dir <- "results"
cov_path <- file.path(results_dir, "coverage.csv")
rep_path <- file.path(results_dir, "replicates.csv")
meta_path <- file.path(results_dir, "metadata.csv")

if (!file.exists(cov_path)) {
  stop(
    "Results not found. Run:\n",
    "  Rscript run_experiment.R --smoke   # smoke\n",
    "  Rscript run_experiment.R           # default (B = 200)\n",
    "from the experiment directory first."
  )
}

cov <- read.csv(cov_path, stringsAsFactors = FALSE)
rep <- read.csv(rep_path, stringsAsFactors = FALSE)
meta <- read.csv(meta_path, stringsAsFactors = FALSE)

meta_lookup <- function(k) meta$value[meta$key == k]
B <- as.integer(meta_lookup("reps"))
ns <- as.integer(strsplit(meta_lookup("ns"), ",")[[1]])
smoke <- as.logical(meta_lookup("smoke"))
mc_se_cov <- function(p, B) sqrt(p * (1 - p) / B)
```

## Question

Do the IPW influence-function Wald CIs and the rank-truncated FIML
Louis-information Wald CIs for Conger's $\kappa$ achieve calibrated
standard errors at moderate sample sizes, across MCAR (DGPs A and B)
and the inherited DGP C?

Three findings, in priority order for the manuscript.

**1. IPW Wald SEs remain close to calibrated under MCAR.** On DGPs A
and B, with `B = 50` replicates per cell, the empirical 95% coverage
runs in the 0.92 – 0.98 band across `n ∈ {500, 2000, 8000}`. The
standardised residual `z = (κ̂ − κ) / SE` is approximately `N(0, 1)`
(`SD(z)` in 0.91 – 1.07). Mean SE is close to the Monte Carlo SD. The
IPW influence-function variance in the companion manuscript remains
good as written.

**2. The FIML Louis SE explosion is fixed by rank truncation, but
finite-sample MLE bias still matters.** The library now forms the
symmetrised Louis information eigendecomposition and drops eigenvalues
below `info_rcond * lambda_max` (`info_rcond = 5e-5` by default). On
DGP A, the mean FIML SE / Monte Carlo SD ratio is now 1.22, 1.09, and
1.12 at `n = 500, 2000, 8000`; previously it was 3.6, 7.9, and 9.3.
On DGP B, the corresponding ratios are 1.00, 1.03, and 1.10. The
runaway `n^{-1/4}` standard-error behaviour is gone. Coverage can
still fall below nominal in moderate samples because the FIML point
estimate has finite-sample bias, especially on DGP A/B at `n <= 2000`;
that is a point-estimation caveat rather than the old Louis-inverse
pathology.

**3. The MAR claim in Section 6 still needs reframing.** Under DGP C the FIML
estimator is *biased*, not merely conservative: coverage collapses
from 0.94 at `n = 500` to 0.08 at `n = 2000` to 0.00 at `n = 8000`,
with mean `z` growing as `~√n` (0.5, 3.1, 7.7). The reason: DGP C
generates missingness conditional on the latent truth `T_i`, not on
observed entries. Conditional on the full data `X*`, the missingness
mechanism then depends on `X*` through the posterior `P(T | X*)`, and
since `T` is not in the FIML model, this is MNAR with respect to the
`X*`-only model FIML actually fits. Rubin ignorability does not apply
to DGP C. Section 6 of the manuscript currently states "FIML is
approximately unbiased under MAR (DGP C)" — that claim is wrong in
spirit. Either replace DGP C with a cleanly MAR mechanism (missingness
conditional on observed ratings of other raters) or rewrite the prose
to flag that DGP C is technically MNAR and the residual FIML bias is
much smaller than IPW/AC/Gwet bias but not zero.

## Evidence

### Empirical 95% coverage by DGP, method, and `n`

```{r coverage-table}
cov$mc_se_cov95 <- mc_se_cov(cov$cov_95, cov$B)
tab <- cov[, c("dgp", "method", "n", "B", "truth",
               "cov_95", "mc_se_cov95",
               "mean_width_95", "mean_se", "mc_sd_est",
               "mean_z", "sd_z")]
colnames(tab) <- c("DGP", "Method", "n", "B", "True $\\kappa_C$",
                   "Cov$_{.95}$", "MC SE",
                   "Mean Width", "Mean SE", "MC SD",
                   "Mean $z$", "SD $z$")
kable(tab, digits = 4,
      caption = paste0(
        "Empirical 95% Wald coverage of Conger's $\\kappa_C$ ",
        "(identity loss), B = ", B, " replicates per cell. ",
        "`Mean z` and `SD z` summarise the standardised residual ",
        "$(\\hat\\kappa - \\kappa) / \\widehat{\\mathrm{SE}}$, ",
        "which should be approximately N(0, 1)."
      ))
```

The headline columns are `Cov$_{.95}$` (nominal target 0.95) and the
`Mean z` / `SD z` pair (target 0 and 1). The Monte Carlo SE on the
coverage estimate at nominal 0.95 with B replicates is
$\sqrt{0.95 \cdot 0.05 / B}$ ($\approx$ `r round(mc_se_cov(0.95, B), 3)`
at the current B).

### Standardised residual distribution

```{r zhist, fig.width=7, fig.height=4.5}
rep$z <- (rep$est - rep$truth) / rep$se
rep <- rep[is.finite(rep$z), ]
op <- par(mfrow = c(length(unique(rep$dgp)), length(unique(rep$method))),
          mar = c(3, 3, 2, 1), mgp = c(1.7, 0.5, 0))
xlim <- range(c(-4, 4, rep$z), na.rm = TRUE)
for (d in sort(unique(rep$dgp))) {
  for (m in unique(rep$method)) {
    sub <- rep[rep$dgp == d & rep$method == m, ]
    hist(sub$z, breaks = seq(min(xlim) - 0.5, max(xlim) + 0.5, by = 0.5),
         xlim = xlim, freq = FALSE,
         main = sprintf("DGP %s -- %s", d, toupper(m)),
         xlab = "z", col = "grey85", border = "white")
    curve(dnorm(x), add = TRUE, col = "black", lwd = 1.5)
    abline(v = c(-1.96, 1.96), col = "grey40", lty = 2)
  }
}
par(op)
```

The dashed lines mark the nominal 95% Wald rejection boundaries
$\pm 1.96$. Under nominal coverage the histogram should sit inside the
standard normal density (solid curve) and the mass beyond the dashed
lines should be near 5%.

## Caveats

- Single coefficient (Conger's $\kappa$, identity loss). Coverage for
  Fleiss / Brennan-Prediger could be added by extending `fit_one()` in
  `run_experiment.R`. The same `vcov` already carries all three diagonal
  entries.
- The rank-truncation mechanism is diagnosed separately in
  `experiments/10-louis-spectrum/`, which decomposes the Louis variance
  by information eigen-direction.
- DGPs A / B / C are inherited from
  `papers/combined/scripts/simulations_raw_three_estimators.R`; they use a
  constrained latent-truth-plus-guess rater model. Coverage under a
  Dawid-Skene rater model is exercised in
  `experiments/02-rater-model-sensitivity`.
- DGP C's missingness depends on the latent truth `T_i`, which is not
  cleanly MAR (the truth is only partially recovered from observed
  raters when guessing is in the mix). A cleanly MAR DGP belongs in
  experiment 02 or as a separate cell here.
- FIML uses `em_options = list(max_iter = 50000, tol = 1e-7)`. Failed-
  to-converge replicates count as `NA` and are excluded from the
  coverage calculation; `n_ok` in `results/coverage.csv` records the
  retained count.

## Reproduce

```bash
cd experiments/01-coverage-iif-louis
Rscript run_experiment.R --smoke        # ~5s, B = 5, n = 200
Rscript run_experiment.R --reps 50      # ~minutes
Rscript run_experiment.R                # default (B = 200, ns up to 8000); hour-ish
quarto render report.qmd
```

Smoke output sets `smoke = TRUE` in `results/metadata.csv` so this
report renders against whichever last-run grade is on disk. Inspect
`results/metadata.csv` to see which grade produced the displayed
numbers.
