---
title: "Experiment 07: Quadratic Kappa Coverage"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/07-quadratic-edgeworth-coverage/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

display_methods <- c("basic", "arcsine", "fisher", "cf_student_minus")
focus <- summary[
  summary$coefficient %in% c("Conger", "Fleiss") &
    summary$method %in% display_methods,
]
focus <- focus[order(focus$target, focus$coefficient, focus$R, focus$n, focus$method), ]
tab <- data.frame(
  Target = focus$target,
  Coefficient = focus$coefficient,
  R = focus$R,
  n = focus$n,
  Method = focus$method,
  `Valid Reps` = focus$valid_reps,
  Coverage = fmt(focus$coverage),
  `Miss Below` = fmt(focus$miss_below),
  `Miss Above` = fmt(focus$miss_above),
  `Mean Length` = fmt(focus$mean_length),
  Bias = fmt(focus$bias),
  `Mean SE` = fmt(focus$mean_se),
  `Mean IF Skew` = fmt(focus$mean_skew_if),
  check.names = FALSE
)

worst <- focus[order(focus$coverage), ][1, ]
best_by_cell <- do.call(rbind, lapply(split(
  focus,
  interaction(focus$target, focus$coefficient, focus$R, focus$n)
), function(d) {
  d[which.min(abs(d$coverage - 0.95)), ]
}))
best_tab <- data.frame(
  Target = best_by_cell$target,
  Coefficient = best_by_cell$coefficient,
  R = best_by_cell$R,
  n = best_by_cell$n,
  Method = best_by_cell$method,
  `Valid Reps` = best_by_cell$valid_reps,
  Coverage = fmt(best_by_cell$coverage),
  `Mean Length` = fmt(best_by_cell$mean_length),
  check.names = FALSE
)
```

## Question

Can a simple Edgeworth-like correction improve small-sample Wald coverage for complete-data categorical quadratic kappa?

## Short Answer

This is still exploratory, but the second pass clarifies the shape of the problem. The worst coverage in this run is `r fmt(worst$coverage)` for target `r worst$target`, `r worst$coefficient`, `R = r worst$R`, `n = r worst$n`, method `r worst$method`. The Moss (2024) pattern is easy to reproduce: the untransformed interval is fragile at small `n`, especially as the true coefficient approaches the upper boundary. Fisher and arcsine transformations help more reliably than the one-term skewness correction. The promising Cornish-Fisher sign, `cf_student_minus`, mostly shifts misses from the upper tail, but it is not yet a standalone fix.

## Evidence

`Miss Below` means the interval upper endpoint is below the true coefficient; `Miss Above` means the lower endpoint is above it.

```{r}
knitr::kable(tab, align = "rlrrlrrrrrrrr")
```

Closest-to-nominal method by cell:

```{r}
knitr::kable(best_tab, align = "rlrrlrrr")
```

## Methods

The simulation follows the complete categorical Perreault-Leigh setup used in Moss (2024): five uniform categories, complete rectangular ratings, and a knowledge probability of `sqrt(target)` so the true quadratic coefficient equals `target`. The ordinary intervals are basic Wald, arcsine, and Fisher transforms. The extra intervals apply Cornish-Fisher quantile shifts from the estimated influence-function skewness:

- `cf_if`: `z + gamma / 6 * (z^2 - 1)`;
- `cf_student_plus`: `z + gamma / 6 * (2 z^2 + 1)`;
- `cf_student_minus`: the same studentized term with the opposite sign.

These are deliberately labelled as probes, not final theory.

## Caveats

No missingness is included here. The experiment uses the categorical available-case implementation with complete data, so IPW and available-case coincide. The Edgeworth variants only use the first influence-skewness term and do not yet account for curvature of the kappa map or studentized standard-error estimation in a derived way.

## Reproduce

Smoke run:

```sh
Rscript experiments/07-quadratic-edgeworth-coverage/run_experiment.R --smoke
```

Longer exploratory run:

```sh
Rscript experiments/07-quadratic-edgeworth-coverage/run_experiment.R --reps 1000 --target-grid 0.5,0.8,0.9 --n-grid 10,20,40,100 --r-grid 2,5 --progress
```

This report used n grid `r meta[["n_grid"]]`, R grid `r meta[["r_grid"]]`, reps `r meta[["reps"]]`, target grid `r meta[["target_grid"]]`, and seed base `r meta[["seed_base"]]`.
