---
title: "Clean MAR replacement DGPs"
subtitle: "Experiment 12"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
results_dir <- "results"
baseline_path <- file.path(results_dir, "baseline_summary.csv")
candidate_path <- file.path(results_dir, "candidate_summary.csv")
ranking_path <- file.path(results_dir, "candidate_ranking.csv")
mechanisms_path <- file.path(results_dir, "mechanisms.csv")
metadata_path <- file.path(results_dir, "metadata.csv")

if (!file.exists(baseline_path)) {
  stop(
    "Results not found. Run:\n",
    "  Rscript run_experiment.R --smoke   # smoke\n",
    "  Rscript run_experiment.R           # default\n",
    "from the experiment directory first."
  )
}

baseline <- read.csv(baseline_path, stringsAsFactors = FALSE)
candidate <- read.csv(candidate_path, stringsAsFactors = FALSE)
ranking <- read.csv(ranking_path, stringsAsFactors = FALSE)
mechanisms <- read.csv(mechanisms_path, stringsAsFactors = FALSE)
meta <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta_kv <- setNames(meta$value, meta$key)

method_order <- c("available", "ipw", "fiml", "gwet")
method_label <- c(available = "AC", ipw = "IPW", fiml = "FIML", gwet = "Gwet")
fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_signed <- function(x, digits = 4) {
  ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
}
fmt_prop <- function(x) fmt(x, 3)
```

## Question

Can the old latent-truth DGP C be replaced by a genuinely MAR mechanism
that demonstrates the methodological point cleanly, and can the gap
between FIML and the non-MAR estimators be made larger without leaving
MAR?

## Short Answer

Yes. The baseline `anchor_moderate` DGP is a clean, fairly mundane MAR
replacement: rater 1 is always observed, and raters 2--4 are observed
with probabilities depending on rater 1's observed category. The
missingness is value-dependent but MAR because the value controlling
missingness is always observed.

At `n = 8000`, the non-FIML estimators retain an asymptotic positive
bias around `r fmt_signed(max(abs(subset(baseline, n == max(baseline$n) & method != "fiml")$bias)), 3)`,
while FIML is near truth (`r fmt_signed(subset(baseline, n == max(baseline$n) & method == "fiml")$bias, 3)`).
This is not a theatrical effect, but it is exactly the methodological
story: under clean MAR the observed ratings are not representative of
the full-data rating distribution, and the observed-data likelihood
uses the incomplete patterns correctly.

More dramatic examples are possible. The candidate search finds
sequential MAR designs where the largest non-FIML bias is about
`r fmt_signed(ranking$largest_non_fiml_bias[1], 3)` while the FIML bias is
`r fmt_signed(ranking$fiml_bias[1], 3)`. These designs are still MAR, but
they are less mundane: later raters are much less likely to be observed
after hard or already-missing previous ratings.

## Why These Mechanisms Are MAR

The old DGP C used missingness probabilities depending on latent truth
`T_i`. That is not MAR for the FIML model used here because `T_i` is not
observed and is not part of the fitted full-data vector.

The replacement mechanisms condition only on observed information:

- **Anchor MAR:** rater 1 is always observed; for raters `j > 1`,
  `P(M_j = 1 | X*) = pi[X_1]`. Since `X_1` is observed for every subject,
  the missingness probability is a function of the observed data.
- **Sequential MAR:** rater 1 is always observed; for raters `j > 1`,
  observation depends on the observed value of rater `j - 1`, or on a
  constant fallback if rater `j - 1` is missing. This is a monotone-style
  observed-history mechanism.

Both preserve positivity in the committed grid: all category-specific
observation probabilities are strictly between zero and one.

## Baseline Evidence

```{r baseline-table}
baseline_display <- baseline
baseline_display$Method <- method_label[baseline_display$method]
baseline_display <- baseline_display[order(baseline_display$n,
                                           match(baseline_display$method, method_order)), ]
tab <- data.frame(
  n = baseline_display$n,
  Method = baseline_display$Method,
  Truth = fmt(baseline_display$truth, 4),
  Mean = fmt(baseline_display$mean_est, 4),
  Bias = fmt_signed(baseline_display$bias, 4),
  `MC SD` = fmt(baseline_display$sd_est, 4),
  `MC SE(Bias)` = fmt(baseline_display$mc_se_bias, 4),
  `Observed Fraction` = fmt_prop(baseline_display$mean_observed_fraction),
  check.names = FALSE
)
knitr::kable(tab, align = "rlrrrrrr",
             caption = paste0(
               "Baseline anchor-MAR DGP, Conger's kappa with identity loss. ",
               "B = ", meta_kv[["baseline_reps"]], " replicates per n."
             ))
```

The gap stabilises rather than vanishes for AC, IPW, and Gwet. FIML has
larger Monte Carlo SD in this small support, but the mean moves toward
the full-data target as `n` increases.

## Candidate Search

```{r ranking-table}
ranking_display <- ranking
ranking_display$Mechanism <- ranking_display$mechanism
ranking_display$Spec <- ranking_display$spec
ranking_display <- ranking_display[order(-ranking_display$abs_bias_gap), ]
rank_tab <- data.frame(
  Spec = ranking_display$Spec,
  Mechanism = ranking_display$Mechanism,
  C = ranking_display$C,
  R = ranking_display$R,
  `Observed Fraction` = fmt_prop(ranking_display$mean_observed_fraction),
  `Largest Non-FIML` = method_label[ranking_display$largest_non_fiml_method],
  `Largest Non-FIML Bias` = fmt_signed(ranking_display$largest_non_fiml_bias, 4),
  `FIML Bias` = fmt_signed(ranking_display$fiml_bias, 4),
  `Abs Bias Gap` = fmt(ranking_display$abs_bias_gap, 4),
  check.names = FALSE
)
knitr::kable(rank_tab, align = "llrrrlrrr",
             caption = paste0(
               "Candidate clean-MAR designs at n = ", meta_kv[["search_n"]],
               ", B = ", meta_kv[["search_reps"]], ". Sorted by the gap ",
               "between the largest non-FIML absolute bias and FIML absolute bias."
             ))
```

The sequential candidates create the largest visual separation. The
mechanism is not invalid: it conditions on observed rating history. But
it is a stronger design choice than the anchor baseline because one
hard or missing rating can make later ratings sparse.

```{r top-candidate-methods}
top_specs <- head(ranking$spec, 3L)
top <- subset(candidate, spec %in% top_specs)
top$method <- factor(top$method, levels = method_order)
top <- top[order(match(top$spec, top_specs), top$method), ]
top_tab <- data.frame(
  Spec = top$spec,
  Method = method_label[as.character(top$method)],
  Truth = fmt(top$truth, 4),
  Mean = fmt(top$mean_est, 4),
  Bias = fmt_signed(top$bias, 4),
  `MC SD` = fmt(top$sd_est, 4),
  check.names = FALSE
)
knitr::kable(top_tab, align = "llrrrr",
             caption = "Method-level summaries for the three most separated candidates.")
```

## Mechanism Parameters

```{r mechanism-table}
mech_tab <- mechanisms[, c("spec", "mechanism", "C", "R", "p",
                           "rho_base", "rho_truth_mult", "pi_anchor",
                           "pi_fallback")]
names(mech_tab) <- c("Spec", "Mechanism", "C", "R", "p",
                     "rho_base", "rho_truth_mult", "pi_anchor",
                     "pi_fallback")
knitr::kable(mech_tab, align = "llrrlllll")
```

## Caveats

- The candidate search is deliberately small and qualitative. It shows
  existence, not an optimal or exhaustive stress test.
- The most dramatic designs are close to the edge of practical
  positivity: they keep nonzero observation probabilities, but some
  categories are observed by later raters only rarely.
- The baseline is probably the better main-text replacement for old DGP
  C. The severe sequential cases are useful as a supplement or a
  robustness paragraph if the paper needs a more visible demonstration.
- This experiment reports point-estimate bias only. Coverage under the
  clean-MAR replacements belongs in a follow-up of experiment 01 if the
  final manuscript wants Wald interval evidence.

## Reproduce

```bash
cd experiments/12-clean-mar-dgp
Rscript run_experiment.R --smoke
Rscript run_experiment.R
quarto render report.qmd
```
