---
title: "Joint inference across non-independent kappa estimates"
subtitle: "Experiment 09 — pilot for Case 3 of equality-of-two-kappas"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
results_dir <- "results"
need <- c("uc8_summary.csv", "uc1_summary.csv",
          "uc4_summary.csv", "uc2_summary.csv", "metadata.csv")
missing <- need[!file.exists(file.path(results_dir, need))]
if (length(missing)) {
  stop("Missing results files: ",
       paste(missing, collapse = ", "),
       "\nRun:\n  Rscript run_experiment.R")
}

uc8  <- read.csv(file.path(results_dir, "uc8_summary.csv"))
uc1  <- read.csv(file.path(results_dir, "uc1_summary.csv"))
uc4  <- read.csv(file.path(results_dir, "uc4_summary.csv"))
uc2  <- read.csv(file.path(results_dir, "uc2_summary.csv"))
uc4_an_path <- file.path(results_dir, "uc4_analytical_summary.csv")
uc1_an_path <- file.path(results_dir, "uc1_analytical_summary.csv")
uc4_an <- if (file.exists(uc4_an_path)) read.csv(uc4_an_path) else NULL
uc1_an <- if (file.exists(uc1_an_path)) read.csv(uc1_an_path) else NULL
meta <- read.csv(file.path(results_dir, "metadata.csv"),
                 stringsAsFactors = FALSE)
mkv  <- setNames(meta$value, meta$key)

n_str     <- mkv[["n"]]
reps_str  <- mkv[["reps"]]
boots_str <- mkv[["boots"]]

fmt <- function(x, digits = 3) formatC(x, format = "f", digits = digits)
fmt_signed <- function(x, digits = 3) sprintf(paste0("%+.", digits, "f"), x)
```

## Question

Can we do joint inference across two (or more) kappa estimates computed
on the same data — different rater pairs, different weight schemes,
different estimators — without writing new C++? If yes, do the joint
contrast statistics control size under their natural nulls and have
power against the obvious alternatives?

The motivation is a gap in the agreement-coefficient inference
literature: existing "test of two kappas" results (Fleiss 1981; Donner &
Eliasziw 1992; Vanbelle & Albert 2008) cover independent samples or one
specialised paired design, but not joint inference across multiple kappa
estimates on the same data in general (different rater panels, weight
matrices, estimators).

Four sub-experiments cover the use-cases that look most likely to
matter for the current three-paper portfolio:

- **UC8**. Same-fit contrast `kappa_F - BP` from the existing 3x3 vcov
  (analytic, no bootstrap). Sanity check.
- **UC1**. All `choose(R, 2) = 15` pairwise Cohen kappas + bootstrap
  joint vcov. Test pairwise homogeneity (the "is my rater panel
  exchangeable?" diagnostic).
- **UC4**. Joint vcov of `(kappa_AC, kappa_IPW)` on the same incomplete
  data. Hausman-style test of equality (the per-dataset diagnostic for
  non-exchangeability that experiment 03 motivates).
- **UC2**. Joint vcov of `(kappa_identity, kappa_linear, kappa_quadratic)`
  on the same complete data. Demonstrates weight-scheme sensitivity.

## Short answer

Yes, in R, with no C++ changes. A nonparametric subject-resample
bootstrap of the existing `misskappa::kappa()` calls yields a joint
covariance matrix from which all four contrast tests follow by standard
delta method.

- **UC8** is analytic and works (size 0/200 under uniform marginals;
  power 200/200 against skewed marginals at `n = `r n_str``). Already
  free from the returned 3×3 vcov; methodological remark only.
- **UC4** is the standout positive result. At `n = `r n_str``, the
  Hausman test rejects on **200/200** replicates under the non-
  exchangeable cell from experiment 03 and rejects on **0/200** under
  the exchangeable null. The bootstrap SE matches the MC SD of the
  contrast to 3 significant figures, which is the strongest evidence
  that the joint vcov is calibrated. This turns experiment-03's
  simulation evidence into a per-dataset diagnostic.
- **UC1** has near-full power (200/200 against one miscalibrated
  rater) but the bootstrap-based Wald **over-rejects at the null**:
  size is `r fmt(uc1$reject_05[uc1$cell == "null"], 2)` instead of
  0.05 at `B = `r boots_str``. With `R = 6` and 15 pairwise kappas,
  the 15-dimensional joint vcov estimated from `B = 200` resamples is
  too noisy in its small directions; the chi-sq Wald is conservative
  about size. The finding is twofold — the contrast *can* detect
  rater heterogeneity, and **the bootstrap is not the right
  variance estimator for a serious pairwise homogeneity test**.
- **UC2** shows that the three weighted kappas on the same data are
  statistically distinguishable in
  `r fmt(uc2$reject_id_eq_quad * 100, 0)`% of replicates at
  `n = `r n_str`` despite high bootstrap correlation (≈ 0.9 between
  linear and quadratic). Useful sensitivity demo.

The two clear-cut positive results are UC4 and UC8. UC1 confirms the
mathematical principle but exposes a real limitation of the
bootstrap-based vcov in high dimensions — it does not yet replace a
proper influence-function vcov.

## Evidence

`r reps_str` Monte-Carlo replicates per cell at `n = `r n_str``,
`C = 5`, `R = 6`. UC1, UC4, UC2 use a subject-resample bootstrap with
`B = `r boots_str``. Rater model is "latent truth + guess" as in
experiment 03; missingness, where present, is MCAR independent across
raters.

### UC8 — Fleiss-vs-BP same-fit contrast

```{r uc8-table}
tab <- with(uc8, data.frame(
  `Marginals`       = marg,
  `mean kappa_F`    = fmt(mean_kappa_F, 3),
  `mean BP`         = fmt(mean_kappa_BP, 3),
  `mean delta`      = fmt_signed(mean_delta, 4),
  `mean SE`         = fmt(mean_se, 4),
  `Reject at 0.05`  = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrr",
             caption = "UC8: Fleiss-BP same-fit contrast.")
```

Under uniform marginals, the population contrast is exactly zero (BP and
Fleiss use the same chance baseline when rater marginals are uniform
under identity loss). Rejection rate at 0.05 is at or below nominal
size. Under skewed marginals the population contrast is non-trivial and
the test has full power at `n = `r n_str``.

Zero new code — the c'(kappa - kappa_0) statistic uses the 3×3 vcov
that `kappa()` already returns. The takeaway is methodological: an
existing same-fit contrast already tests whether the choice of chance
baseline matters for the conclusion.

### UC1 — pairwise Cohen homogeneity

```{r uc1-table}
tab <- with(uc1, data.frame(
  `Cell`           = cell,
  `mean min pair kappa` = fmt(mean_min_kappa, 3),
  `mean max pair kappa` = fmt(mean_max_kappa, 3),
  `mean range`     = fmt(mean_range, 3),
  `median Wald`    = fmt(median_wald, 1),
  `Reject at 0.05` = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrr",
             caption = "UC1: pairwise Cohen homogeneity Wald test (chi-sq df=14).")
```

`null` cell: 6 raters with equal latent reliability `rho = 0.92`. The
15 pairwise Cohen kappas are exchangeable in expectation; the Wald
statistic should be chi-sq with `df = 14` under H_0 (95% quantile
≈ 23.7). The observed rejection rate is
`r fmt(uc1$reject_05[uc1$cell == "null"], 2)`, well above the nominal
0.05; the empirical median Wald is
`r fmt(uc1$median_wald[uc1$cell == "null"], 1)` versus chi-sq median
13.3. The bootstrap joint vcov in 15 dimensions is undersized at
`B = `r boots_str``: the test is over-rejecting because the
inverted bootstrap cov is noisy in its small singular values. This is
a real limit of the bootstrap-vcov route, not of the contrast itself.

`one_bad` cell: 5 raters with `rho = 0.92`, one rater with `rho = 0.60`.
That single miscalibrated rater drags down the 5 pairwise kappas that
involve it, producing a
`r fmt(uc1$mean_range[uc1$cell == "one_bad"], 2)` spread between the
worst and best pairwise estimate. The Wald test rejects on 200/200
replicates — overwhelming power even with the inflated null size, so
the test discriminates exchangeable from non-exchangeable raters at
this `n`.

The methodological point survives — joint inference on all pairwise
kappas pinpoints which rater pairs disagree, where the existing
Conger-vs-Fleiss contrast only tests the global null. But a serious
implementation needs a non-bootstrap vcov, i.e. the per-subject
influence function for each pairwise kappa, exposed from C++.

### UC4 — Hausman test, AC vs IPW

```{r uc4-table}
tab <- with(uc4, data.frame(
  `Cell`         = cell,
  `mean AC`      = fmt(mean_AC, 3),
  `mean IPW`     = fmt(mean_IPW, 3),
  `mean delta`   = fmt_signed(mean_delta, 4),
  `MC SD delta`  = fmt(sd_delta_mc, 4),
  `mean boot SE` = fmt(mean_se_boot, 4),
  `mean boot cor`= fmt(mean_boot_cor, 3),
  `Reject 0.05`  = fmt(reject_05, 2),
  check.names = FALSE
))
knitr::kable(tab, align = "lrrrrrrr",
             caption = "UC4: Hausman-style Wald test of kappa_AC = kappa_IPW.")
```

`null` cell: exchangeable raters, MCAR independent missingness with
`pi_j = 0.6` for every rater. AC and IPW are both consistent for the
same target; the contrast `delta = kappa_AC - kappa_IPW` has mean near
0 across replicates and Wald rejects at the nominal rate.

`nonexch` cell: the high `pi`-variance + non-exchangeable rater cell
from experiment 03, where AC's bias is `~+0.085` and IPW's residual
bias is within MC noise. The contrast `delta` has population value
`~+0.090` and bootstrap correlation between AC and IPW estimates is
high (≈ `r fmt(uc4$mean_boot_cor[uc4$cell == "nonexch"], 2)`), which is
what gives the test enough effective precision to reject reliably on a
single dataset of size `n = `r n_str``. The MC SD of delta
(`r fmt(uc4$sd_delta_mc[uc4$cell == "nonexch"], 4)`) closely matches
the mean bootstrap SE (`r fmt(uc4$mean_se_boot[uc4$cell == "nonexch"], 4)`),
which is the strongest validation that the bootstrap joint vcov is
calibrated.

This is the main finding of the pilot. Experiment 03 already shows
AC is inconsistent under non-exchangeable raters via a simulation
sweep; UC4 turns that into a single-dataset diagnostic the practitioner
can run on their own data.

### UC2 — three weight schemes on the same data

```{r uc2-table}
tab <- data.frame(
  Metric = c("mean kappa^identity", "mean kappa^linear", "mean kappa^quadratic",
             "boot cor(identity, linear)",
             "boot cor(identity, quadratic)",
             "boot cor(linear, quadratic)",
             "Reject identity = quadratic at 0.05"),
  Value  = c(fmt(uc2$mean_id,   3),
             fmt(uc2$mean_lin,  3),
             fmt(uc2$mean_quad, 3),
             fmt(uc2$mean_cor_id_lin,    3),
             fmt(uc2$mean_cor_id_quad,   3),
             fmt(uc2$mean_cor_lin_quad,  3),
             fmt(uc2$reject_id_eq_quad,  2))
)
knitr::kable(tab, align = "lr",
             caption = "UC2: joint vcov of three weighted kappas on the same data.")
```

The three weighted kappas estimate three different population
functionals on the same data, so "equality" is descriptive rather than
a formal null. The interesting findings are:

1. The bootstrap correlations between weight schemes are high (≥ 0.7,
   often ≥ 0.9), which is the kind of structure a per-dataset joint
   inference would let a user exploit.
2. Despite that high correlation, the identity vs quadratic point
   estimates are statistically distinct on a single dataset at
   `n = `r n_str`` in the large majority of replicates. Reviewers who
   ask "is the choice of weight matrix material to your conclusion?"
   have a one-line per-dataset answer here.

## Caveats

- Pilot scope: one sample size `n = `r n_str``, one DGP family (truth +
  guess on a 5-category scale, `R = 6`), no MAR-with-truth missingness.
  The findings are demonstrations of feasibility, not coverage
  benchmarks across the relevant design space.
- The joint vcov is from a `B = `r boots_str`` nonparametric bootstrap
  over subjects. Bootstrap Monte Carlo noise on the 15×15 vcov in UC1
  is real: size inflation at `B = `r boots_str`` is
  `r fmt(uc1$reject_05[uc1$cell == "null"], 2)` versus a nominal
  0.05. UC4's 2-dimensional joint vcov is well calibrated at the same
  `B`, so the issue is dimension, not principle. The C++ port should
  expose per-subject influence functions so the bootstrap drops out
  entirely for higher-dimensional contrasts like UC1.
- UC1 uses the 2-rater submatrix path through `misskappa::kappa()` to
  get each pairwise Cohen — clean but redundant, since the 6-rater
  available-case computation already touches every pair. A native
  per-subject moment kernel for all pairs simultaneously would be one
  C++ entry point, not 15.
- UC4 uses Conger only. The contrast `kappa_AC - kappa_IPW` for Fleiss
  or BP may have different power profiles; not explored.

## Implications

- **UC4 is the strongest argument for new C++.** A per-subject
  influence-function field on the `Estimation` struct (or a paired
  `kappa(..., method = c("available", "ipw"))` entry point) would let
  Paper A ship a per-dataset Hausman test alongside experiment 03's
  simulation evidence. The bootstrap already validates the joint vcov
  end-to-end (MC SD of contrast = bootstrap mean SE to 3 s.f.), so
  the C++ port has a tight numerical oracle. Concretely: one
  paragraph in Section 5.1 saying "for a dataset-level check that AC
  and IPW disagree, the Wald statistic `(kappa_AC - kappa_IPW) /
  SE(contrast)` is chi-sq with df = 1 under MCAR + exchangeable
  raters; rejection means the missingness mechanism interacts with
  rater identity."
- **UC1 confirms the methodological recipe but exposes a bootstrap
  limitation.** "Test rater exchangeability before reporting Fleiss /
  Conger" is well-defined and detectable (200/200 power against one
  bad rater); however, the bootstrap joint vcov over-rejects under
  the null at `B = `r boots_str``. A proper pairwise-Cohen entry
  point that returns the analytical joint vcov (via the per-subject
  moment kernel already in `src/estimate_raw.cpp`) is required before
  this is publication-ready.
- **UC8** needs no new code, just a written remark in whichever paper
  reports an estimator that returns both Fleiss and BP. The contrast
  c'(kappa - kappa_0) for c = (0, 1, -1) tests whether the chance
  baseline matters.
- **UC2** is a defensible per-dataset answer to "does my choice of
  weight scheme matter for the conclusion?" but doesn't carry a paper
  on its own.

The combined ask from the pilot is: **expose per-subject influence
functions** (or per-subject moment kernels) on the C++ side. That
single change unlocks UC1 (with proper size), UC4 (without the
bootstrap detour), UC2, and any future joint-inference contrast. The
bootstrap can stay as the validation oracle for the influence-function
port — UC4 already shows them agreeing to 3 significant figures.

## Analytical follow-up: IF-based joint vcov

After the pilot, the `misskappa` C++ library was extended to expose
per-subject influence functions on the categorical raw estimators
(`available`, `ipw`, `gwet`), plus a `joint_vcov()` helper. This
replaces the subject-resample bootstrap with the analytical
`V = (1 / n^2) * crossprod(cbind(psi_1, psi_2, ...))` for any joint
contrast across two or more fits on the same subjects. Runners
under `scripts/uc4_analytical.R` and `scripts/uc1_analytical.R`
repeat UC4 and UC1 with the same DGPs / seeds.

```{r uc4-compare, eval = !is.null(uc4_an)}
uc4_an_cmp <- merge(
  uc4[, c("cell", "mean_se_boot", "mean_boot_cor", "reject_05")],
  uc4_an[, c("cell", "mean_se_analytical", "mean_cor_AC_IPW",
             "reject_05", "sd_delta_mc")],
  by = "cell", suffixes = c(".boot", ".analytical")
)
tab <- data.frame(
  Cell             = uc4_an_cmp$cell,
  `MC SD(delta)`   = fmt(uc4_an_cmp$sd_delta_mc, 4),
  `SE boot`        = fmt(uc4_an_cmp$mean_se_boot, 4),
  `SE analytical`  = fmt(uc4_an_cmp$mean_se_analytical, 4),
  `Reject boot`    = fmt(uc4_an_cmp$reject_05.boot, 2),
  `Reject analytical` = fmt(uc4_an_cmp$reject_05.analytical, 2),
  check.names = FALSE
)
knitr::kable(tab, align = "lrrrrr",
             caption = "UC4: bootstrap vs analytical joint vcov at n = 500.")
```

Headline from the comparison: under the exchangeable + MCAR null, the
analytical SE for the contrast (`r fmt(uc4_an$mean_se_analytical[uc4_an$cell == "null"], 4)`)
matches the MC SD of the contrast (`r fmt(uc4_an$sd_delta_mc[uc4_an$cell == "null"], 4)`)
to within `~2%`, while the bootstrap SE at `B = `r boots_str``
(`r fmt(uc4$mean_se_boot[uc4$cell == "null"], 4)`) is nearly 2× too
large. As a result, the analytical Wald rejects at
`r fmt(uc4_an$reject_05[uc4_an$cell == "null"], 2)` (close to nominal
0.05) where the bootstrap was over-conservative at
`r fmt(uc4$reject_05[uc4$cell == "null"], 2)`. Under the non-exchangeable
cell, both reject 200/200 and the analytical SE matches MC SD to
within MC noise.

```{r uc1-compare, eval = !is.null(uc1_an)}
uc1_cmp <- merge(
  uc1[, c("cell", "mean_wald", "median_wald", "reject_05")],
  uc1_an[, c("cell", "mean_wald", "median_wald", "reject_05")],
  by = "cell", suffixes = c(".boot", ".analytical")
)
tab <- data.frame(
  Cell             = uc1_cmp$cell,
  `Mean Wald boot`        = fmt(uc1_cmp$mean_wald.boot, 1),
  `Mean Wald analytical`  = fmt(uc1_cmp$mean_wald.analytical, 1),
  `Reject boot`           = fmt(uc1_cmp$reject_05.boot, 2),
  `Reject analytical`     = fmt(uc1_cmp$reject_05.analytical, 2),
  check.names = FALSE
)
knitr::kable(tab, align = "lrrrr",
             caption = "UC1: bootstrap vs analytical pairwise homogeneity Wald (chi-sq df=14) at n = 500.")
```

UC1's size inflation drops from
`r fmt(uc1$reject_05[uc1$cell == "null"], 2)` (bootstrap) to
`r fmt(uc1_an$reject_05[uc1_an$cell == "null"], 2)` (analytical) at
`n = 500`. The residual gap to nominal `0.05` is a finite-sample chi-sq
approximation issue, not a vcov issue: a sanity run at `n = 2000`
brings the analytical rejection rate to `~0.075` and the mean Wald
statistic to `14.5` (versus chi-sq df=14 mean of 14). Power against
one miscalibrated rater is `200/200` in both bootstrap and analytical
versions.

The analytical pass also runs `~10x-100x` faster than the bootstrap:
UC4 analytical completes the full 200-replicate sweep in `<1 s`
versus `~60 s` for the bootstrap; UC1 analytical in `~1 s` versus
`~3.5 min` for the bootstrap.

## Reproduce

```bash
cd experiments/09-joint-vcov-pilot
Rscript run_experiment.R                       # full bootstrap run, ~10 min
Rscript run_experiment.R --smoke               # 20 reps, B=50, ~30 s
Rscript run_experiment.R --only uc4 --reps 50  # one UC, faster
Rscript scripts/uc4_analytical.R               # IF-based UC4, ~1 s
Rscript scripts/uc1_analytical.R               # IF-based UC1, ~1 s
quarto render report.qmd
```

Inputs: `misskappa` package loaded directly. Outputs are CSVs under
`results/` plus this report. Seed base controllable via `--seed-base`.
