---
title: "AC vs IPW efficiency for Conger's kappa"
subtitle: "Experiment 03"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
results_dir <- "results"
mse_path        <- file.path(results_dir, "mse.csv")
replicates_path <- file.path(results_dir, "replicates.csv")
metadata_path   <- file.path(results_dir, "metadata.csv")

if (!file.exists(mse_path)) {
  stop("results/mse.csv not found. Run:\n  Rscript run_experiment.R")
}

mse  <- read.csv(mse_path, stringsAsFactors = FALSE)
meta <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta_kv <- setNames(meta$value, meta$key)

n_str    <- meta_kv[["n"]]
reps_str <- meta_kv[["reps"]]
seed_str <- meta_kv[["seed_base"]]
smoke    <- meta_kv[["smoke"]] == "TRUE"

pi_order   <- c("zero", "mid", "high")
corr_order <- c("none", "mod", "high")
mse$pi_var_label <- factor(mse$pi_var_label, levels = pi_order)
mse$corr_M_label <- factor(mse$corr_M_label, levels = corr_order)
mse$exch_label   <- factor(mse$exch_label,   levels = c("exch", "nonexch"))

fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
fmt_signed <- function(x, digits = 4) {
  ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
}
```

## Question

Under what conditions on the missingness probabilities `pi_j` and rater
exchangeability does IPW's MSE actually beat AC's? The paper's framing
treats AC as an inefficient estimator; the smoke run of
`papers/combined/scripts/simulations_raw_three_estimators.R` has IPW SD `> AC SD`
in the cell where AC is also consistent, which suggests the framing is
backwards in finite samples.

## Short answer

IPW wins on MSE exactly when AC is inconsistent (raters with
heterogeneous skill). Under exchangeable raters AC is unbiased and has
the lower variance, so AC dominates IPW on MSE everywhere across the
`pi_j`-variance and within-subject correlation grid. Variance of `pi_j`
amplifies AC's bias under non-exchangeability and inflates IPW's variance
penalty under exchangeability. Within-pair correlation of missingness
indicators shifts variance but does not flip the ordering. The
manuscript's "AC is inefficient" prose should be reworded around the
bias under non-exchangeability, not finite-sample variance.

## Evidence

`r reps_str` replicates per cell, `n = `r n_str``, `C = 5`, `R = 6`,
Conger's kappa with identity loss. Rater model: latent truth + guess,
matching the paper script. Missingness: Gaussian-copula latent with
equicorrelated within-subject structure at tetrachoric correlation
`rho_M`. The mechanism is independent of the latent truth in every
cell, so IPW is consistent throughout; AC is consistent only under
exchangeable raters.

### Headline: which estimator wins on MSE

```{r mse-winner}
wide <- reshape(
  mse[, c("exch_label", "pi_var_label", "corr_M_label", "method",
          "bias", "sd_est", "mse")],
  idvar     = c("exch_label", "pi_var_label", "corr_M_label"),
  timevar   = "method",
  direction = "wide"
)
wide$mse_ratio <- wide$mse.available / wide$mse.ipw
wide$winner    <- ifelse(wide$mse.available < wide$mse.ipw, "AC", "IPW")
wide <- wide[order(wide$exch_label, wide$pi_var_label, wide$corr_M_label), ]

tab <- data.frame(
  Exchangeability = ifelse(wide$exch_label == "exch", "Exchangeable", "Non-exchangeable"),
  `pi_j variance` = as.character(wide$pi_var_label),
  `corr_M`        = as.character(wide$corr_M_label),
  `Bias AC`       = fmt_signed(wide$bias.available, 4),
  `Bias IPW`      = fmt_signed(wide$bias.ipw, 4),
  `SD AC`         = fmt(wide$sd_est.available, 4),
  `SD IPW`        = fmt(wide$sd_est.ipw, 4),
  `MSE AC`        = fmt(wide$mse.available, 5),
  `MSE IPW`       = fmt(wide$mse.ipw, 5),
  `Winner`        = wide$winner,
  check.names = FALSE
)
knitr::kable(tab, align = "lllrrrrrrl",
             caption = sprintf("Bias, MC SD, and MSE for AC and IPW at n=%s (B=%s).",
                               n_str, reps_str))
```

### Reading the table

Under **exchangeable raters** (top half) AC's bias is at MC-noise level
in every cell and AC's SD is the same as or slightly below IPW's. AC
wins MSE in every cell. The gap grows with `var(pi_j)`: as `pi_j` spans
a wider range, IPW's weights vary more and its variance penalty grows,
while AC stays unbiased.

Under **non-exchangeable raters** (bottom half) AC is heavily biased
once `pi_j` varies. With `pi_var = mid` and no within-subject
correlation, AC's bias is around `r fmt(wide$bias.available[wide$exch_label == "nonexch" & wide$pi_var_label == "mid" & wide$corr_M_label == "none"], 3)`
and IPW's is at MC-noise level. The bias squared dominates AC's MSE in
every non-exchangeable cell with non-zero `pi_var`. IPW wins MSE in
every such cell, often by an order of magnitude.

The exception is the `pi_var = zero` row under non-exchangeability:
when all raters have the same `pi_j`, AC's pair weighting is uniform
and matches IPW asymptotically. AC is consistent and slightly tighter
in this degenerate column.

### Within-subject correlation in M

`corr_M` shifts both estimators' SDs in the same direction within a
row, often non-monotonically (positive correlation reduces effective
sample size for pair counts but also reduces some variance terms). It
does not flip the AC vs IPW ordering in any cell. The bias of AC under
non-exchangeability is essentially unchanged across `corr_M`, since
bias depends on the marginal `pi_j` schedule, not on cross-rater
dependence in M.

### MSE picture

```{r mse-plot, fig.width=7, fig.height=4}
library(ggplot2)

plot_df <- mse
plot_df$method <- factor(toupper(plot_df$method), levels = c("AC", "IPW"))
plot_df$exch_label <- factor(
  ifelse(plot_df$exch_label == "exch", "Exchangeable", "Non-exchangeable"),
  levels = c("Exchangeable", "Non-exchangeable")
)
plot_df$corr_M_label <- factor(
  paste0("corr_M = ", plot_df$corr_M_label),
  levels = paste0("corr_M = ", corr_order)
)
plot_df$pi_var_label <- factor(plot_df$pi_var_label, levels = pi_order)

ggplot(plot_df,
       aes(x = pi_var_label, y = mse, colour = method,
           linetype = method, shape = method, group = method)) +
  geom_line() +
  geom_point(size = 2) +
  scale_y_log10() +
  scale_colour_manual(values = c(AC = "#000000", IPW = "#0072B2")) +
  scale_linetype_manual(values = c(AC = "solid", IPW = "dashed")) +
  scale_shape_manual(values = c(AC = 16, IPW = 17)) +
  facet_grid(exch_label ~ corr_M_label) +
  labs(x = "Variance of pi_j across raters",
       y = "MSE for Conger's kappa (log scale)",
       colour = NULL, linetype = NULL, shape = NULL) +
  theme_classic(base_size = 11) +
  theme(strip.background = element_blank(),
        legend.position = "top")
```

The top row (exchangeable) shows AC at or below IPW everywhere, with the
gap widening as `var(pi_j)` grows. The bottom row (non-exchangeable)
shows AC's MSE shooting up by an order of magnitude as soon as `pi_j`
varies, while IPW's MSE stays nearly flat across `pi_j`-variance.

### Bias vs MC SE check

```{r bias-vs-se}
sig <- with(mse, data.frame(
  Exchangeability = ifelse(exch_label == "exch", "Exchangeable", "Non-exchangeable"),
  `pi_j variance` = as.character(pi_var_label),
  `corr_M`        = as.character(corr_M_label),
  Method          = toupper(method),
  Bias            = fmt_signed(bias, 4),
  `MC SE (bias)`  = fmt(mc_se_bias, 5),
  `|t|`           = fmt(abs(bias) / mc_se_bias, 1),
  check.names = FALSE
))
sig <- sig[order(sig[[1]], sig[[2]], sig[[3]], sig[[4]]), ]
knitr::kable(sig, row.names = FALSE,
             caption = "Bias relative to Monte Carlo SE on the bias.")
```

AC's biases under non-exchangeability with non-zero `pi_var` are well
above MC noise (`|t|` in the tens). IPW's residual biases are within MC
noise across all cells. This confirms that the MSE pattern is driven by
bias, not by Monte Carlo accident.

## Caveats

- One sample size (`n = `r n_str``). The bias-variance trade-off shifts
  with `n`. For `n` small enough, AC's lower variance can dominate
  IPW's bias correction even under non-exchangeability. The MSE ratios
  here are at one moderate point on that curve.
- Conger only, identity loss. Fleiss and Brennan-Prediger weight the
  rater margins differently and may show different magnitudes; the
  qualitative ordering is expected to carry over because the bias
  mechanism is the same.
- The rater model is "latent truth + guess", not Dawid-Skene. Whether
  this matters for the AC vs IPW comparison is the subject of
  experiment 02.
- Missingness is MCAR with respect to the latent truth in every cell.
  The experiment says nothing about IPW under MAR mechanisms that
  depend on the truth.

## What this implies for the manuscript

The current Section 5.1 prose around "AC is inefficient" should soften.
The cleaner contrast in the simulation is:

- AC is consistent under exchangeability and asymptotically efficient
  relative to IPW. The variance penalty IPW pays for unused reweighting
  is visible at `n = `r n_str`` and grows with `var(pi_j)`.
- AC is inconsistent under non-exchangeable raters. Its bias dominates
  any variance gap once `pi_j` varies, and IPW wins MSE.

The experiment supports replacing the "inefficient" framing with an
"inconsistent under non-exchangeability" framing in the body of the
paper.

## Reproduce

```bash
cd experiments/03-ac-vs-ipw-efficiency
Rscript run_experiment.R          # full run, ~10-15 min on a laptop
Rscript run_experiment.R --smoke  # 10 replicates per cell, < 1 min
quarto render report.qmd
```

Inputs: `misskappa` R package loaded directly with `library(misskappa)`.
Outputs: `results/mse.csv`, `results/replicates.csv`, `results/metadata.csv`.
Seed base controllable with `--seed-base`.
