---
title: "Experiment 21: Kelley & Pornprasertmanit (2016) Design, Categorical Items"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
if (!dir.exists(res_dir)) {
  res_dir <- file.path("experiments", "21-alpha-kp2016-categorical", "results")
}
needed <- c("truth.csv", "summary.csv", "metadata.csv")
missing <- needed[!file.exists(file.path(res_dir, needed))]
if (length(missing)) {
  stop(
    "Missing results: ", paste(missing, collapse = ", "),
    ". Run: Rscript experiments/21-alpha-kp2016-categorical/run_experiment.R --smoke"
  )
}
truth <- read.csv(file.path(res_dir, "truth.csv"), stringsAsFactors = FALSE)
summary <- read.csv(file.path(res_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)

meta_val <- function(k) {
  v <- metadata$value[metadata$key == k]
  if (length(v)) v[[1L]] else NA_character_
}
fmt <- function(x, d = 3) ifelse(is.na(x), "", formatC(x, format = "f", digits = d))
is_smoke <- identical(meta_val("smoke"), "TRUE")
```

## Question

Kelley & Pornprasertmanit (2016, *Psychological Methods* 21:69--92) is the
modern complete-data evaluation of confidence-interval methods for composite
reliability. Its continuous Study 1 crosses sample size, number of items,
factor-loading pattern, population reliability, and item distribution. **Does
that design, adapted to ordered-categorical items, reproduce the alpha-missing
estimator split --- and where is saturated Cat-FIML even feasible?**

The adaptation keeps the K&P factors but (i) generates ordered-categorical
items by thresholding a single-factor Gaussian copula, (ii) calibrates loadings
to a target *latent* alpha, and (iii) reports the exact post-categorisation
alpha as truth (the Maydeu-Olivares, Coffman & Hartmann 2007 convention). The
congeneric ("graded", loadings .2 -> .8) pattern is the primary baseline; the
equal/tau pattern is a contrast.

## Short answer

```{r short}
n_dgp <- nrow(truth)
feas <- aggregate(cat_em_feasible ~ categories, data = truth, FUN = function(z) sum(z) / length(z))
max_feasible_J <- tapply(truth$n_items[truth$cat_em_feasible],
                         truth$categories[truth$cat_em_feasible],
                         max)
clamped <- sum(truth$loading_clamped)
```

- The design runs end to end: **`r n_dgp` DGP cells** were generated and all
  four estimators (pairwise, Cat-FIML, normal FIML, listwise) fit under each
  mechanism.
- **Cat-FIML is the binding constraint.** The saturated `C^J` table is only
  tractable for small scales; in this grid Cat-FIML is feasible up to
  `r paste(sprintf("J=%s for C=%s", max_feasible_J, names(max_feasible_J)), collapse=", ")`.
  Beyond that it is skipped, by design. Pairwise and normal FIML carry the
  longer K&P scales; this is itself the estimator-selection story.
- Two of the four estimators (pairwise, normal FIML) need no categorical
  structure, so the experiment can sweep the full K&P item-count range for
  them while Cat-FIML is confined to the short-scale corner.
- `r if (is_smoke) "These results are a **smoke run** (tiny grid, few reps); coverage/bias numbers are illustrative, not conclusive." else "Coverage and bias are summarised below; see results/summary.csv for the full grid."`

## Evidence

### DGP catalogue and Cat-FIML feasibility

```{r truth-table, results='asis'}
tt <- truth[order(truth$categories, truth$n_items, truth$loading,
                  truth$target_latent_alpha), ]
show <- data.frame(
  DGP = tt$dgp,
  Items = tt$n_items,
  Loadings = tt$loading,
  Cats = tt$categories,
  Shape = tt$shape,
  `Item skew` = fmt(tt$item_skewness, 2),
  `Target a` = fmt(tt$target_latent_alpha, 2),
  `Latent a` = fmt(tt$latent_alpha, 3),
  `Pop alpha` = fmt(tt$alpha, 3),
  Clamped = ifelse(tt$loading_clamped, "yes", ""),
  `C^J cells` = tt$cells,
  `Cat-FIML` = ifelse(tt$cat_em_feasible, "run", "skip"),
  check.names = FALSE
)
knitr::kable(show, align = "l")
```

The `Pop alpha` column is the truth used for bias and coverage. Where
`Clamped = yes`, the target latent alpha was infeasible for that
items/loading combination (a standardized single-factor model caps the
loadings), so the largest attainable value was used instead --- this bites
only for short *graded* scales.

### Coverage and bias by estimator

```{r summary-table, results='asis'}
keep <- c("mechanism", "method", "n", "n_items", "categories", "shape",
          "truth", "bias", "se_over_sd", "coverage95", "n_valid", "skipped")
s <- summary[, keep]
s <- s[order(s$mechanism, s$method, s$categories, s$n_items, s$n), ]
disp <- data.frame(
  Mechanism = s$mechanism, Method = s$method, n = s$n, Items = s$n_items,
  Cats = s$categories, Shape = s$shape, Truth = fmt(s$truth, 3),
  Bias = fmt(s$bias, 4), `SE/SD` = fmt(s$se_over_sd, 2),
  `Cover .95` = fmt(s$coverage95, 3), Valid = s$n_valid, Skipped = s$skipped,
  check.names = FALSE
)
knitr::kable(disp, align = "l")
```

`SE/SD` near 1 means the estimated standard error matches the Monte-Carlo
spread of the estimator; `Cover .95` should sit near 0.95. Skipped cells are
Cat-FIML cases where `C^J` exceeded the feasibility cap.

## Caveats

- **Smoke vs. full.** `r if (is_smoke) "This render is from a smoke run." else "This render is from a full run."` The default
  grid is large; the headline numbers need the full design and several hundred
  replications to be trustworthy.
- **Loading clamp at small J.** A short *graded* congeneric scale cannot reach
  high reliability in a standardized model, so high targets clamp. Use the
  equal/tau pattern, or more items, to span the full reliability range. The
  realized `Pop alpha` (not the target label) is always the estimand.
- **Cat-FIML cost.** The saturated multinomial likelihood scales as `C^J`; it
  is omitted, not approximated, on the larger scales. This is a property of the
  estimator, and a reason the paper leans on pairwise and normal FIML for wide
  scales.
- This experiment is a design *skeleton*, not a parity replication of K&P; it
  targets coefficient alpha (the paper's estimand), not omega.

## Reproduce

```bash
# Smoke check (seconds):
Rscript experiments/21-alpha-kp2016-categorical/run_experiment.R --smoke --progress

# Full K&P-range run (overnight; tune as needed):
Rscript experiments/21-alpha-kp2016-categorical/run_experiment.R \
  --items 4,8,12,16,20 --n-grid 50,100,200,400,1000 \
  --alphas 0.7,0.8,0.9 --loadings equal,graded \
  --categories 2,5 --shapes symmetric,skew \
  --mechanisms complete,mcar30,mar_anchor30 \
  --reps 500 --progress
```

```{r provenance, results='asis'}
cat(sprintf("\n*Generated %s; misskappa %s; %s DGPs; seed base %s.*\n",
            meta_val("generated_at"), meta_val("misskappa_version"),
            meta_val("n_dgps"), meta_val("seed_base")))
```
