---
title: "Experiment 08: Literature Bootstrap Intervals"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/archive/pre-redesign/08-quadratic-bootstrap-literature/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

method_order <- c(
  "moss_basic", "moss_arcsine", "moss_fisher",
  "boot_percentile", "boot_basic", "boot_bca", "boot_t",
  "boot_t_arcsine", "boot_t_fisher"
)
summary$method <- factor(summary$method, levels = method_order)
focus <- summary[summary$coefficient %in% c("Conger", "Fleiss"), ]
focus <- focus[order(focus$target, focus$coefficient, focus$R, focus$n, focus$method), ]

overall <- aggregate(
  cbind(coverage, miss_below, miss_above, mean_length, valid_reps) ~ target + method,
  data = focus,
  FUN = mean
)
overall <- overall[order(overall$target, overall$method), ]
overall_tab <- data.frame(
  Target = overall$target,
  Method = overall$method,
  Coverage = fmt(overall$coverage),
  `Miss Below` = fmt(overall$miss_below),
  `Miss Above` = fmt(overall$miss_above),
  `Mean Length` = fmt(overall$mean_length),
  `Mean Valid Reps` = fmt(overall$valid_reps, 1),
  check.names = FALSE
)

worst <- focus[order(focus$coverage), ][1, ]
best_by_cell <- do.call(rbind, lapply(split(
  focus,
  interaction(focus$target, focus$coefficient, focus$R, focus$n)
), function(d) {
  d[which.min(abs(d$coverage - 0.95)), ]
}))
best_tab <- data.frame(
  Target = best_by_cell$target,
  Coefficient = best_by_cell$coefficient,
  R = best_by_cell$R,
  n = best_by_cell$n,
  Method = best_by_cell$method,
  Coverage = fmt(best_by_cell$coverage),
  `Mean Length` = fmt(best_by_cell$mean_length),
  check.names = FALSE
)

worst_tab <- head(focus[order(focus$coverage), ], 12)
worst_tab <- data.frame(
  Target = worst_tab$target,
  Coefficient = worst_tab$coefficient,
  R = worst_tab$R,
  n = worst_tab$n,
  Method = worst_tab$method,
  Coverage = fmt(worst_tab$coverage),
  `Miss Below` = fmt(worst_tab$miss_below),
  `Miss Above` = fmt(worst_tab$miss_above),
  `Mean Length` = fmt(worst_tab$mean_length),
  check.names = FALSE
)
```

## Question

Do standard literature bootstrap intervals improve the high-agreement coverage problem for complete-data categorical quadratic kappa?

## Short Answer

This is a deliberately small probe. Moss (2024) compares basic, arcsine, and Fisher transformed t intervals, and notes Efron's bootstrap-t as the natural second-order candidate while warning that categorical quadratic kappas have discrete, irregular sampling distributions. In this run the warning matters: the bootstrap family is not automatically better. The worst coverage is `r fmt(worst$coverage)` for target `r worst$target`, `r worst$coefficient`, `R = r worst$R`, `n = r worst$n`, method `r worst$method`.

The original-scale bootstrap-t has the best average coverage among the bootstrap methods, but it pays with much longer intervals and some invalid replicates in the hardest cells. The transformed bootstrap-t variants are shorter, but they are not clearly better than the Moss Fisher interval. Plain percentile/basic intervals inherit the upper-boundary discreteness instead of repairing it; BCa helps in some larger-rater cells, but it does not fix the worst `R = 2`, `n = 10` corner.

## Evidence

`Miss Below` means the interval upper endpoint is below the true coefficient; `Miss Above` means the lower endpoint is above it.

Average over coefficients, rater counts, and sample sizes:

```{r}
knitr::kable(overall_tab, align = "rlrrrrr")
```

Worst cells:

```{r}
knitr::kable(worst_tab, align = "rlrrlrrrr")
```

Closest-to-nominal method by cell:

```{r}
knitr::kable(best_tab, align = "rlrrlrr")
```

## Methods

The data-generating model is the complete categorical Perreault-Leigh setup used by Moss (2024): five uniform categories, complete rectangular ratings, and knowledge probability `sqrt(target)`, so the population quadratic coefficient equals `target`.

The analytic methods are the Moss (2024) t intervals on the identity, arcsine, and Fisher scales. The bootstrap methods resample items nonparametrically:

- `boot_percentile`: ordinary percentile interval.
- `boot_basic`: reverse-percentile/basic bootstrap interval.
- `boot_bca`: bias-corrected and accelerated interval using leave-one-item jackknife acceleration.
- `boot_t`: Efron's studentized bootstrap-t interval on the original scale.
- `boot_t_arcsine`, `boot_t_fisher`: the same studentized bootstrap-t construction after the arcsine or Fisher transform.

## Caveats

The stored run is intentionally small because each Monte Carlo replicate nests a bootstrap and a jackknife. The aim is to decide which literature methods deserve a larger run, not to estimate coverage to publication precision.

## Reproduce

Smoke run:

```sh
Rscript experiments/archive/pre-redesign/08-quadratic-bootstrap-literature/run_experiment.R --smoke
```

Exploratory run used here:

```sh
Rscript experiments/archive/pre-redesign/08-quadratic-bootstrap-literature/run_experiment.R --reps 120 --boot-reps 149 --target-grid 0.8,0.9 --n-grid 10,20,40 --r-grid 2,5 --progress
```

This report used n grid `r meta[["n_grid"]]`, R grid `r meta[["r_grid"]]`, reps `r meta[["reps"]]`, bootstrap reps `r meta[["boot_reps"]]`, target grid `r meta[["target_grid"]]`, and seed base `r meta[["seed_base"]]`.
