---
title: "Counts-FIML Under Sampling Misspecification"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- tryCatch(knitr::current_input(), error = function(e) NA_character_)
base_dir <- if (is.na(input)) getwd() else dirname(normalizePath(input, mustWork = FALSE))
candidate_dirs <- c(
  file.path(base_dir, "results"),
  "results",
  file.path("experiments", "04-counts-sampling-misspec", "results")
)
result_dir <- candidate_dirs[file.exists(file.path(candidate_dirs, "summary.csv"))][1L]
if (is.na(result_dir)) {
  stop("Missing results. Run: Rscript experiments/04-counts-sampling-misspec/run_experiment.R")
}

summary <- read.csv(file.path(result_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(result_dir, "metadata.csv"), stringsAsFactors = FALSE)
category_shift <- read.csv(file.path(result_dir, "category_shift.csv"), stringsAsFactors = FALSE)
mechanisms <- read.csv(file.path(result_dir, "mechanisms.csv"), stringsAsFactors = FALSE)

meta <- function(key) metadata$value[match(key, metadata$key)]
fmt <- function(x, digits = 4) ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
fmt_bias_sd <- function(b, s) paste0(fmt(b), " (", fmt(s), ")")

mechanism_label <- setNames(mechanisms$label, mechanisms$mechanism)
estimator_label <- c(
  counts_fiml = "Counts FIML",
  raw_fiml = "Raw FIML",
  raw_ipw = "Raw IPW"
)

get_bias <- function(mechanism, estimator, coefficient = "Fleiss") {
  idx <- summary$mechanism == mechanism &
    summary$estimator == estimator &
    summary$coefficient == coefficient
  summary$bias[idx][1L]
}

m1_counts <- get_bias("m1_rater_mcar", "counts_fiml")
m1_raw_fiml <- get_bias("m1_rater_mcar", "raw_fiml")
m1_raw_ipw <- get_bias("m1_rater_mcar", "raw_ipw")
m2_counts <- get_bias("m2_value_dropout", "counts_fiml")
m2_raw_fiml <- get_bias("m2_value_dropout", "raw_fiml")
m2_raw_ipw <- get_bias("m2_value_dropout", "raw_ipw")
```

## Question

How badly does counts-FIML break when Assumption (S), uniform
hypergeometric subsampling of observed counts from full counts, fails?

The experiment uses two failure mechanisms. M1 has heterogeneous raters
with rater-specific MCAR observation rates, then discards rater labels
before fitting counts-FIML. M2 has exchangeable raters but value-dependent
dropout, so low categories are less often observed.

## Short Answer

In M1, counts-FIML is the estimator that moves. Fleiss bias is
`r fmt(m1_counts)`, compared with `r fmt(m1_raw_fiml)` for raw FIML and
`r fmt(m1_raw_ipw)` for raw IPW. This is the clean Assumption (S)
failure. Rater labels contain the information needed to undo the
rater-specific observation process.

In M2, the missingness is value-dependent, so the problem is NMAR for
all three estimators. Fleiss bias is `r fmt(m2_counts)` for counts-FIML,
`r fmt(m2_raw_fiml)` for raw FIML, and `r fmt(m2_raw_ipw)` for raw IPW.
This cell is a boundary case rather than a counts-specific repair story.

## Evidence

```{r}
summary_display <- summary
summary_display$Mechanism <- mechanism_label[summary_display$mechanism]
summary_display$Estimator <- estimator_label[summary_display$estimator]
summary_display$`Coefficient` <- summary_display$coefficient
summary_display$`Truth` <- fmt(summary_display$truth)
summary_display$`Mean Estimate` <- fmt(summary_display$mean_estimate)
summary_display$`Bias (SD)` <- fmt_bias_sd(summary_display$bias, summary_display$sd)
summary_display$`RMSE` <- fmt(summary_display$rmse)
summary_display$`Failures` <- summary_display$failures
summary_display <- summary_display[order(
  summary_display$Mechanism, summary_display$Coefficient, summary_display$Estimator
), c("Mechanism", "Estimator", "Coefficient", "Truth", "Mean Estimate", "Bias (SD)", "RMSE", "Failures")]
knitr::kable(summary_display, align = "lllrrrrr")
```

The category marginals below show the sampling distortion after rater
labels are removed.

```{r}
shift_display <- category_shift
shift_display$Mechanism <- mechanism_label[shift_display$mechanism]
shift_display$Category <- shift_display$category
shift_display$`Full Share` <- fmt(shift_display$full_share)
shift_display$`Observed Share` <- fmt(shift_display$observed_share)
shift_display$`Observed - Full` <- fmt(shift_display$shift)
shift_display <- shift_display[order(shift_display$Mechanism, shift_display$Category),
                               c("Mechanism", "Category", "Full Share", "Observed Share", "Observed - Full")]
knitr::kable(shift_display, align = "lrrrr")
```

## Caveats

The truth values are analytic expectations under the complete-data
rating model. The experiment used `r meta("reps")` replicates at
$n = `r meta("n")`$, so the Monte Carlo standard error of the reported
bias is the `mc_se_bias` column in `results/summary.csv`.

M1 is designed to isolate a counts-model failure after aggregation. M2 is
designed to show the broader boundary under value-dependent dropout.

## Reproduce

```sh
Rscript experiments/04-counts-sampling-misspec/run_experiment.R
```

For a cheap check:

```sh
Rscript experiments/04-counts-sampling-misspec/run_experiment.R --smoke
```
