---
title: "Experiment 18: Alpha Missing Paper Simulation"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
if (!dir.exists(res_dir)) {
  res_dir <- file.path("experiments", "studies", "18-alpha-paper-simulation", "results")
}
needed <- c("truth.csv", "summary.csv", "metadata.csv")
missing <- needed[!file.exists(file.path(res_dir, needed))]
if (length(missing)) {
  stop(
    "Missing results: ", paste(missing, collapse = ", "),
    ". Run: Rscript experiments/studies/18-alpha-paper-simulation/run_experiment.R --smoke"
  )
}
truth <- read.csv(file.path(res_dir, "truth.csv"), stringsAsFactors = FALSE)
summary <- read.csv(file.path(res_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)

fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
```

## Question

For the standalone alpha-missing paper, do scored-ordinal item models and clean
MCAR/MAR mechanisms produce the expected qualitative split: pairwise alpha has
usable inference under MCAR but is vulnerable under MAR, while saturated
categorical FIML and normal FIML are the MAR comparators?

The item responses are three-category scored ordinal variables obtained by
thresholding a latent Gaussian copula. Population alpha is computed from the
exact scored-category covariance induced by the thresholds and latent
correlations, not from the latent normal covariance.

## Short Answer

```{r short-answer}
smoke <- metadata$value[metadata$key == "smoke"]
reps <- metadata$value[metadata$key == "reps"]
```

This is a `r ifelse(identical(smoke, "TRUE"), "smoke", "configured")` run with
`r reps` replications per cell. It is mainly an API and design check. The
important thing to inspect is whether the MCAR rows keep pairwise, Cat-FIML, and
normal FIML centered on the same truth, while observed-anchor MAR starts
separating the pairwise/listwise estimates from the likelihood estimators.

## Truths

```{r truths}
truth_tab <- truth[, c("dgp", "family", "p", "categories", "truth_alpha",
                       "min_latent_cor", "max_latent_cor",
                       "min_score_cov", "max_score_cov")]
truth_tab$truth_alpha <- fmt(truth_tab$truth_alpha)
truth_tab$min_latent_cor <- fmt(truth_tab$min_latent_cor)
truth_tab$max_latent_cor <- fmt(truth_tab$max_latent_cor)
truth_tab$min_score_cov <- fmt(truth_tab$min_score_cov)
truth_tab$max_score_cov <- fmt(truth_tab$max_score_cov)
knitr::kable(
  truth_tab,
  col.names = c("DGP", "Family", "Items", "Categories", "Alpha",
                "Min Latent Cor", "Max Latent Cor",
                "Min Score Cov", "Max Score Cov")
)
```

## Evidence

```{r evidence-table}
display <- summary[, c("dgp", "mechanism", "n", "method", "n_valid",
                       "truth", "bias", "sd_estimate", "mean_se",
                       "coverage95", "mean_observed_fraction",
                       "median_elapsed_ms")]
display$truth <- fmt(display$truth)
display$bias <- fmt(display$bias)
display$sd_estimate <- fmt(display$sd_estimate)
display$mean_se <- fmt(display$mean_se)
display$coverage95 <- fmt(display$coverage95, 3)
display$mean_observed_fraction <- fmt(display$mean_observed_fraction, 3)
display$median_elapsed_ms <- fmt(display$median_elapsed_ms, 1)
knitr::kable(
  display,
  col.names = c("DGP", "Mechanism", "n", "Method", "Valid",
                "Truth", "Bias", "MC SD", "Mean SE", "Coverage",
                "Obs Frac", "Median Time (ms)")
)
```

```{r bias-plot, fig.width=7, fig.height=4}
old_par <- par(no.readonly = TRUE)
on.exit(par(old_par), add = TRUE)
plot_df <- summary[summary$method %in% c("pairwise", "cat_em", "normal_fiml", "listwise"), ]
plot_df$cell <- paste(plot_df$dgp, plot_df$mechanism, sep = "\n")
cols <- c(pairwise = "#000000", cat_em = "#009E73",
          normal_fiml = "#0072B2", listwise = "#D55E00")
methods <- intersect(names(cols), unique(plot_df$method))
x <- seq_along(unique(plot_df$cell))
cells <- unique(plot_df$cell)
ylim <- range(plot_df$bias, na.rm = TRUE)
plot(x, rep(NA_real_, length(x)), ylim = ylim, xaxt = "n",
     xlab = "", ylab = "Bias", main = "Bias by DGP and mechanism")
abline(h = 0, col = "gray70", lty = 2)
offsets <- seq(-0.18, 0.18, length.out = length(methods))
for (m_idx in seq_along(methods)) {
  m <- methods[m_idx]
  g <- plot_df[plot_df$method == m, ]
  points(match(g$cell, cells) + offsets[m_idx], g$bias,
         pch = 16, col = cols[m])
}
axis(1, at = x, labels = cells, las = 2, cex.axis = 0.75)
legend("topright", legend = methods, pch = 16, col = cols[methods], bty = "n")
```

## Caveats

- This is the scored-ordinal arm only, capped to six three-category items so the
  saturated categorical FIML state space stays small enough for repeated fits.
- Pairwise coverage uses the package's overlap-subsample SE from
  `alpha(method = "available")`.
- `cat_em` is the saturated categorical FIML estimator from
  `alpha(method = "fiml")`; listwise is retained only as a sanity check.
- `normal_fiml` treats the scored ordinal categories as continuous and uses
  `alpha_continuous(se_type = "sandwich")`.
- Smoke results are not paper evidence. Use them to catch implementation and
  design problems before the `B = 500` or larger pass.

## Reproduce

Smoke:

```sh
Rscript experiments/studies/18-alpha-paper-simulation/run_experiment.R --smoke --progress
quarto render experiments/studies/18-alpha-paper-simulation/report.qmd
```

Main ordinal pass:

```sh
Rscript experiments/studies/18-alpha-paper-simulation/run_experiment.R \
  --reps 500 \
  --n-grid 200,500,1000 \
  --dgps ordinal_tau6,ordinal_congeneric6,ordinal_twofactor6 \
  --mechanisms mcar15,mcar30,mar_anchor15,mar_anchor30 \
  --methods pairwise,cat_em,normal_fiml,listwise \
  --progress
```

```{r metadata}
knitr::kable(metadata, col.names = c("Key", "Value"))
```
