---
title: "Experiment 16: Alpha Calibration Sweep"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- tryCatch(knitr::current_input(), error = function(e) NA_character_)
base_dir <- if (is.na(input)) getwd() else dirname(normalizePath(input, mustWork = FALSE))
candidate_dirs <- c(
  file.path(base_dir, "results"),
  "results",
  file.path("experiments", "16-alpha-calibration-sweep", "results")
)
result_dir <- candidate_dirs[file.exists(file.path(candidate_dirs, "summary.csv"))][1L]
if (is.na(result_dir)) {
  stop("Missing results. Run: Rscript experiments/16-alpha-calibration-sweep/run_experiment.R --smoke",
       call. = FALSE)
}

summary <- read.csv(file.path(result_dir, "summary.csv"), stringsAsFactors = FALSE)
truth <- read.csv(file.path(result_dir, "truth.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(result_dir, "metadata.csv"), stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
fmt_ms <- function(x) ifelse(is.na(x), "", formatC(x, format = "f", digits = 1))
fmt_sci <- function(x, digits = 2) ifelse(is.na(x), "", formatC(x, format = "e", digits = digits))

total_failures <- sum(summary$failures)
max_if_error <- max(summary$max_psi_vcov_abs_error, na.rm = TRUE)
if (!is.finite(max_if_error)) max_if_error <- NA_real_

largest_profile <- truth[which.max(truth$patterns), ]
largest_timing <- subset(
  summary,
  profile == largest_profile$profile[1L] & method == "fiml" & mechanism == "mar"
)
largest_timing <- if (nrow(largest_timing)) {
  largest_timing[which.max(largest_timing$n), ]
} else {
  data.frame(median_elapsed_ms = NA_real_)
}

fiml_rows <- subset(summary, method == "fiml")
worst_fiml <- if (nrow(fiml_rows)) {
  fiml_rows[which.max(abs(fiml_rows$bias)), ]
} else {
  data.frame(profile_label = "", mechanism = "", n = NA_integer_, bias = NA_real_)
}
```

## Question

Against fixed population alpha truths, how do `alpha(method =
"available")` and saturated categorical `alpha(method = "fiml")` behave
under complete data, MCAR missingness, and an ignorable anchor-item MAR
mechanism across small `C x R x n` grids?

The population truth is computed by one-dimensional quadrature under the
latent ordinal item model, rather than by a large simulated complete
sample. The Louis-information path is checked externally through mean
estimated SE, Wald coverage, timing, and reconstruction of `vcov()` from
the exported per-subject influence rows.

## Short Answer

This is a capped calibration run, not the final paper sweep. It found
`r total_failures` fit failures in the committed run, and the largest
influence-to-vcov reconstruction error was `r fmt_sci(max_if_error)`.
For the largest profile in this run (`r largest_profile$profile_label[1]`,
`C^R = r largest_profile$patterns[1]`), median FIML time in the MAR cell
at the largest displayed `n` was `r fmt_ms(largest_timing$median_elapsed_ms[1])`
ms.

The largest absolute FIML bias in the committed run was
`r fmt(worst_fiml$bias[1])` in the `r worst_fiml$profile_label[1]`,
`r worst_fiml$mechanism[1]`, `n = r worst_fiml$n[1]` cell. That is a
finite-sample/sparse-support warning rather than a conclusion; it is
the cell to revisit with a larger `n` grid and replication budget.

Use this to check mechanics and choose the full replication budget. The
coverage columns are still Monte Carlo noisy unless `--reps` is raised.

## Truths

```{r}
truth_display <- truth
truth_display$`Profile` <- truth_display$profile_label
truth_display$`C` <- truth_display$C
truth_display$`R` <- truth_display$R
truth_display$`C^R` <- truth_display$patterns
truth_display$`Alpha` <- fmt(truth_display$truth_alpha)
truth_display$`Total Var` <- fmt(truth_display$total_variance)
truth_display$`Item Var Sum` <- fmt(truth_display$item_variance_sum)
truth_display <- truth_display[, c("Profile", "C", "R", "C^R", "Alpha",
                                   "Total Var", "Item Var Sum")]
knitr::kable(truth_display, row.names = FALSE, align = "lrrrrrr")
```

## Calibration

```{r}
cal <- summary[, c(
  "profile_label", "mechanism", "n", "method", "truth", "reps",
  "n_valid", "bias", "mc_se_bias", "sd_estimate", "mean_se",
  "se_over_sd", "coverage95", "mean_z", "sd_z", "median_elapsed_ms",
  "mean_observed_fraction", "failures"
)]
cal <- cal[order(cal$profile_label, cal$mechanism, cal$n, cal$method), ]
cal$Profile <- cal$profile_label
cal$Mechanism <- cal$mechanism
cal$Method <- cal$method
cal$Truth <- fmt(cal$truth)
cal$Bias <- fmt(cal$bias)
cal$`MC SE Bias` <- fmt(cal$mc_se_bias)
cal$`SD Est` <- fmt(cal$sd_estimate)
cal$`Mean SE` <- fmt(cal$mean_se)
cal$`SE/SD` <- fmt(cal$se_over_sd, 2)
cal$`Cover 95` <- fmt(cal$coverage95)
cal$`Mean z` <- fmt(cal$mean_z, 2)
cal$`SD z` <- fmt(cal$sd_z, 2)
cal$`Median ms` <- fmt_ms(cal$median_elapsed_ms)
cal$`Obs Frac` <- fmt(cal$mean_observed_fraction)
cal$Failures <- cal$failures
cal <- cal[, c("Profile", "Mechanism", "n", "Method", "Truth", "reps",
               "n_valid", "Bias", "MC SE Bias", "SD Est", "Mean SE",
               "SE/SD", "Cover 95", "Mean z", "SD z", "Median ms",
               "Obs Frac", "Failures")]
knitr::kable(cal, row.names = FALSE, align = "llrlrrrrrrrrrrrrrr")
```

## Diagnostics

```{r}
diag <- summary[, c("profile_label", "mechanism", "n", "method",
                    "min_pair_count_min", "min_observed_categories",
                    "max_psi_vcov_abs_error")]
diag <- diag[order(diag$profile_label, diag$mechanism, diag$n, diag$method), ]
diag$Profile <- diag$profile_label
diag$Mechanism <- diag$mechanism
diag$Method <- diag$method
diag$`Min Pair n` <- diag$min_pair_count_min
diag$`Min Categories` <- diag$min_observed_categories
diag$`Max IF Vcov Error` <- formatC(diag$max_psi_vcov_abs_error, format = "e", digits = 2)
diag <- diag[, c("Profile", "Mechanism", "n", "Method", "Min Pair n",
                 "Min Categories", "Max IF Vcov Error")]
knitr::kable(diag, row.names = FALSE, align = "llrlrrr")
```

## Caveats

- The alpha-specific Louis eigenspectrum is not exposed by the R package
  yet. This runner checks SE calibration and IF reconstruction, but a
  true spectrum diagnostic would need a small alpha analogue of
  `rcpp_fiml_louis_spectrum`.
- The MAR mechanism observes item 1 for everyone and lets the other
  items' observation probabilities depend on item 1. That is ignorable
  for the saturated FIML model, but it is only one MAR design.
- If a finite sample drops an entire category, the R wrapper estimates
  on the observed support. The summary records the minimum observed
  category count so support loss is visible.
- The committed run uses `--paper-n-grid 250,1000`. An uncapped
  `paper5x6`, `n = 4000`, `B = 200` pass was interrupted after it
  stayed in the first high-dimensional `n = 4000` cell for several
  minutes. Treat that as a feasibility finding: `5^6` needs separate
  scheduled benchmarking or a smaller replication grid.
- The committed run is intended to exercise the full pipeline and give
  first-order calibration signals. Raise `--reps` further before using
  coverage or SE ratios as manuscript evidence.

## Reproduce

```bash
Rscript experiments/16-alpha-calibration-sweep/run_experiment.R --smoke
```

Starter sweep:

```bash
Rscript experiments/16-alpha-calibration-sweep/run_experiment.R --profiles tiny3x4,ordinal4x5,paper5x6 --n-grid 150,500 --reps 8 --progress
```

Fuller first pass:

```bash
Rscript experiments/16-alpha-calibration-sweep/run_experiment.R \
  --reps 200 \
  --n-grid 250,1000,4000 \
  --paper-n-grid 250,1000 \
  --progress
```

```{r}
cat(sprintf(
  "Metadata: smoke=%s; profiles=%s; mechanisms=%s; n_grid=%s; paper_n_grid=%s; reps=%s; seed_base=%s; truth=%s; misskappa=%s\n",
  meta[["smoke"]], meta[["profiles"]], meta[["mechanisms"]], meta[["n_grid"]],
  meta[["paper_n_grid"]], meta[["reps"]], meta[["seed_base"]], meta[["truth_method"]],
  meta[["misskappa_version"]]
))
```
