---
title: "Experiment 18: Alpha Missing Paper Simulation"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
if (!dir.exists(res_dir)) {
  res_dir <- file.path("experiments", "18-alpha-paper-simulation", "results")
}
needed <- c("truth.csv", "summary.csv", "metadata.csv")
missing <- needed[!file.exists(file.path(res_dir, needed))]
if (length(missing)) {
  stop(
    "Missing results: ", paste(missing, collapse = ", "),
    ". Run: Rscript experiments/18-alpha-paper-simulation/run_experiment.R --smoke"
  )
}
truth <- read.csv(file.path(res_dir, "truth.csv"), stringsAsFactors = FALSE)
summary <- read.csv(file.path(res_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)

fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
```

## Question

For the standalone alpha-missing paper, do the paper-backed continuous item
models and clean MCAR/MAR mechanisms produce the expected qualitative split:
pairwise alpha is acceptable under MCAR but vulnerable under MAR, while
normal-FIML is the MAR workhorse?

This first pass exercises the continuous arm. Pairwise coverage is intentionally
blank because the package does not yet expose the overlapping-subsample
pairwise SE.

## Short Answer

```{r short-answer}
smoke <- metadata$value[metadata$key == "smoke"]
reps <- metadata$value[metadata$key == "reps"]
```

This is a `r ifelse(identical(smoke, "TRUE"), "smoke", "configured")` run with
`r reps` replications per cell. It is mainly an API and design check. The
important thing to inspect is whether the MCAR rows keep pairwise and FIML
centered on the same truth, while observed-anchor MAR starts separating the
pairwise/listwise estimates from FIML.

## Truths

```{r truths}
truth_tab <- truth[, c("dgp", "family", "p", "truth_alpha",
                       "min_loading", "max_loading",
                       "min_residual_var", "max_residual_var")]
truth_tab$truth_alpha <- fmt(truth_tab$truth_alpha)
truth_tab$min_loading <- fmt(truth_tab$min_loading)
truth_tab$max_loading <- fmt(truth_tab$max_loading)
truth_tab$min_residual_var <- fmt(truth_tab$min_residual_var)
truth_tab$max_residual_var <- fmt(truth_tab$max_residual_var)
knitr::kable(
  truth_tab,
  col.names = c("DGP", "Family", "Items", "Alpha",
                "Min Loading", "Max Loading",
                "Min Residual Var", "Max Residual Var")
)
```

## Evidence

```{r evidence-table}
display <- summary[, c("dgp", "mechanism", "n", "method", "n_valid",
                       "truth", "bias", "sd_estimate", "mean_se",
                       "coverage95", "mean_observed_fraction",
                       "median_elapsed_ms")]
display$truth <- fmt(display$truth)
display$bias <- fmt(display$bias)
display$sd_estimate <- fmt(display$sd_estimate)
display$mean_se <- fmt(display$mean_se)
display$coverage95 <- fmt(display$coverage95, 3)
display$mean_observed_fraction <- fmt(display$mean_observed_fraction, 3)
display$median_elapsed_ms <- fmt(display$median_elapsed_ms, 1)
knitr::kable(
  display,
  col.names = c("DGP", "Mechanism", "n", "Method", "Valid",
                "Truth", "Bias", "MC SD", "Mean SE", "Coverage",
                "Obs Frac", "Median Time (ms)")
)
```

```{r bias-plot, fig.width=7, fig.height=4}
old_par <- par(no.readonly = TRUE)
on.exit(par(old_par), add = TRUE)
plot_df <- summary[summary$method %in% c("pairwise", "fiml_sandwich", "listwise"), ]
plot_df$cell <- paste(plot_df$dgp, plot_df$mechanism, sep = "\n")
cols <- c(pairwise = "#000000", fiml_sandwich = "#009E73", listwise = "#D55E00")
methods <- intersect(names(cols), unique(plot_df$method))
x <- seq_along(unique(plot_df$cell))
cells <- unique(plot_df$cell)
ylim <- range(plot_df$bias, na.rm = TRUE)
plot(x, rep(NA_real_, length(x)), ylim = ylim, xaxt = "n",
     xlab = "", ylab = "Bias", main = "Bias by DGP and mechanism")
abline(h = 0, col = "gray70", lty = 2)
offsets <- seq(-0.18, 0.18, length.out = length(methods))
for (m_idx in seq_along(methods)) {
  m <- methods[m_idx]
  g <- plot_df[plot_df$method == m, ]
  points(match(g$cell, cells) + offsets[m_idx], g$bias,
         pch = 16, col = cols[m])
}
axis(1, at = x, labels = cells, las = 2, cex.axis = 0.75)
legend("topright", legend = methods, pch = 16, col = cols[methods], bty = "n")
```

## Caveats

- This is the continuous arm only. The ordinal/categorical arm remains the next
  addition, capped to six five-category items.
- Pairwise coverage is pending the package port of the pairwise overlapping
  subsample SE. Until then, pairwise rows report point-estimate behavior and
  Monte Carlo SD only.
- `fiml_normal` is the normal-theory SE; `fiml_sandwich` is the default robust
  SE and the main continuous-MAR method.
- Smoke results are not paper evidence. Use them to catch implementation and
  design problems before the `B = 500` or larger pass.

## Reproduce

Smoke:

```sh
Rscript experiments/18-alpha-paper-simulation/run_experiment.R --smoke --progress
quarto render experiments/18-alpha-paper-simulation/report.qmd
```

Main continuous pass:

```sh
Rscript experiments/18-alpha-paper-simulation/run_experiment.R \
  --reps 500 \
  --n-grid 200,500,1000 \
  --dgps tau6_essential,congeneric6_zy,congeneric8_gradient,twofactor8 \
  --mechanisms mcar15,mcar30,mar_zy_light,mar_anchor15,mar_anchor30 \
  --methods pairwise,fiml_sandwich,fiml_normal,listwise \
  --progress
```

```{r metadata}
knitr::kable(metadata, col.names = c("Key", "Value"))
```
