---
title: "35 - Cat-FIML SQUAREM Calibration"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
results_dir <- file.path(dirname(knitr::current_input()), "results")
needed <- file.path(results_dir, c("summary.csv", "comparison.csv", "metadata.csv"))
if (!all(file.exists(needed))) {
  stop(
    "Missing results. Run: Rscript experiments/workbench/35-cat-fiml-squarem-calibration/run_experiment.R --smoke --progress",
    call. = FALSE
  )
}
summary <- read.csv(file.path(results_dir, "summary.csv"), stringsAsFactors = FALSE)
comparison <- read.csv(file.path(results_dir, "comparison.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(results_dir, "metadata.csv"), stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.finite(x), formatC(x, digits = digits, format = "fg", flag = "#"), NA)
}
fmt_sci <- function(x) {
  ifelse(is.finite(x), formatC(x, digits = 2, format = "e"), NA)
}
```

## Question

Does SQUAREM acceleration change the strict-ML categorical FIML estimates,
standard errors, or interval coverage enough to offset the runtime gain seen in
the phase-timed performance probe?

## Short Answer

This experiment is the calibration harness for that decision. It compares two
local C++ variants on matched simulated datasets: ordinary active-set EM, which
is checksum-identical to canonical strict ML in the performance probe, and the
same estimator with SQUAREM wrapped around the EM fixed-point map. The default
smoke run is only a mechanical check; use `--calibration` or a larger `--reps`
override before treating the coverage columns as policy evidence.

```{r}
headline <- comparison[comparison$coefficient == "Conger" &
                         comparison$weight_label == "quadratic",
                       c("dgp", "mechanism", "n", "paired_reps",
                         "median_abs_estimate_diff", "median_se_ratio",
                         "coverage_wald_t_em", "coverage_wald_t_squarem",
                         "coverage_fisher_t_em", "coverage_fisher_t_squarem",
                         "median_speedup", "mean_iter_ratio")]
headline <- headline[order(headline$dgp, headline$mechanism, headline$n), ]
headline_out <- transform(
  headline,
  median_abs_estimate_diff = fmt_sci(median_abs_estimate_diff),
  median_se_ratio = fmt(median_se_ratio),
  coverage_wald_t_em = fmt(coverage_wald_t_em),
  coverage_wald_t_squarem = fmt(coverage_wald_t_squarem),
  coverage_fisher_t_em = fmt(coverage_fisher_t_em),
  coverage_fisher_t_squarem = fmt(coverage_fisher_t_squarem),
  median_speedup = fmt(median_speedup),
  mean_iter_ratio = fmt(mean_iter_ratio)
)
names(headline_out) <- c("DGP", "Mechanism", "n", "Paired Reps",
                         "Median |Delta Est|", "Median SE Ratio",
                         "Wald-t EM", "Wald-t SQUAREM",
                         "Fisher-t EM", "Fisher-t SQUAREM",
                         "Median Speedup", "Mean Iter Ratio")
knitr::kable(headline_out)
```

## Evidence

The paired comparison is computed within replicate, DGP, mechanism, sample
size, weight, and coefficient. The estimate-difference columns are therefore
algorithmic differences on the same data, not Monte Carlo noise from different
datasets.

```{r}
est_dev <- comparison[order(-comparison$p95_abs_estimate_diff),
                      c("dgp", "mechanism", "n", "weight_label", "coefficient",
                        "paired_reps", "p95_abs_estimate_diff",
                        "median_abs_estimate_diff_over_em_se",
                        "median_se_ratio", "p10_se_ratio", "p90_se_ratio")]
est_dev <- utils::head(est_dev, 12L)
est_dev_out <- transform(
  est_dev,
  p95_abs_estimate_diff = fmt_sci(p95_abs_estimate_diff),
  median_abs_estimate_diff_over_em_se = fmt(median_abs_estimate_diff_over_em_se),
  median_se_ratio = fmt(median_se_ratio),
  p10_se_ratio = fmt(p10_se_ratio),
  p90_se_ratio = fmt(p90_se_ratio)
)
names(est_dev_out) <- c("DGP", "Mechanism", "n", "Weight", "Coefficient",
                        "Paired Reps", "P95 |Delta Est|",
                        "Median |Delta| / EM SE", "Median SE Ratio",
                        "P10 SE Ratio", "P90 SE Ratio")
knitr::kable(est_dev_out)
```

```{r}
runtime <- comparison[comparison$coefficient == "Conger",
                      c("dgp", "mechanism", "n", "weight_label",
                        "median_speedup", "mean_iter_ratio",
                        "mean_support_delta")]
runtime <- runtime[order(runtime$dgp, runtime$mechanism, runtime$n, runtime$weight_label), ]
runtime_out <- transform(
  runtime,
  median_speedup = fmt(median_speedup),
  mean_iter_ratio = fmt(mean_iter_ratio),
  mean_support_delta = fmt(mean_support_delta)
)
names(runtime_out) <- c("DGP", "Mechanism", "n", "Weight",
                        "Median Speedup", "Mean Iter Ratio",
                        "Mean Support Delta")
knitr::kable(runtime_out)
```

## Caveats

The DGPs are study-29-shaped stress cells, not a literal source of study 29's
full simulation grid. They cover balanced, high-agreement, and skew-sparse
latent-class ratings with MCAR, anchor-MAR, and planned two-rater missingness.
Truth is complete-data Monte Carlo truth, so increase `--truth-n` for final
calibration runs.

The SQUAREM variant uses the active-set implementation from the performance
probe. The ordinary EM control uses the same active-set implementation, so the
statistical comparison isolates acceleration rather than the mechanical
active-set rewrite.

## Reproduce

Generated at `r meta[["generated_at"]]` with:

```{r}
cat(meta[["command"]], "\n")
```

Useful outputs:

- `results/replicates.csv`: one row per variant, replicate, coefficient, and weight.
- `results/summary.csv`: bias, RMSE, SE calibration, interval coverage, runtime, and EM support by variant.
- `results/comparison.csv`: paired EM-vs-SQUAREM estimate, SE, coverage, support, speed, and iteration ratios.
