---
title: "Equality of two dependent alphas: a worked example (cocron::knowledge)"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
read_results <- function(f) {
  p <- file.path("results", f)
  if (!file.exists(p))
    stop("Missing ", p, ". Run: Rscript run_experiment.R", call. = FALSE)
  read.csv(p, check.names = FALSE)
}
bench <- read_results("bench.csv")
miss  <- read_results("missing_summary.csv")
meta  <- read_results("metadata.csv")
mv    <- function(k) meta$value[meta$key == k]
fmt   <- function(x, d = 3) formatC(as.numeric(x), digits = d, format = "f")
```

## Question

Can the influence-function machinery in the alpha-missing paper deliver a
real-data **test of equal coefficient alpha** for two forms taken by the
same subjects (a *paired* contrast), and does that test survive
item-level missingness?

The data are Diedenhofen & Musch's `knowledge` set from the `cocron`
package: `r mv("n_subjects")` testees who each took two
`r mv("n_items_per_form")`-item binary knowledge tests. Because the same
people took both forms, the two alpha estimates are dependent.

## Short answer

Yes on both counts. `misskappa::alpha(estimator = "pairwise")` reproduces
`cocron`'s alpha for each form to the printed precision, and the
IF-based paired Wald test `alpha_test(fit_a, fit_b, paired = TRUE)`
rejects equal alpha (p ≈ `r fmt(bench$p_value[bench$test == "paired (dependent IF)"], 3)`),
agreeing in direction with `cocron`'s normal-theory dependent test. The
two differ in magnitude for a principled reason: our test uses the
robust sandwich covariance of the two alpha estimators, while `cocron`
uses the Feldt-family normal-theory variance. Under MCAR amputation the
available-case contrast stays essentially unbiased and its standard
error inflates smoothly, so the test loses power gracefully rather than
breaking.

## Evidence

### Complete data: misskappa vs cocron

```{r bench-table}
b <- bench
b$`alpha A` <- fmt(b$alpha_A, 4)
b$`alpha B` <- fmt(b$alpha_B, 4)
b$`diff`    <- fmt(b$diff, 4)
b$`corr(A,B)` <- ifelse(is.na(b$corr), "--", fmt(b$corr, 3))
b$`chi-square` <- fmt(b$statistic, 3)
b$`p-value`    <- fmt(b$p_value, 4)
knitr::kable(
  b[, c("source", "test", "alpha A", "alpha B", "diff",
        "corr(A,B)", "chi-square", "p-value")],
  align = "llrrrrrr",
  caption = "Equal-alpha tests on the complete knowledge data (df = 1)."
)
```

Two points stand out.

* **Point estimates agree exactly.** The available-case alpha equals
  `cocron`'s for both forms, so any difference in the test is about the
  *variance*, not the estimate.
* **Dependence matters, and the variance model matters.** Treating the
  forms as independent (ignoring that the same subjects took both)
  roughly halves the chi-square for both packages. Among the dependent
  tests, the IF/sandwich statistic
  (`r fmt(bench$statistic[bench$test == "paired (dependent IF)"], 2)`) is
  smaller than `cocron`'s normal-theory one
  (`r fmt(bench$statistic[bench$test == "dependent (Feldt/normal)"], 2)`):
  the sandwich estimator is robust to the non-normality of binary items
  and ties the dependence to the actual correlation of the two alpha
  estimators (≈ `r fmt(bench$corr[bench$test == "paired (dependent IF)"], 2)`),
  whereas the Feldt construction leverages the higher inter-form score
  correlation. Both reject at the 5% level.

### Missing data: MCAR robustness of the available-case contrast

```{r miss-table}
m <- miss
out <- data.frame(
  `MCAR rate`        = fmt(m$rate, 2),
  `mean alpha A`     = fmt(m$mean_alpha_A, 4),
  `mean alpha B`     = fmt(m$mean_alpha_B, 4),
  `mean diff`        = fmt(m$mean_diff, 4),
  `bias of diff`     = fmt(m$bias_diff, 4),
  `mean SE (dep)`    = fmt(m$mean_se_dep, 4),
  `power @ .05`      = fmt(m$power_05, 2),
  check.names = FALSE
)
knitr::kable(out, align = "rrrrrrr",
  caption = "Available-case paired test under MCAR amputation; bias is relative to the complete-data contrast.")
```

As the item-missing rate rises, the mean available-case contrast tracks
the complete-data value (bias near zero), the estimated standard error
of the contrast inflates monotonically, and rejection power falls
accordingly. This is exactly the behaviour the paper's MCAR theory
predicts: the estimator stays valid, it just has less information.

## Caveats

* This is an illustration, not a calibration study. The two real forms
  genuinely differ in alpha, so the "power" column is power against the
  true alternative, not a size check. A null-DGP size study lives in the
  simulation experiments (18--23), not here.
* `mean SE (dep)` is the full estimated standard error of the contrast;
  `mc_sd_diff` in `results/missing_summary.csv` is only the
  amputation-induced spread (data held fixed), so the two are not
  comparable.
* Binary items make `cocron`'s normal-theory assumption least
  favourable; the gap between the two dependent tests should narrow for
  approximately normal items.
* Saturated categorical FIML is deliberately *not* used here: a 30-item
  binary table has $2^{30}$ cells. The available-case (pairwise
  covariance) estimator is the right tool for these widths under MCAR.

## Reproduce

```bash
cd experiments/24-alpha-equal-cocron
Rscript run_experiment.R --smoke      # fast (reps = 20)
Rscript run_experiment.R              # full (reps = 300)
quarto render report.qmd
```

Run with `misskappa` `r mv("misskappa_version")`, `cocron`
`r mv("cocron_version")`, R `r mv("R_version")`.
