---
title: "Getting started with misskappa"
vignette: >
  %\VignetteIndexEntry{Getting started with misskappa}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
format: html
---

```{r}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`misskappa` estimates weighted agreement coefficients (Conger, Fleiss,
Brennan–Prediger) and coefficient alpha when some ratings are **missing**, and
attaches an asymptotic covariance matrix to every estimate so you get standard
errors, confidence intervals, and equality tests for free. Each estimator is
chosen with a single `estimator=` argument. You never have to remember a
combination of `method` and `type` flags.

```{r}
#| message: false
library(misskappa)
```

## A first agreement coefficient

`dat.gwet2014` is a small categorical rating matrix (20 subjects, five
raters, four categories) with some ratings missing (`NA`). The
inverse-probability-weighted (IPW) estimator is consistent under
missing-completely-at-random missingness:

```{r}
fit <- kappa(dat.gwet2014, estimator = "ipw", weight = "linear")
fit
```

The print method shows each coefficient with its standard error and a 95%
confidence interval. The object carries everything you need:

```{r}
coef(fit)
sqrt(diag(vcov(fit)))
```

### Confidence intervals

`confint()` gives natural-scale Wald intervals, or a variance-stabilizing
Fisher *z* interval that always stays inside `(-1, 1)`, useful when a
coefficient is close to the boundary:

```{r}
confint(fit)
confint(fit, transform = "fisher")
```

## Choosing an estimator

The right `estimator=` depends on the data and the missingness assumption:

| Input | `estimator` | Assumption |
|---|---|---|
| Categorical ratings | `"ipw"` | MCAR, rater-varying observation rates |
| Categorical ratings | `"cat_fiml"` | Ignorable (MCAR or MAR), saturated multinomial |
| Scored ratings (quadratic) | `"pairwise"` | MCAR, distribution-free |
| Scored ratings (quadratic) | `"nt_fiml"` | Ignorable, normal-theory |

The scored estimators treat the categories as numbers and use the quadratic
loss. Pass the scores directly, or integer codes together with `values`.

```{r}
kappa(dat.zapf2016, estimator = "pairwise")
```

## Counts-format data

When raters are interchangeable and you only have per-subject category counts,
use `kappa_counts()`. `dat.fleiss1971` is the classic psychiatric-diagnosis
table (30 subjects, six raters, five categories):

```{r}
kappa_counts(dat.fleiss1971, estimator = "pairwise")
```

## Coefficient alpha

`alpha()` estimates Cronbach's coefficient alpha for a subjects-by-items score
battery, again with missing entries allowed. `psych::bfi` is a 25-item Big Five
questionnaire answered by 2800 people, many of whom skipped at least one item.
Treating the five-item Neuroticism scale as the battery, the normal-theory FIML
estimator handles the missing responses under ignorable missingness:

```{r}
#| eval: !expr requireNamespace("psych", quietly = TRUE)
data(bfi, package = "psych")
neuroticism <- as.matrix(bfi[, paste0("N", 1:5)])
sum(!complete.cases(neuroticism))      # respondents missing at least one item
alpha(neuroticism, estimator = "nt_fiml")
```

## Beyond pairwise: g-wise and vector-valued agreement

The same `kappa()` reaches two generalizations. Set `g > 2` for *g-wise*
agreement, where the disagreement kernel compares `g` raters at once (`g = 2` is
ordinary pairwise kappa):

```{r}
kappa(dat.gwet2014, estimator = "ipw", weight = "nominal", g = 3)
```

And when each rating is a *vector*, pass a subjects-by-raters-by-features array.
`dat.vanbelle2019` holds 28 observers classifying crackles at six chest sites
for 20 patients. A component loss (`weight`) measures vector disagreement:

```{r}
kappa(dat.vanbelle2019[, 1:4, ], estimator = "pairwise")
```

The [g-wise agreement](articles/gwise-agreement.html) article develops the
multirater kernels in full. Vector-valued agreement is **frontier / experimental**
ongoing research — see the
[vector-valued agreement](articles/vector-valued-agreement.html) article.

## Testing whether coefficients are equal

A common question is whether a coefficient is the *same* across two groups, two
estimators, two timepoints, or several rater pairs. `kappa_test()` and
`alpha_test()` answer it and return a standard `htest`. Use `paired = TRUE`
(the default) when the fits share the same subjects (the dependence is taken
from the per-subject influence functions) and `paired = FALSE` for
independent samples.

```{r}
kappa_test(AB = kappa(dat.zapf2016[, 1:2], estimator = "pairwise"),
           CD = kappa(dat.zapf2016[, 3:4], estimator = "pairwise"),
           coef = "Conger")
```

See the [equality tests](articles/equality-tests.html) article for the
one-sample, independent, paired, and *G*-way homogeneity cases worked out on
real data, and the [mathematical guides](articles/index.html) for the
loss-matrix formulation and the missing-data estimators.
