\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{What \texttt{method = "gwet"} Means in \texttt{misskappa}}
\date{}

\begin{document}
\maketitle

This note documents what the \texttt{"gwet"} method does in the current
\texttt{misskappa} implementation (as opposed to the descriptive label in the
manuscript).

\section*{Setup}

Let $X_{ij}$ be the rating for subject $i \in \{1,\dots,n\}$ by rater
$j \in \{1,\dots,R\}$, and let $M_{ij} = 1[X_{ij}\ \text{is observed}]$ indicate
non-missingness.

Define the estimated marginal and pairwise observation rates
\[
  \hat\pi_j = \frac{1}{n}\sum_{i=1}^n M_{ij},
  \qquad
  \hat\pi_{jk} = \frac{1}{n}\sum_{i=1}^n M_{ij}M_{ik}.
\]
Let $\ell(\cdot,\cdot)$ be the (possibly weighted) loss used to define
disagreement.

\section*{Where \texttt{"gwet"} differs from \texttt{"available"}}

In \texttt{misskappa}, the methods \texttt{"available"}, \texttt{"ipw"}, and
\texttt{"gwet"} differ only through the weights used in the nonparametric
estimators (raw categorical and continuous backends).

\subsection*{Observed disagreement $\hat d$ (within-subject pairs)}

The observed disagreement estimate is computed from within-subject rater pairs:
\[
  \hat d
  \;=\;
  \frac{
    \sum_{i=1}^n \sum_{1\le j<k\le R} M_{ij}M_{ik}\,\ell(X_{ij},X_{ik})\,w^{(\mathrm{obs})}_{jk}
  }{
    \sum_{i=1}^n \sum_{1\le j<k\le R} M_{ij}M_{ik}\,w^{(\mathrm{obs})}_{jk}
  }.
\]
The weight choice is:
\[
  w^{(\mathrm{obs})}_{jk} =
  \begin{cases}
    1 & \texttt{method = "available"} \\
    \hat\pi_{jk}^{-1} & \texttt{method = "ipw"} \\
    1 & \texttt{method = "gwet"}.
  \end{cases}
\]
So \texttt{"gwet"} and \texttt{"available"} use the same observed-disagreement
component in the current code.

\subsection*{Chance disagreement $\hat d_C, \hat d_F$ (cross-subject pairs)}

The chance-disagreement components (used for Conger/Fleiss kappas) are based on
cross-subject pairings and use product weights:
\[
  w^{(\mathrm{exp})}_{jk} =
  \begin{cases}
    1 & \texttt{method = "available"} \\
    \hat\pi_j^{-1}\hat\pi_k^{-1} & \texttt{method = "ipw"} \\
    \hat\pi_j^{-1}\hat\pi_k^{-1} & \texttt{method = "gwet"}.
  \end{cases}
\]
Thus, in \texttt{misskappa}, \texttt{"gwet"} is \emph{not} the same as
\texttt{"available"}: it leaves $\hat d$ unchanged but reweights the chance term
using inverse \emph{marginal} observation rates (and does \emph{not} use inverse
pairwise observation rates in $\hat d$ as \texttt{"ipw"} does).

\section*{Practical consequences}

\begin{itemize}
\item With complete data ($M_{ij}\equiv 1$), $\hat\pi_j \equiv 1$ and
  \texttt{"available"} and \texttt{"gwet"} coincide.
\item If missingness rates differ by rater (some columns have more NAs), then
  \texttt{"gwet"} differs from \texttt{"available"} through the chance term.
\item For counts data (\texttt{kappa\_counts()}), \texttt{"gwet"} is not
  supported because the counts format does not retain per-rater missingness.
\end{itemize}

\end{document}

