---
title: "Experiment 11: Quadratic Delta Variance"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/archive/pre-redesign/11-quadratic-delta-variance/run_experiment.R --smoke",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

method_order <- c(
  "basic_if_n", "basic_if_unbiased", "basic_delta2_var", "basic_delta2_center_var",
  "arcsine_if_n", "arcsine_if_unbiased", "arcsine_delta2_var", "arcsine_delta2_center_var",
  "fisher_if_n", "fisher_if_unbiased", "fisher_delta2_var", "fisher_delta2_center_var"
)
summary$method <- factor(summary$method, levels = method_order)
focus <- summary[summary$coefficient %in% c("Conger", "Fleiss"), ]
focus <- focus[order(focus$target, focus$coefficient, focus$R, focus$n, focus$method), ]

overall <- aggregate(
  cbind(coverage, miss_below, miss_above, mean_length) ~ target + method,
  data = focus,
  FUN = mean
)
overall <- overall[order(overall$target, overall$method), ]
overall_value <- function(target, method, column) {
  overall[overall$target == target & as.character(overall$method) == method, column][[1]]
}
overall_tab <- data.frame(
  Target = overall$target,
  Method = overall$method,
  Coverage = fmt(overall$coverage),
  `Miss Below` = fmt(overall$miss_below),
  `Miss Above` = fmt(overall$miss_above),
  `Mean Length` = fmt(overall$mean_length),
  check.names = FALSE
)

variance_diag <- aggregate(
  cbind(mean_variance_ratio_unbiased, mean_variance_ratio_delta2, mean_bias_delta2, mean_se_if_n) ~ target,
  data = focus[focus$method == "basic_if_n", ],
  FUN = mean
)
variance_diag <- variance_diag[order(variance_diag$target), ]
variance_tab <- data.frame(
  Target = variance_diag$target,
  `n-1 Var Ratio` = fmt(variance_diag$mean_variance_ratio_unbiased),
  `Delta2 Var Ratio` = fmt(variance_diag$mean_variance_ratio_delta2),
  `Delta2 Bias` = fmt(variance_diag$mean_bias_delta2, 4),
  `Mean IF SE` = fmt(variance_diag$mean_se_if_n),
  check.names = FALSE
)

worst <- focus[order(focus$coverage), ][1, ]
worst_tab <- head(focus[order(focus$coverage), ], 14)
worst_tab <- data.frame(
  Target = worst_tab$target,
  Coefficient = worst_tab$coefficient,
  R = worst_tab$R,
  n = worst_tab$n,
  Method = worst_tab$method,
  Coverage = fmt(worst_tab$coverage),
  `Miss Below` = fmt(worst_tab$miss_below),
  `Miss Above` = fmt(worst_tab$miss_above),
  `Mean Length` = fmt(worst_tab$mean_length),
  check.names = FALSE
)

best_by_cell <- do.call(rbind, lapply(split(
  focus,
  interaction(focus$target, focus$coefficient, focus$R, focus$n)
), function(d) {
  d[which.min(abs(d$coverage - 0.95)), ]
}))
best_tab <- data.frame(
  Target = best_by_cell$target,
  Coefficient = best_by_cell$coefficient,
  R = best_by_cell$R,
  n = best_by_cell$n,
  Method = best_by_cell$method,
  Coverage = fmt(best_by_cell$coverage),
  `Mean Length` = fmt(best_by_cell$mean_length),
  check.names = FALSE
)
```

## Question

Can finite-sample analytic corrections to the influence-function variance improve coverage for complete categorical quadratic kappa?

## Short Answer

This directly probes the variance idea, without bootstrap. The `if_unbiased` correction is just the empirical influence covariance with denominator `n - 1`. The `delta2_var` correction adds the second-order normal delta variance term for the ratio map `theta = 1 - A / B`. The `delta2_center_var` correction also subtracts the corresponding second-order delta bias estimate from the interval center.

The short version: variance inflation helps a little, but it is small. The `n - 1` correction mostly behaves like a mild interval-length bump. The second-order delta variance is also modest in this Perreault-Leigh grid: on the Fisher scale, average coverage moves from `r fmt(overall_value(0.9, "fisher_if_n", "coverage"))` to `r fmt(overall_value(0.9, "fisher_delta2_center_var", "coverage"))` at target `0.9`, and from `r fmt(overall_value(0.95, "fisher_if_n", "coverage"))` to `r fmt(overall_value(0.95, "fisher_delta2_center_var", "coverage"))` at target `0.95`. The bias correction is not a clear win, which suggests the worst upper-boundary failures are not mainly caused by a missing smooth-ratio bias term. The worst coverage is `r fmt(worst$coverage)` for target `r worst$target`, `r worst$coefficient`, `R = r worst$R`, `n = r worst$n`, method `r worst$method`.

## Evidence

`Miss Below` means the interval upper endpoint is below the true coefficient; `Miss Above` means the lower endpoint is above it.

Average over coefficients, rater counts, and sample sizes:

```{r}
knitr::kable(overall_tab, align = "rlrrrr")
```

Average size of the variance and bias corrections:

```{r}
knitr::kable(variance_tab, align = "rrrrr")
```

Worst cells:

```{r}
knitr::kable(worst_tab, align = "rlrrlrrrr")
```

Closest-to-nominal method by cell:

```{r}
knitr::kable(best_tab, align = "rlrrlrr")
```

## Methods

The data-generating model is the complete categorical Perreault-Leigh setup used by Moss (2024): five uniform categories, complete rectangular ratings, and knowledge probability `sqrt(target)`, so the population quadratic coefficient equals `target`.

Write the coefficient as `theta = 1 - A / B`, where `A` is the observed disagreement and `B` is the chance disagreement. The baseline influence variance uses the first derivative of this ratio map and the empirical covariance of the estimated influence functions for `(A, B)`.

The second-order variance uses the normal delta approximation

```text
Var[g(T)] ~= grad' Sigma grad / n + trace(H Sigma H Sigma) / (2 n^2),
```

where `g(A, B) = 1 - A / B`, `grad` is its gradient, `H` is its Hessian, and `Sigma` is estimated from the item-level influence values using denominator `n - 1`. The second-order bias estimate is `trace(H Sigma) / (2 n)`.

Each correction is applied on three scales: identity/basic, Moss's arcsine transform, and Moss's Fisher transform.

## Caveats

This is still a smooth approximation to a discrete bounded statistic. It does not include exact finite-sample U-statistic covariance terms, category-count discreteness, or a lattice/boundary correction. Those may be the next places to look if the smooth delta terms remain too small.

## Reproduce

Smoke run:

```sh
Rscript experiments/archive/pre-redesign/11-quadratic-delta-variance/run_experiment.R --smoke
```

Exploratory run used here:

```sh
Rscript experiments/archive/pre-redesign/11-quadratic-delta-variance/run_experiment.R --reps 500 --target-grid 0.5,0.8,0.9,0.95 --n-grid 10,20,40 --r-grid 2,5 --progress
```

This report used n grid `r meta[["n_grid"]]`, R grid `r meta[["r_grid"]]`, reps `r meta[["reps"]]`, target grid `r meta[["target_grid"]]`, and seed base `r meta[["seed_base"]]`.
