---
title: "Study 33: nt-FIML Degenerate Sigma Rescue"
format:
  html:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
results_dir <- Sys.getenv("STUDY33_RESULTS_DIR", "results/smoke")
needed <- file.path(results_dir, c("recovery.csv", "timing.csv", "calibration.csv"))
if (!all(file.exists(needed))) {
  stop(
    "Missing result CSVs. Run: Rscript run_experiment.R --smoke ",
    "or set STUDY33_RESULTS_DIR to a completed output directory.",
    call. = FALSE
  )
}
recovery <- read.csv(file.path(results_dir, "recovery.csv"))
timing <- read.csv(file.path(results_dir, "timing.csv"))
calibration <- read.csv(file.path(results_dir, "calibration.csv"))
```

## Question

Does the experimental `em_options$soo_rcond` rescue for singular observed-block
covariances in normal-theory FIML buy trustworthy inference for small-sample
quadratic kappa?

## Short Answer

No. The rescue can turn numerical failures into point estimates, but the
degenerate rescued stratum reports standard errors that are much too small.
The default policy should remain to error on the singular covariance path.

## Evidence

Recovery:

```{r}
knitr::kable(recovery, digits = 3)
```

Median fit timings:

```{r}
knitr::kable(timing, digits = 3)
```

Calibration:

```{r}
knitr::kable(calibration, digits = 3)
```

## Caveats

This study uses one latent-normal rater model with five raters, five categories,
and one correlation level. It is an engineering and inference diagnostic for the
rescued boundary path, not a broad simulation study for the quadratic paper.

## Reproduce

```bash
Rscript run_experiment.R --smoke --load-all
Rscript run_experiment.R --full --load-all
```
