---
title: "Experiment 14: MCAR verification under the Perreault–Leigh DGP"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop("Missing results. Run: Rscript experiments/14-quadratic-mcar-verification/run_experiment.R",
       call. = FALSE)
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
```

## Question

Under the Perreault–Leigh judge-skill DGP with MCAR missingness, does
the closed-form quadratic moment estimator
(`misskappa::kappa_quadratic`) deliver negligible bias and approximately
nominal Wald coverage for Fleiss and Conger across a small
$(n, R)$ grid? How does empirical $\widehat{\Gamma}$ compare to normal
$\Gamma_{N}$? And how does the pairwise-available estimator compare to
listwise deletion in RMSE?

## Short answer

- **Verification.** Bias is small at $n \ge 50$ and substantial at
  $n = 10$. Coverage under empirical $\widehat{\Gamma}$ reaches the
  nominal $0.95$ by $n = 200$ and undercovers at $n = 10$. Normal
  $\Gamma_{N}$ slightly undercovers at large $n$ and fails to fit on
  many small-$n$ draws.
- **Listwise efficiency.** At $R = 4$ the pairwise-available estimator
  has visibly lower RMSE (ratio ~0.75). At $R = 2$ the two estimators
  are roughly comparable (ratio ~1.07): pairwise-available picks up
  singleton-observed rows for the rater means but not for the
  covariance, and that trade does not always pay off at small $n$.

## Evidence

```{r}
focus <- summary[, c("estimator", "vcov_kind", "coefficient", "R", "n",
                     "p_pair", "truth", "reps", "mean_n_used",
                     "bias", "se_over_sd", "cov95", "rmse")]
focus <- focus[order(focus$estimator, focus$vcov_kind, focus$coefficient,
                     focus$R, focus$n, focus$p_pair), ]
focus$bias <- fmt(focus$bias)
focus$se_over_sd <- fmt(focus$se_over_sd, 2)
focus$cov95 <- fmt(focus$cov95)
focus$rmse <- fmt(focus$rmse)
focus$truth <- fmt(focus$truth)
focus$mean_n_used <- fmt(focus$mean_n_used, 1)
knitr::kable(focus, row.names = FALSE,
             col.names = c("estimator", "vcov", "coef", "R", "n", "p_pair",
                           "truth", "reps", "n_used",
                           "bias", "SE/SD", "cov95", "RMSE"),
             align = "llllrrrrrrrr")
```

## Caveats

- Single DGP (Perreault–Leigh, $p_{k} = 0.9$, $C = \{-2,\dots,2\}$).
- Single pairwise observation rate ($0.6$).
- Wald intervals only. Fisher and arcsine transforms from
  \citet{Moss2024-cc} improve small-$n$ coverage but are not
  exercised.
- Brennan–Prediger is reported by the package but not displayed.
- Normal $\Gamma_{N}$ returns a non-finite standard error on draws
  where the empirical covariance is ill-conditioned (the Wald
  quadratic form picks up a small negative diagonal entry). At
  $n = 10$ a large fraction of draws are affected, so the
  normal-$\Gamma$ coverage at small $n$ is reported on a selected
  subset and should be read with caution.

## Reproduce

```bash
Rscript experiments/14-quadratic-mcar-verification/run_experiment.R --progress
```

Smoke variant:

```bash
Rscript experiments/14-quadratic-mcar-verification/run_experiment.R --smoke
```

```{r}
cat(sprintf("Metadata: dgp=%s; p_k=%s; n_grid=%s; r_grid=%s; miss=%s; reps=%s; seed_base=%s; misskappa=%s\n",
            meta[["dgp"]], meta[["knowledge_prob"]],
            meta[["n_grid"]], meta[["r_grid"]], meta[["miss_grid"]],
            meta[["reps"]], meta[["seed_base"]], meta[["misskappa_version"]]))
```
