---
title: "Experiment 14: MCAR verification for the closed-form quadratic kappa"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
input <- knitr::current_input()
base_dir <- if (!is.null(input)) dirname(normalizePath(input)) else getwd()
summary_path <- file.path(base_dir, "results", "summary.csv")
metadata_path <- file.path(base_dir, "results", "metadata.csv")

if (!file.exists(summary_path) || !file.exists(metadata_path)) {
  stop(
    "Missing experiment results. Run: Rscript experiments/14-quadratic-mcar-verification/run_experiment.R",
    call. = FALSE
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)

fmt <- function(x, digits = 3) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
```

## Question

Does the closed-form quadratic moment estimator
(`misskappa::kappa_quadratic`) deliver negligible bias and approximately
nominal Wald coverage for the Fleiss and Conger coefficients under
MCAR, across a small $(n, R, \text{pairwise rate})$ grid? And does the
normal-theory $\Gamma_N$ option track the empirical $\hat\Gamma$ when
the data-generating process is multivariate normal?

## Short answer

Yes on both counts.

- Bias is small (around $-0.02$ to $-0.03$ at $n = 50$, below $0.01$ at
  $n \ge 200$) and shrinks at the expected $1/n$ rate.
- $\widehat{SE}/\mathrm{SD}$ sits near 1 at $n \ge 200$ for both
  $\Gamma$ variants and shows mild under-estimation at $n = 50$ for
  empirical $\hat\Gamma$.
- 95% Wald coverage is at nominal at $n = 200$ and above; mild
  undercoverage at $n = 50$ (between 0.89 and 0.94 in the harder
  cells).
- Normal $\Gamma_N$ buys slightly better small-$n$ coverage on a
  multivariate-normal DGP but agrees with empirical $\hat\Gamma$
  asymptotically.

## Evidence

```{r}
focus <- summary[, c("dgp", "vcov_kind", "coefficient", "R", "n",
                     "p_pair", "truth", "bias", "se_over_sd", "cov95",
                     "reps")]
focus <- focus[order(focus$dgp, focus$coefficient, focus$vcov_kind,
                     focus$R, focus$n, focus$p_pair), ]
focus$bias <- fmt(focus$bias)
focus$se_over_sd <- fmt(focus$se_over_sd, 2)
focus$cov95 <- fmt(focus$cov95)
focus$truth <- fmt(focus$truth)
knitr::kable(focus, row.names = FALSE,
             col.names = c("DGP", "vcov", "coef", "R", "n", "p_pair",
                           "truth", "bias", "SE/SD", "cov95", "reps"),
             align = "lllrrrrrrrr")
```

## Caveats

- Multivariate normal DGP only. Heavy-tailed or skewed marginals are
  out of scope; pseudo-elliptical $\Gamma$ would be the right vehicle
  and is left to a follow-up.
- Wald intervals only. Transform-based intervals (Fisher, arcsine) are
  available in `misskappa` but not exercised here.
- Brennan--Prediger is reported by the package but not displayed: under
  quadratic loss it is an affine rescaling of the same observed
  disagreement that drives Fleiss, so its bias and coverage track
  Fleiss closely.

## Reproduce

```bash
Rscript experiments/14-quadratic-mcar-verification/run_experiment.R --progress
```

Smoke variant:

```bash
Rscript experiments/14-quadratic-mcar-verification/run_experiment.R --smoke
```

```{r}
cat(sprintf("Metadata: n_grid=%s; r_grid=%s; miss_grid=%s; reps=%s; seed_base=%s; misskappa=%s\n",
            meta[["n_grid"]], meta[["r_grid"]], meta[["miss_grid"]],
            meta[["reps"]], meta[["seed_base"]], meta[["misskappa_version"]]))
```
