---
title: "Validation of normal-FIML quadratic kappa (`kappa_continuous(method = \"fiml\")`)"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
val_path <- file.path(res_dir, "validation.csv")
if (!file.exists(val_path)) {
  stop("No results. Run `Rscript run_experiment.R` first.")
}
validation <- read.csv(val_path, stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)
```

## Question

Does the normal-FIML quadratic kappa --- the Conger and Fleiss coefficients read
off a saturated multivariate-normal covariance fitted by FIML/EM under ignorable
missingness, with a delta-method standard error --- agree with the established
references and the bootstrap truth?

This is the kappa counterpart of experiment 17 (alpha). The estimator reuses the
same saturated-EM covariance, casewise scores and sandwich machinery as
`alpha_continuous()`; the only new piece is the coefficient gradient, which
(unlike alpha) carries a **mean block** because the kappas depend on
`t3 = sum (mu_j - mubar)^2`. The DGP gives the raters distinct means so `t3 > 0`
and Conger differs from Fleiss, exercising that block. The EM covariance is
checked against **magmaan**'s `estimate_saturated_em_moments()`, and the sandwich
SE against the same routine's exposed sandwich `acov = H^-1 J H^-1` --- so the
influence-function variance is confirmed through magmaan's independent FIML path
without any change to magmaan. The point and SE are also checked against the
pairwise-available estimator `kappa_continuous(method = "available")` and a
nonparametric **case bootstrap** (the ground truth). The `nonnormal/mcar` cell is
the discriminator.

## Short answer

Yes. On complete data the sandwich SE equals the pairwise-available ADF SE to
numerical precision, and the EM covariance matches magmaan to machine epsilon.
Under non-normality the **normal-theory** SE undershoots while the **sandwich**
SE tracks the bootstrap --- the robustness payoff, and the reason the estimator
defaults to `se_type = "sandwich"`. Under MCAR the FIML sandwich SE is marginally
tighter than the pairwise-available SE, the expected small efficiency gain.

## Moments

Maximum absolute difference between the EM covariance and magmaan's oracle, per
cell (Conger and Fleiss share the fit, so one row per cell):

```{r moments}
m <- validation[validation$coef == "Conger",
                c("dist", "mech", "max_cov_err_magmaan")]
knitr::kable(m, digits = 9, row.names = FALSE,
             col.names = c("Dist", "Mech", "Max cov. error vs magmaan"))
```

The covariance matches magmaan to roughly `1e-6` under missingness and to machine
epsilon on complete data.

## Standard errors

```{r ses}
v <- validation[, c("dist", "mech", "coef", "est", "se_sandwich", "se_normal",
                    "se_available", "se_magmaan", "se_boot")]
knitr::kable(v, digits = 5, row.names = FALSE,
             col.names = c("Dist", "Mech", "Coef", "Estimate", "SE sandwich",
                           "SE normal", "SE available", "SE magmaan",
                           "SE bootstrap"))
```

Reading the table:

- On **complete** data the sandwich SE reproduces the pairwise-available ADF SE
  (`SE available`) to four digits --- the two are the same delta-method estimator
  when no row is incomplete, an independent confirmation through a different code
  path (the C++ closed form).
- The **`SE magmaan`** column is the saturated-moment sandwich SE obtained from
  magmaan's exposed `acov = H^-1 J H^-1`, contracted with the same gradient.
  It reproduces the misskappa **`SE sandwich`** column to numerical precision in
  every cell, an independent confirmation of the influence-function variance
  through magmaan's separate FIML implementation.
- Under **normal MCAR/MAR** the sandwich, normal-theory and available SEs are all
  on the same empirical scale as the bootstrap, with the FIML sandwich SE
  slightly **tighter** than the pairwise-available SE.
- In the **`nonnormal/mcar`** cell the normal-theory delta SE undershoots the
  sandwich and bootstrap scale, while the sandwich SE stays close to the
  bootstrap. This is the expected behaviour and the reason the estimator defaults
  to `se_type = "sandwich"`; `se_type = "normal"` is exposed for the
  multivariate-normal regime.

## Caveats

This is a single-realization validation per cell with a bootstrap truth, not a
coverage study: it confirms the SE *scale* and the moment/reference agreement,
not Wald interval coverage over repeated samples. The bootstrap in the checked-in
smoke run (`--smoke`) is a gross-scale sanity check; paper-facing coverage and
efficiency claims should come from a larger Monte Carlo. Only the quadratic kernel
is in scope --- the Conger/Fleiss kappas are functions of `(mu, Sigma)` only under
quadratic weighting. Exchangeable-count and categorical-FIML variants are not
covered here.

## Reproduce

```bash
Rscript run_experiment.R --smoke     # cheap: n=400, boot=100
Rscript run_experiment.R             # default: n=800, boot=600
```

## Provenance

```{r meta}
knitr::kable(metadata, col.names = c("key", "value"))
```
