---
title: "Experiment 20: Normal-FIML Stress Check for Ordinal MAR Alpha"
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
res_dir <- "results"
if (!dir.exists(res_dir)) {
  res_dir <- file.path("experiments", "probes", "20-alpha-nt-fiml-stress", "results")
}
needed <- c("truth.csv", "summary.csv", "metadata.csv")
missing <- needed[!file.exists(file.path(res_dir, needed))]
if (length(missing)) {
  stop(
    "Missing results: ", paste(missing, collapse = ", "),
    ". Run: Rscript experiments/probes/20-alpha-nt-fiml-stress/run_experiment.R --smoke --progress"
  )
}
truth <- read.csv(file.path(res_dir, "truth.csv"), stringsAsFactors = FALSE)
summary <- read.csv(file.path(res_dir, "summary.csv"), stringsAsFactors = FALSE)
metadata <- read.csv(file.path(res_dir, "metadata.csv"), stringsAsFactors = FALSE)

fmt <- function(x, digits = 4) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}
label_method <- function(x) {
  out <- x
  out[out == "pairwise"] <- "Pairwise"
  out[out == "cat_fiml"] <- "Cat-FIML"
  out[out == "normal_fiml"] <- "Normal FIML"
  out[out == "listwise"] <- "Listwise"
  out
}
label_mechanism <- function(x) {
  out <- x
  out[out == "mcar30"] <- "MCAR 30%"
  out[out == "mar_linear30"] <- "Linear Anchor MAR 30%"
  out[out == "mar_lowgate45"] <- "Low-Anchor MAR 45%"
  out
}
```

## Question

Can the surprisingly small bias of normal-theory FIML in the paper-facing
ordinal simulation be broken by changing only the ordinal margins, latent
response distribution, and MAR shape?

The control condition is the current thresholded congeneric Gaussian copula with
balanced three-category margins. The stress conditions keep six scored ordinal
items but use skewed margins, replace the Gaussian copula with a two-class
ordinal mixture, and add an anchor-group mixture in which item 1 indexes groups
with different target-item covariance. Missingness is still MAR because all
target-item response probabilities depend only on item 1, which is always
observed.

## Short Answer

```{r short-answer}
normal <- summary[summary$method == "normal_fiml", ]
mar <- normal[normal$mechanism != "mcar30", ]
worst <- mar[which.max(abs(mar$bias)), ]
reps <- metadata$value[metadata$key == "reps"]
n <- metadata$value[metadata$key == "n"]
```

This run uses `r reps` replications per cell at `n = r n`. The largest absolute
normal-FIML MAR bias is `r fmt(worst$bias, 3)` in the
`r worst$dgp_label` / `r label_mechanism(worst$mechanism)` cell. Treat this as a
stress diagnostic, not final paper evidence: the useful question is whether the
direction and size persist when we later increase replications or move to a
larger population-style run.

## Truths

```{r truths}
truth_tab <- truth[, c("dgp", "family", "truth_alpha", "min_score_cov",
                       "max_score_cov", "dgp_note")]
truth_tab$truth_alpha <- fmt(truth_tab$truth_alpha)
truth_tab$min_score_cov <- fmt(truth_tab$min_score_cov)
truth_tab$max_score_cov <- fmt(truth_tab$max_score_cov)
knitr::kable(
  truth_tab,
  col.names = c("DGP", "Family", "Alpha", "Min Score Cov",
                "Max Score Cov", "DGP Note")
)
```

## Evidence

```{r evidence-table}
display <- summary[, c("dgp", "mechanism", "method", "n_valid", "truth",
                       "bias", "mc_se_bias", "mse", "coverage95",
                       "mean_complete_rows")]
display$mechanism <- label_mechanism(display$mechanism)
display$method <- label_method(display$method)
display$truth <- fmt(display$truth)
display$bias <- fmt(display$bias)
display$mc_se_bias <- fmt(display$mc_se_bias)
display$mse <- fmt(display$mse)
display$coverage95 <- fmt(display$coverage95, 3)
display$mean_complete_rows <- fmt(display$mean_complete_rows, 1)
knitr::kable(
  display,
  col.names = c("DGP", "Mechanism", "Method", "Valid", "Truth",
                "Bias", "MC SE Bias", "MSE", "Coverage", "Complete Rows")
)
```

```{r normal-focus}
normal_tab <- summary[summary$method == "normal_fiml",
                      c("dgp", "mechanism", "bias", "mc_se_bias", "mse",
                        "coverage95", "mean_ci_length")]
normal_tab$mechanism <- label_mechanism(normal_tab$mechanism)
normal_tab$bias <- fmt(normal_tab$bias)
normal_tab$mc_se_bias <- fmt(normal_tab$mc_se_bias)
normal_tab$mse <- fmt(normal_tab$mse)
normal_tab$coverage95 <- fmt(normal_tab$coverage95, 3)
normal_tab$mean_ci_length <- fmt(normal_tab$mean_ci_length)
knitr::kable(
  normal_tab,
  col.names = c("DGP", "Mechanism", "Bias", "MC SE Bias", "MSE",
                "Coverage", "CI Length")
)
```

```{r bias-plot, fig.width=7, fig.height=4.5}
old_par <- par(no.readonly = TRUE)
on.exit(par(old_par), add = TRUE)
plot_df <- summary
plot_df$cell <- paste(plot_df$dgp, label_mechanism(plot_df$mechanism), sep = "\n")
cols <- c(pairwise = "#000000", cat_fiml = "#009E73",
          normal_fiml = "#0072B2", listwise = "#D55E00")
methods <- intersect(names(cols), unique(plot_df$method))
cells <- unique(plot_df$cell)
x <- seq_along(cells)
ylim <- range(plot_df$bias, na.rm = TRUE)
plot(x, rep(NA_real_, length(x)), ylim = ylim, xaxt = "n",
     xlab = "", ylab = "Bias", main = "Bias by DGP and Missingness")
abline(h = 0, col = "gray70", lty = 2)
offsets <- seq(-0.20, 0.20, length.out = length(methods))
for (m_idx in seq_along(methods)) {
  m <- methods[m_idx]
  g <- plot_df[plot_df$method == m, ]
  points(match(g$cell, cells) + offsets[m_idx], g$bias,
         pch = 16, col = cols[m])
}
axis(1, at = x, labels = cells, las = 2, cex.axis = 0.65)
legend("topright", legend = label_method(methods), pch = 16,
       col = cols[methods], bty = "n")
```

## Caveats

- This is deliberately a stress check. The low-anchor MAR mechanism is plausible
  as an early-item dropout proxy, but it is stronger and less smooth than the
  paper's current MAR design.
- Cat-FIML is saturated over the observed ordinal response space. At `n = 100`,
  it can still have finite-sample bias from sparse cells even when its model is
  correctly specified.
- Normal FIML uses the scored categories as continuous and reports sandwich
  intervals. The sandwich does not remove pseudo-model bias.
- The latent-class DGP is not calibrated to match the Gaussian-copula covariance
  exactly. The anchor-group DGP is even more diagnostic because it makes the
  conditional covariance of the target items depend on the observed anchor. These
  are candidate additions, not replacement paper DGPs yet.

## Reproduce

Smoke:

```sh
Rscript experiments/probes/20-alpha-nt-fiml-stress/run_experiment.R --smoke --progress
quarto render experiments/probes/20-alpha-nt-fiml-stress/report.qmd
```

Configured diagnostic:

```sh
Rscript experiments/probes/20-alpha-nt-fiml-stress/run_experiment.R \
  --reps 200 \
  --n 100 \
  --progress
quarto render experiments/probes/20-alpha-nt-fiml-stress/report.qmd
```

```{r metadata}
knitr::kable(metadata, col.names = c("Key", "Value"))
```
