---
title: "Kappa Planned-Missingness Smoke"
subtitle: "Experiment 29"
bibliography: report.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
pick_results_dir <- function() {
  override <- Sys.getenv("KAPPA29_RESULTS_DIR", unset = "")
  if (nzchar(override)) return(override)
  result_dirs <- list.dirs("results", full.names = TRUE, recursive = FALSE)
  if (length(result_dirs)) {
    info <- file.info(file.path(result_dirs, "summary.csv"))
    result_dirs <- result_dirs[!is.na(info$mtime)]
    if (length(result_dirs)) return(result_dirs[which.max(info$mtime)])
  }
  "results"
}

results_dir <- pick_results_dir()
summary_path <- file.path(results_dir, "summary.csv")
metadata_path <- file.path(results_dir, "metadata.csv")
dgps_path <- file.path(results_dir, "dgps.csv")
mechanisms_path <- file.path(results_dir, "mechanisms.csv")

if (!file.exists(summary_path)) {
  stop(
    "No summary.csv found. Run, for example:\n",
    "  Rscript run_experiment.R --smoke --progress\n",
    "or set KAPPA29_RESULTS_DIR to a result directory before rendering."
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
dgps <- read.csv(dgps_path, stringsAsFactors = FALSE)
mechanisms <- read.csv(mechanisms_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)
meta_value <- function(key, default = "") {
  val <- meta[[key]]
  if (is.null(val) || length(val) == 0L || is.na(val)) default else val
}
fmt <- function(x, digits = 3) ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
fmt_signed <- function(x, digits = 3) ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
fmt_int <- function(x) formatC(x, format = "d", big.mark = ",")

method_label <- c(
  ipw_nominal = "IPW",
  cat_fiml_nominal = "Cat-FIML",
  listwise_nominal = "Listwise",
  ipw_absolute = "IPW",
  cat_fiml_absolute = "Cat-FIML",
  listwise_absolute = "Listwise",
  ipw_quadratic = "IPW",
  cat_fiml_quadratic = "Cat-FIML",
  pairwise_quadratic = "Pairwise",
  nt_fiml_quadratic = "NT-FIML",
  listwise_quadratic = "Listwise"
)
summary$method_label <- unname(method_label[summary$method])
summary$method_label[is.na(summary$method_label)] <- summary$method[is.na(summary$method_label)]
summary$abs_bias <- abs(summary$bias)
```

## Question

This study is the first paper-facing screen for kappa under accidental MCAR,
clean observed MAR, and planned missing ratings. It adds small sample sizes,
planned rater-subset designs, listwise deletion, high-agreement Moss-style
cells, ordinal normal-copula ratings, and Dawid-Skene-style severity raters.

The generic planned-missingness framing follows the idea that measurement burden
can be deliberately distributed over subjects while preserving ignorable
observed-data likelihoods [@Graham2006PlannedMissingness].

The design additions here are narrower and more agreement-specific than the
generic planned-missingness literature. The ICC planned-missingness papers treat
ratings as exchangeable and ask how to spend a fixed rating budget across
subjects, raters, and overlap patterns [@vanDerArkJorgensenTenHove2023Planned;
@tenHoveJorgensenVanDerArk2025ICC]. We therefore include a random two-rater
assignment (`designed_random2`) and a block assignment (`designed_block2`) to
separate ordinary loss of information from the fixed-rater identifiability
problem created by disconnected co-observation graphs. Fanshawe et al.'s breast
cancer grading application makes the same graph-connectivity issue visible for
latent-trait agreement models [@Fanshawe2008AgreementMissing], so the runner also
includes a sparse connected ring (`designed_ring2`) where every rater is linked
but many rater pairs are never co-observed.

The kappa-native planned design is the balanced incomplete block idea in
Abraira and Perez de Vargas [@AbrairaPerezDeVargas1999KappaIncomplete]:
`designed_bib3` assigns three raters per subject, and with six raters uses ten
blocks in which every rater pair appears exactly twice. The form and anchor
mechanisms riff on the three-form and anchor-rater designs discussed for
planned incomplete reliability data [@Graham2006PlannedMissingness;
@tenHoveJorgensenVanDerArk2025ICC]. `designed_forms4` cycles four-rater
subsets corresponding to two of three rater-form groups; `designed_anchor3`
keeps rater 1 on every subject and rotates two additional raters; and
`designed_two_phase` adds periodic fully rated calibration subjects to the
anchor design. These are planned and ignorable by construction, unlike the
accidental MCAR/MAR mechanisms in the same grid. Schouten's fixed-versus-varying
observer distinction [@Schouten1986NominalAgreement] and the modern deletion
comparisons by De Raadt et al. and Warrens et al. motivate keeping listwise
deletion and accidental missingness in the same study
[@DeRaadt2019KappaMissing; @Warrens2025WeightedKappaMissing].

## Run

The report is reading `r results_dir`.

```{r run-summary}
rep_path <- file.path(results_dir, "replicates.csv")
replicate_rows <- if (file.exists(rep_path)) nrow(read.csv(rep_path)) else NA_integer_

run_tab <- data.frame(
  Quantity = c(
    "Mode",
    "Replications requested",
    "Replications in shard",
    "Sample size",
    "Truth sample size",
    "DGPs",
    "Missingness mechanisms",
    "Estimator rows",
    "Summary rows",
    "Replicate rows",
    "Failures",
    "Elapsed seconds"
  ),
  Value = c(
    meta_value("mode"),
    meta_value("reps"),
    meta_value("reps_in_shard"),
    meta_value("n_grid"),
    meta_value("truth_n"),
    meta_value("dgps"),
    meta_value("mechanisms"),
    meta_value("methods"),
    fmt_int(nrow(summary)),
    if (is.na(replicate_rows)) "not checked" else fmt_int(replicate_rows),
    fmt_int(sum(summary$failures[summary$interval == "wald_t"])),
    fmt(as.numeric(meta_value("elapsed_seconds", NA)), 2)
  ),
  check.names = FALSE
)
knitr::kable(run_tab, align = "ll")
```

```{r dgp-table}
dgp_tab <- dgps[, c("dgp", "family", "C", "R", "support_size", "category_prob", "details")]
names(dgp_tab) <- c("DGP", "Family", "Categories", "Raters", "Full patterns",
                    "Category probabilities", "Details")
knitr::kable(dgp_tab, align = "llrrrll")
```

```{r mechanism-table}
mech_cols <- intersect(c("mechanism", "family", "description", "source"),
                       names(mechanisms))
mech_tab <- mechanisms[mechanisms$mechanism %in% unique(summary$mechanism),
                       mech_cols]
names(mech_tab) <- c("Mechanism", "Family", "Description", "Source")[seq_along(mech_cols)]
knitr::kable(mech_tab, align = paste(rep("l", length(mech_cols)), collapse = ""))
```

## Smoke Diagnostics

The smoke run is meant to catch mechanical failures and rough timing problems.
Coverage at this replication count is not interpretable.

```{r quadratic-smoke}
quad <- subset(summary, coefficient == "Conger" &
                 weight_label == "quadratic" &
                 interval == "wald_t")
quad <- quad[order(quad$dgp, quad$mechanism, quad$method_label), ]
quad_tab <- data.frame(
  DGP = quad$dgp,
  Mechanism = quad$mechanism,
  Method = quad$method_label,
  Valid = quad$n_valid,
  Failures = quad$failures,
  Bias = fmt_signed(quad$bias, 3),
  RMSE = fmt(quad$rmse, 3),
  `Mean SE` = fmt(quad$mean_se, 3),
  `Mean n_eff` = fmt(quad$mean_n_eff, 1),
  `Fit ms` = fmt(quad$median_elapsed_ms, 0),
  check.names = FALSE
)
knitr::kable(quad_tab, align = "lllrrrrrrr")
```

## Interval Check

```{r interval-check}
int <- subset(summary, coefficient == "Conger" & weight_label == "quadratic")
int <- aggregate(
  cbind(coverage95, mean_ci_length, miss_below, miss_above) ~ interval + method_label,
  int,
  mean,
  na.rm = TRUE
)
int <- int[order(int$method_label, int$interval), ]
int_tab <- data.frame(
  Method = int$method_label,
  Interval = int$interval,
  Coverage = fmt(int$coverage95, 2),
  Length = fmt(int$mean_ci_length, 3),
  `Miss Below` = fmt(int$miss_below, 2),
  `Miss Above` = fmt(int$miss_above, 2),
  check.names = FALSE
)
knitr::kable(int_tab, align = "llrrrr")
```

## Reproduce

```bash
cd experiments/studies/29-kappa-planned-missingness

Rscript run_experiment.R --smoke --reps 5 --progress \
  --out-dir results/smoke-YYYYMMDD-HHMM

KAPPA29_RESULTS_DIR=results/smoke-YYYYMMDD-HHMM quarto render report.qmd
```
