---
title: "Kappa Planned-Missingness Smoke"
subtitle: "Experiment 29"
bibliography: report.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
pick_results_dir <- function() {
  override <- Sys.getenv("KAPPA29_RESULTS_DIR", unset = "")
  if (nzchar(override)) return(override)
  result_dirs <- list.dirs("results", full.names = TRUE, recursive = FALSE)
  if (length(result_dirs)) {
    info <- file.info(file.path(result_dirs, "summary.csv"))
    result_dirs <- result_dirs[!is.na(info$mtime)]
    if (length(result_dirs)) return(result_dirs[which.max(info$mtime)])
  }
  "results"
}

results_dir <- pick_results_dir()
summary_path <- file.path(results_dir, "summary.csv")
metadata_path <- file.path(results_dir, "metadata.csv")
dgps_path <- file.path(results_dir, "dgps.csv")
mechanisms_path <- file.path(results_dir, "mechanisms.csv")

if (!file.exists(summary_path)) {
  stop(
    "No summary.csv found. Run, for example:\n",
    "  Rscript run_experiment.R --smoke --progress\n",
    "or set KAPPA29_RESULTS_DIR to a result directory before rendering."
  )
}

summary <- read.csv(summary_path, stringsAsFactors = FALSE)
metadata <- read.csv(metadata_path, stringsAsFactors = FALSE)
dgps <- read.csv(dgps_path, stringsAsFactors = FALSE)
mechanisms <- read.csv(mechanisms_path, stringsAsFactors = FALSE)
meta <- setNames(metadata$value, metadata$key)
meta_value <- function(key, default = "") {
  val <- meta[[key]]
  if (is.null(val) || length(val) == 0L || is.na(val)) default else val
}
fmt <- function(x, digits = 3) ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
fmt_signed <- function(x, digits = 3) ifelse(is.na(x), "", sprintf(paste0("%+.", digits, "f"), x))
fmt_int <- function(x) formatC(x, format = "d", big.mark = ",")

method_label <- c(
  ipw_nominal = "IPW",
  cat_fiml_nominal = "Cat-FIML",
  listwise_nominal = "Listwise",
  ipw_absolute = "IPW",
  cat_fiml_absolute = "Cat-FIML",
  listwise_absolute = "Listwise",
  ipw_quadratic = "IPW",
  cat_fiml_quadratic = "Cat-FIML",
  pairwise_quadratic = "Pairwise",
  nt_fiml_quadratic = "NT-FIML",
  listwise_quadratic = "Listwise",
  fleiss_counts_nominal = "Fleiss-counts",
  fleiss_counts_quadratic = "Fleiss-counts",
  fleiss_counts_fiml_nominal = "Fleiss-counts-FIML",
  fleiss_counts_fiml_quadratic = "Fleiss-counts-FIML"
)
summary$method_label <- unname(method_label[summary$method])
summary$method_label[is.na(summary$method_label)] <- summary$method[is.na(summary$method_label)]
summary$abs_bias <- abs(summary$bias)
```

## Question

This study is the first paper-facing screen for kappa under accidental MCAR,
clean observed MAR, and planned missing ratings. It adds small sample sizes,
planned rater-subset designs, listwise deletion, high-agreement Moss-style
cells, ordinal normal-copula ratings, and Dawid-Skene-style severity raters.

The generic planned-missingness framing follows the idea that measurement burden
can be deliberately distributed over subjects while preserving ignorable
observed-data likelihoods [@Graham2006PlannedMissingness].

### Where the planned designs come from, and why the estimand splits in two

The planned rater-subset designs are taken from the ICC planned-incomplete-data
literature [@vanDerArkJorgensenTenHove2023Planned;
@tenHoveJorgensenVanDerArk2025ICC; @Fanshawe2008AgreementMissing;
@AbrairaPerezDeVargas1999KappaIncomplete]. The key point for this study is *how*
that literature gets identifiability from sparse data. ten Hove, Jorgensen and
van der Ark [-@tenHoveJorgensenVanDerArk2025ICC] estimate interrater reliability
as an **intraclass correlation** through a **structured** model --- a
random-effects / variance-components ANOVA model, or equivalently a one-factor
SEM --- in which raters are **exchangeable and interchangeable** ($y_{sr} = \mu +
\mu_s + \mu_r + \mu_{sr}$, with the ICC a ratio of variance components). That
structural assumption is exactly what lets them use partially crossed, even
disconnected, designs: the shared subject and rater variance components are
pooled across *all* observed subject--rater combinations, so rater identity is
never needed and the co-observation graph may be sparse. They flag the boundary
themselves --- the common-factor variant has convergence trouble "when there is
little-to-no overlap among raters... minimal information about shared variance."

`misskappa` takes the opposite stance: its weighted kappa is **distribution-free
and fixed-rater** (no variance-components structure, rater identity retained), so
it cannot borrow strength the way an ICC model does. Consequently the *same*
planned designs split into two identifiability regimes:

- **Fixed-rater kappa (Conger / Cohen).** Each disagreement term is a functional
  of a rater *pair's* bivariate law, so the coefficient is identified only if the
  rater co-observation graph is **complete** (every pair co-observed at least
  once). Mere connectivity does not suffice --- the structured-design connectivity
  that rescues an ICC does not rescue a saturated kappa.
- **Exchangeable Fleiss (`kappa_counts`).** Discarding rater identity collapses
  the pairwise terms to a pooled within-subject quantity, and identifiability
  relaxes to **each subject contributing $\ge 2$ ratings**; the graph may be
  disconnected. This is the categorical analogue of ten Hove et al.'s one-way /
  exchangeable ICC, and it is the *only* target estimable on the sparse designs.

This study therefore carries **two estimand tracks**. The fixed-rater estimators
(IPW, Cat-FIML, pairwise, NT-FIML, listwise) target Conger/Fleiss and are valid
on the complete-graph mechanisms; the exchangeable `kappa_counts` track
(`Fleiss-counts`, coefficient `Fleiss_exch`) runs on *every* mechanism, including
the sparse designs. On the three sparse designs the fixed-rater cells double as an
**identifiability demonstration**: the co-observation guard fires and those cells
are recorded as failures, confirming empirically that distribution-free
fixed-rater kappa cannot use what the structured ICC can.

The seven planned designs map onto the two regimes as follows:

| Planned design | Co-observation graph | Fixed-rater track | Exchangeable track |
|---|---|---|---|
| `designed_bib3` (BIB, 3/subject) | complete (every pair $\times 2$) | yes | yes |
| `designed_anchor3` (anchor + 2 rotating) | complete | yes | yes |
| `designed_two_phase` (anchor + calibration) | complete | yes | yes |
| `designed_forms4` (rater-form subsets) | complete | yes | yes |
| `designed_random2` (random pair) | sparse, often incomplete | guard fires | yes |
| `designed_ring2` (adjacent pairs) | connected but pair-incomplete | guard fires | yes |
| `designed_block2` (disjoint teams) | disconnected | guard fires | yes |

The complete-graph BIB idea is Abraira and Perez de Vargas
[-@AbrairaPerezDeVargas1999KappaIncomplete] (`designed_bib3`: three raters per
subject; with six raters, ten blocks in which every rater pair appears twice).
The form and anchor mechanisms riff on the three-form and anchor-rater designs
for planned incomplete reliability data [@Graham2006PlannedMissingness;
@tenHoveJorgensenVanDerArk2025ICC]; `designed_ring2` follows Fanshawe et al.'s
breast-cancer grading application [@Fanshawe2008AgreementMissing], where the same
graph-connectivity issue first becomes visible. These planned mechanisms are
ignorable by construction, unlike the accidental MCAR/MAR mechanisms in the same
grid. Schouten's fixed-versus-varying observer distinction
[@Schouten1986NominalAgreement] and the modern deletion comparisons by De Raadt
et al. and Warrens et al. motivate keeping listwise deletion and accidental
missingness in the same study [@DeRaadt2019KappaMissing;
@Warrens2025WeightedKappaMissing].

## Run

The report is reading `r results_dir`.

```{r run-summary}
rep_path <- file.path(results_dir, "replicates.csv")
replicate_rows <- if (file.exists(rep_path)) nrow(read.csv(rep_path)) else NA_integer_

run_tab <- data.frame(
  Quantity = c(
    "Mode",
    "Replications requested",
    "Replications in shard",
    "Sample size",
    "Truth sample size",
    "DGPs",
    "Missingness mechanisms",
    "Estimator rows",
    "Summary rows",
    "Replicate rows",
    "Failures",
    "Elapsed seconds"
  ),
  Value = c(
    meta_value("mode"),
    meta_value("reps"),
    meta_value("reps_in_shard"),
    meta_value("n_grid"),
    meta_value("truth_n"),
    meta_value("dgps"),
    meta_value("mechanisms"),
    meta_value("methods"),
    fmt_int(nrow(summary)),
    if (is.na(replicate_rows)) "not checked" else fmt_int(replicate_rows),
    fmt_int(sum(summary$failures[summary$interval == "wald_t"])),
    fmt(as.numeric(meta_value("elapsed_seconds", NA)), 2)
  ),
  check.names = FALSE
)
knitr::kable(run_tab, align = "ll")
```

```{r dgp-table}
dgp_tab <- dgps[, c("dgp", "family", "C", "R", "support_size", "category_prob", "details")]
names(dgp_tab) <- c("DGP", "Family", "Categories", "Raters", "Full patterns",
                    "Category probabilities", "Details")
knitr::kable(dgp_tab, align = "llrrrll")
```

```{r mechanism-table}
mech_cols <- intersect(c("mechanism", "family", "description", "source"),
                       names(mechanisms))
mech_tab <- mechanisms[mechanisms$mechanism %in% unique(summary$mechanism),
                       mech_cols]
names(mech_tab) <- c("Mechanism", "Family", "Description", "Source")[seq_along(mech_cols)]
knitr::kable(mech_tab, align = paste(rep("l", length(mech_cols)), collapse = ""))
```

## Smoke Diagnostics

The smoke run is meant to catch mechanical failures and rough timing problems.
Coverage at this replication count is not interpretable.

```{r quadratic-smoke}
quad <- subset(summary, coefficient == "Conger" &
                 weight_label == "quadratic" &
                 interval == "wald_t")
quad <- quad[order(quad$dgp, quad$mechanism, quad$method_label), ]
quad_tab <- data.frame(
  DGP = quad$dgp,
  Mechanism = quad$mechanism,
  Method = quad$method_label,
  Valid = quad$n_valid,
  Failures = quad$failures,
  Bias = fmt_signed(quad$bias, 3),
  RMSE = fmt(quad$rmse, 3),
  `Mean SE` = fmt(quad$mean_se, 3),
  `Mean n_eff` = fmt(quad$mean_n_eff, 1),
  `Fit ms` = fmt(quad$median_elapsed_ms, 0),
  check.names = FALSE
)
knitr::kable(quad_tab, align = "lllrrrrrrr")
```

### Fixed-rater identifiability on sparse designs

The fixed-rater estimators above should show near-total `Failures` on the
disconnected / pair-incomplete designs (`designed_random2`, `designed_block2`,
`designed_ring2`): the co-observation guard fires because Conger/Fleiss are not
identified there. The exchangeable `Fleiss-counts` track below should instead
produce valid estimates on those same designs.

```{r exch-smoke}
exch <- subset(summary, coefficient == "Fleiss_exch" & interval == "wald_t")
if (nrow(exch)) {
  exch <- exch[order(exch$dgp, exch$mechanism, exch$method_label, exch$weight_label), ]
  exch_tab <- data.frame(
    DGP = exch$dgp,
    Mechanism = exch$mechanism,
    Method = exch$method_label,
    Weight = exch$weight_label,
    Valid = exch$n_valid,
    Failures = exch$failures,
    Bias = fmt_signed(exch$bias, 3),
    RMSE = fmt(exch$rmse, 3),
    `Mean SE` = fmt(exch$mean_se, 3),
    `Fit ms` = fmt(exch$median_elapsed_ms, 0),
    check.names = FALSE
  )
  knitr::kable(exch_tab, align = "llllrrrrrr")
} else {
  cat("No exchangeable Fleiss-counts rows in this run.\n")
}
```

## Interval Check

```{r interval-check}
int <- subset(summary, coefficient == "Conger" & weight_label == "quadratic")
int <- aggregate(
  cbind(coverage95, mean_ci_length, miss_below, miss_above) ~ interval + method_label,
  int,
  mean,
  na.rm = TRUE
)
int <- int[order(int$method_label, int$interval), ]
int_tab <- data.frame(
  Method = int$method_label,
  Interval = int$interval,
  Coverage = fmt(int$coverage95, 2),
  Length = fmt(int$mean_ci_length, 3),
  `Miss Below` = fmt(int$miss_below, 2),
  `Miss Above` = fmt(int$miss_above, 2),
  check.names = FALSE
)
knitr::kable(int_tab, align = "llrrrr")
```

## Reproduce

```bash
cd experiments/studies/29-kappa-planned-missingness

Rscript run_experiment.R --smoke --reps 5 --progress \
  --out-dir results/smoke-YYYYMMDD-HHMM

KAPPA29_RESULTS_DIR=results/smoke-YYYYMMDD-HHMM quarto render report.qmd
```
